### 2022-07-21 SIOP_Onet_FA.Rmd
FA with 2, 7, 8, 9 factors

---
title             : "O*Net Factor Analysis Project"
shorttitle        : "SIOP_FA"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

# Methods

## Participants

## Material

## Procedure

# Results

Maybe try hierarchical which 2 overaching factors with the "dimensions" under each).

We are working on a factor analysis of the characteristics themselves. This is based on the larger number of items (97?) with an n of now 568.

We tried different numbers of factors based on a scree plot suggesting we start at 8 factors. We printed only those loadings at .3 or higher to begin. Very few items loaded on more than 1 factor, and most loaded at .3+ on one factor. A few were negative. Need to think carefully about interpretation by exploring what the items themselves are.

Next - tried a 2 factor solution, wondering if items in the "context" and "activity" categories loaded as expected based on Onet categories. Again, most items loaded at .3+. The physical items generally loaded on factor 1.

```{r scree plot}

data <- read.csv("recodedforsiop.csv")[,-1] ## line 1562 of "study1.R" script

data$Structural <-      rowMeans(data[19:31], na.rm=TRUE)
data$Physical <-        rowMeans(data[c(32:60,62)], na.rm=TRUE)
data$Interpersonal <-   rowMeans(data[63:76], na.rm=TRUE)
data$Information.I <-   rowMeans(data[77:81], na.rm=TRUE)
data$Mental.P <-        rowMeans(data[82:91], na.rm=TRUE)
data$WorkOutput <-      rowMeans(data[92:100], na.rm=TRUE)
data$Interact.Others <- rowMeans(data[101:117], na.rm=TRUE)

library(psych)
forfa <- cor(data[c(19:60,62:117)], use="pairwise.complete.obs")
fa.parallel(forfa, n.obs=568)
```

```{r eight_factorFA}
##Here's where we started - it appears that based on scree plot, 8 factors might be appropriate.
efa <- fa(data[c(19:60,62:117)], rotate="oblimin", nfactors=8, n.obs=568, scores=TRUE, alpha=.1) ## changed to data so can get factor scores
summary(efa)

together <- cbind(data, efa$scores)

#res.cor <- correlate(together[c(405:413)], use="pairwise.complete.obs")
#res.cor %>% fashion()
#res.cor %>% shave(upper=FALSE) %>% rplot()

efa$loadings
print(efa$loadings,cut=.3,digits=2)
```

```{r two_factorFA}
##Next, trying the 2 overarching factors we noted first:
efa <- fa(forfa, rotate="oblimin", nfactors=2, n.obs=568, scores=TRUE, alpha=.1)
summary(efa)

efa$loadings
print(efa$loadings,cut=.3,digits=2)
```

```{r seven_factorFA}
## Try 7 factor solution next. 
efa <- fa(forfa, rotate="oblimin", nfactors=7, n.obs=568, scores=TRUE, alpha=.1)
summary(efa)

efa$loadings
print(efa$loadings,cut=.3,digits=2)
```

```{r nine_factorFA}
## Try 9 factor solution next. 
efa <- fa(forfa, rotate="oblimin", nfactors=9, n.obs=568, scores=TRUE, alpha=.1)
summary(efa)

efa$loadings
print(efa$loadings,cut=.3,digits=2)
```
# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup