### 2022-08-15 SIOP_Onet_FA.Rmd
New .Rmd for JDR interaction

---
title             : "O*Net Factor Analysis Project"
shorttitle        : "SIOP_FA"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib", "SIOP_Onet_FA.bib", "jdr.bib", "jobdemands_resources_bibtex.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

csl               : apa7.csl
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

Given the popularity of the Job Demands-Resources Theory [JD-R; @demerouti2001job] in exploring questions related to everything from motivation to job design, we aim to explore the intersection between *perceptions* of job demands and resources, and the broad set of job characteristics provided on O\*Net. This project makes three contributions. We aim to first explore whether ratings of of O\*Net item groupings align with the stated "resources" and "demands" presented in the job demands-resources theory. We then present evidence documenting whether O\*Net job and task descriptors are similarly rated as resources, challenge- or hindrance demands, and lastly, whether such ratings differ across job categories/classifications. 
Across two studies, a series of evaluations were made that used: 1) direct O\*Net terminology (both descriptor and response option), and 2) JD-R influenced ratings of demand, challenge, or hindrance of different types of workers. Prior to a description of results, a brief overview of both the JD-R theory, the stress appraisal process, and O\*Net is provided.

## The Job demands-Resources Theory

The overarching context for this study is that of the job demands-resources theory, which is an expansion of the well-studied job demands-resources model [@demerouti2001job]. One of the major advantages of the job demands-resources theory is that it allows us to model both work environment and job characteristics via job resources and demands. *Resources* include physical, psychological, social, or organizational aspects of the job that may help an employee achieve work goals, reduce job demands, or promote personal growth and development [@demerouti2001job]. In contrast, *demands* include components of a job that require sustained effort, and as such, produce psychological or physiological strain (e.g., high work pressure is frequently cited as a common demand; @demerouti2001job). 
Cognitively, the perception of an element of ones job as a resource or demand activates one of two distinct processes: either health impairment (resulting from demands) or motivation [resulting from resources; @bakker2014job]. Pertinent to the current study, demanding job characteristics are frequently associated with negative outcomes [e.g., @bakker2003dual], whereas job characteristics deemed resources have been associated with positive organizational outcomes like engagement and motivation [@bakker2007job].

## Objective vs. Subjective Nature of Demands and Resources: The Role of Appraisal

@searle2015merits note that the majority of the research on workplace demands is based on apriori classifications of demands. However, the stress experience, or process, described early on by @lazarus1984stress is grounded in the assumption that individual appraisals of stressors/demands vary. Their transactional theory of stress and coping states that people continuously appraise stimuli in their environments. An appraisal is the cognitive process whereby meaning is assigned to a stimulus. If a stimulus is appraised as a stressor (threat, challenge, potentially harmful), emotional distress leads to coping of some kind. This action to cope is also associated with another appraisal about the outcome itself and the process continues if the outcomes is not appraised as favorable [@lazarus1984stress]. The stress appraisal process suggests that classifying a job characteristic or environmental condition as an objective demand or resource might be in error.

We next consider the (limited) empirical evidence on this topic. First, some relatively recent research suggests that job demands and resources may not be universally appraised or assigned as such. Starting with job demands, @webster2011extending, for example, studied workload, role ambiguity, and role conflict demands, and found that while each could be appraised primarily as a challenge or hindrance demand, they could also simultaneously be perceived as being both a challenge and hindrance to different degrees. While their study did include resources, it nonetheless points to individual differences on how people perceive stressors at work. Although part of a much larger study on retirement, @sonnega_comparison_2018 compared self-reported (subjective) ratings of degree of physical demand, stress, and need for intense concentration from the Health and Retirement Study with objective ratings from O*Net. Correlations physical demand (r = .52), stress (r = .10), and need for intense concentration (r = .14), again suggesting perhaps that our objective ratings of job demands (and resources) may be subject to a greater level of individual difference than assumed. Next considering resources, @schmitz_interpreting_2019 also captured subjective and objective resources in their study of retirement. Correlations of composite variables for the resources of autonomy (r = .12. p > .01), recognition of work (r = .07, p > .01), and decision freedom (r = .08, p > .01), while significant, certainly do not reflect high levels of overlap.

We do acknowledge as well, that demands and resources are not necessarily consistent across days, or seasons, for many employees. @downes2021incorporating meta-analysis addresses this reality in depth, although it is beyond the scope of this project.

## O\*Net Resource

Originally, the Advisory Panel for the Dictionary of Occupational Titles recommended a system that would “…promote the effective education, training, counseling, and employment of the American workforce. It should accomplish its purpose by providing a database system that identifies, defines, classifies, and describes occupations in the economy in an accessible and flexible manner” (@advisory1993new, p. 6). The result was the now commonly used O\*NET. The Occupational Information Network (O\*NET; onetonline.org) contains a comprehensive description of occupations [@peterson2001understanding]. This widely accessed database houses hundreds of standardized and occupation-specific descriptors most occupations in the US and these descriptions are continually updated. In fact, there was a call to work with experienced I/O psychologists over the summer to update the content for the [Industrial and Organizational Psychologist listing on O\*Net](https://www.onetonline.org/link/summary/19-3032.00). These data, and the tools provided for free on the website (e.g., Career Exploration Tools, “My Next Move for Veterans”, “My Next Move”, Toolkit for Business) are frequently used by counselors, students, human resources departments, and researchers to assist potential applicants discover the skills and training they need for the job of their choice. It is also useful to employers by providing them with information with which to craft job descriptions and help employees determine what skills are needed for promotion.

Of greatest interest here are statements taken from O\*NET ["activity" and "context" classifications](https://www.O\*NETonline.org/find/descriptor/result/4.A.1.b.3) (e.g., items related to information input, interacting with others, physical work conditions, structural job characteristics). One of the first and basic questions is whether or not the categorical examples of "resources" and "demands" described in the Job Demands-Resources Theory [@demerouti2001job], for example, are generally deemed resources or demands as we objectively define them. The next logical question surrounds how "universal" such ratings are. For instance, it is quite possible, given the theoretical and empirical evidence presented above, that there is wide variability in individual appraisal of work activities and context such that some people may rate a given activity as a resource and others a hindrance. A second study extends the findings from Study 1 to a potentially key moderator - job categories/classifications, examining whether ratings of resources, challenge- and hindrance demands differ by job classification.

# Methods

## Participants

```{r getdata}

data <- read.csv("SIOPcleaned.csv")      ## from "datacleaning.R" script

## O*NET categories

data$onet.resource.ii <- rowMeans(data[c(162:166)], na.rm=TRUE)
data$onet.resource.mp <- rowMeans(data[c(167:176)], na.rm=TRUE)
data$onet.resource.wo <- rowMeans(data[c(177:185)], na.rm=TRUE)
data$onet.resource.io <- rowMeans(data[c(186:202)], na.rm=TRUE)
data$onet.resource.ir <- rowMeans(data[c(149:161)], na.rm=TRUE)
data$onet.resource.pc <- rowMeans(data[c(128:144,146:148)], na.rm=TRUE)
data$onet.resource.sc <- rowMeans(data[c(118:127)], na.rm=TRUE)

data$overall.resource <- rowMeans(data[c(118:144,147:202)], na.rm=TRUE)

data$onet.hindrance.ii <- rowMeans(data[c(246:247,249:252)], na.rm=TRUE)
data$onet.hindrance.mp <- rowMeans(data[c(252:261)], na.rm=TRUE)
data$onet.hindrance.wo <- rowMeans(data[c(262:270)], na.rm=TRUE)
data$onet.hindrance.io <- rowMeans(data[c(271:287)], na.rm=TRUE)
data$onet.hindrance.ir <- rowMeans(data[c(233:245)], na.rm=TRUE)
data$onet.hindrance.pc <- rowMeans(data[c(213:232)], na.rm=TRUE)
data$onet.hindrance.sc <- rowMeans(data[c(203:212)], na.rm=TRUE)

data$overall.hindrance <- rowMeans(data[c(203:247,249:287)], na.rm=TRUE)

data$onet.challenge.ii <- rowMeans(data[c(332:336)], na.rm=TRUE)
data$onet.challenge.mp <- rowMeans(data[c(337:346)], na.rm=TRUE)
data$onet.challenge.wo <- rowMeans(data[c(347:355)], na.rm=TRUE)
data$onet.challenge.io <- rowMeans(data[c(356:372)], na.rm=TRUE)
data$onet.challenge.ir <- rowMeans(data[c(319:331)], na.rm=TRUE)
data$onet.challenge.pc <- rowMeans(data[c(298:307,309:318)], na.rm=TRUE)
data$onet.challenge.sc <- rowMeans(data[c(288:297)], na.rm=TRUE)

data$overall.challenge <- rowMeans(data[c(288:307,309:372)], na.rm=TRUE)

data$burnout      <- rowMeans(data[c(373:376)], na.rm=TRUE)
data$stress       <- rowMeans(data[c(377:379)], na.rm=TRUE)
data$engagement   <- rowMeans(data[c(380:399)], na.rm=TRUE)        ## added 10/12 because analyses too wonky with 3 subscales

```

There were `r nrow(data)` respondents.

```{r PROCESS}

#Next, try a moderated regression

# PROCESS for R version 4.1.1
# Written by Andrew F. Hayes
# www.afhayes.com
# www.processmacro.org
# Copyright 2012-2022 by Andrew F. Hayes ALL RIGHTS RESERVED
# PROCESS workshop schedule at http://haskayne.ucalgary.ca/CCRAM
#
# Distribution of this code in any form except through processmacro.org 
# is prohibited without the permission of the copyright holder, as is
# distribution after modification.
#
# THIS SOFTWARE IS PROVIDED AS IS, WITHOUT WARRANTY OF ANY KIND
# EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
# MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT
# IN NO EVENT SHALL THE COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM
# DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT
# OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE
# SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE
# USE OF THIS SOFTWARE IMPLIES AGREEMENT WITH THESE TERMS
#
# To activate, run this script. It may take a few minutes. 
# Patience is a virtue. This will produce a function called process.
# The other functions it creates cannot be accessed by the user
# but are used by the master process function.

process.bcboot3<-function(databcbt,estmte,xp2,badend,priorlo,priorhi)
{
  databcbt<-as.matrix(sort(databcbt))
  badlo<-0;badhi<-0
  pv<-matrix(as.numeric(databcbt < estmte));
  pv<-sum(pv)/nrow(databcbt);ppv<-pv;
  if (pv > .5){ppv<-(1-pv)}
  y5<-sqrt(-2*log(ppv))
  p0<-(-.322232431088);p1<-(-1);p2<-(-.342242088547);p3<-(-.0204231210245)
  p4<-(-.0000453642210148);q0<-(.0993484626060);q1<-(.588581570495)
  q2<-(.531103462366);q3<-(.103537752850);q4<-(.0038560700634)
  xp<-y5+((((y5*p4+p3)*y5+p2)*y5+p1)*y5+p0)/((((y5*q4+q3)*y5+q2)*y5+q1)*y5+q0)
  if (pv <= .5){xp<-(-xp)}
  cilow<-round(nrow(databcbt)*pnorm(2*xp-xp2))
  cihigh<-trunc(nrow(databcbt)*pnorm(2*xp+xp2))+1
  if (cilow < 1){cilow<-1;booterr<-1;badlo<-1}
  if (cihigh > nrow(databcbt)){cihigh<-nrow(databcbt);booterr<-1;badhi<-1}
  llcit<-databcbt[cilow,1]
  ulcit<-databcbt[cihigh,1]
  if ((badlo==1) & (llcit != priorlo)){priorlo<-llcit;badend<-c(badend,llcit)}
  if ((badhi==1) & (ulcit != priorhi)){priorhi<-ulcit;badend<-c(badend,ulcit)}
  bootse<-sd(databcbt)
  cires<-as.matrix(c(bootse,llcit,ulcit))
  cires<-list(cires,badend,priorlo,priorhi)
  return(cires)
}

process.pboot3<-function(databcbt,lcval,hcval)
{
  databcbt<-as.matrix(sort(databcbt))
  llcit<-databcbt[lcval,1]
  ulcit<-databcbt[hcval,1]
  bootse<-sd(databcbt)
  cires<-as.matrix(c(bootse,llcit,ulcit))
  return(cires)
}

process.llrtest3<-function(lm,y,x,b,basemod,iterate,converge)
{
  lm<-as.matrix(lm)
  btemphld<-b
  llrdat<-matrix(-999,nrow(x),(nrow(lm)-sum(lm)))
  llrdf<-ncol(x)-ncol(llrdat)
  llrcnt<-0
  for (llri in (1:nrow(lm)))
  {
    if (lm[llri,1]==0){llrcnt<-llrcnt+1;llrdat[,llrcnt]<-x[,llri]}
  }
  LL2<-process.modelest(y,llrdat,2,0,xp2,5,iterate,converge)
  b<-btemphld
  pvchi<-(1-pchisq((LL2-basemod),df=llrdf))
  fresult<-cbind((LL2-basemod),llrdf,pvchi)
  return(fresult)
}

process.describ3<-function(descdatf,type=0,quantle=1)
{
  desctmp<-matrix(-999,(8-(4*type)),ncol(descdatf))
  # mean, sd, min, max, 16th, 50th, 84th, dich toggle
  for (jd in c(1:ncol(descdatf)))
  {
    descdat<-descdatf[,jd]
    #get the mean, sd, minimum, and maximum */
    desctmp[1,jd]<-mean(descdat)
    desctmp[2,jd]<-sd(descdat)
    desctmp[3,jd]<-min(descdat)
    desctmp[4,jd]<-max(descdat)
    if (type==0)
    {
      minwarn<-0;maxwarn<-0
      tmp=as.numeric(descdat==desctmp[3,jd])+as.numeric(descdat==desctmp[4,jd])
      desctmp[8,jd]<-as.numeric(sum(tmp)==length(tmp))
      if (desctmp[3,jd]==desctmp[4,jd]){desctmp[8,jd]<-2}
      descdat<-matrix(sort(descdat))
      decval<-c(.16,.5,.84)
      for (kd in c(1:3))
      {
        low<-trunc(decval[kd]*(length(descdat)+1))
        lowdec<-decval[kd]*(length(descdat)+1)-low
        value<-descdat[low,1]+(descdat[(low+1),1]-descdat[low,1])*lowdec
        desctmp[(4+kd),jd]<-value
      }
      mnotev<-(1)
      modvals<-matrix(desctmp[5:7,],ncol=ncol(descdatf))
      if (quantle != 1)
      {
        desctmp[5,jd]<-desctmp[1,jd]-desctmp[2,jd]
        desctmp[6,jd]<-desctmp[1,jd]
        desctmp[7,jd]<-desctmp[1,jd]+desctmp[2,jd]
        modvals<-matrix(desctmp[5:7,],ncol=ncol(descdatf))
        mnotev<-(2)
        if (modvals[1,1] < desctmp[3,1]){modvals[1,1]<-desctmp[3,1];minwarn<-1}
        if (modvals[3,1] > desctmp[4,1]){modvals[3,1]<-desctmp[4,1];maxwarn<-1}
      }
      if (desctmp[8,jd]==1)
      {modvals<-matrix(c(desctmp[3,1],desctmp[4,1]))
      mnotev<-0;minwarn<-0;maxwarn<-0
      }
      descrtrn<-list(desctmp,modvals,minwarn,maxwarn,mnotev)
    }
  }
  if (type==1)
  {descrtrn<-list(desctmp)}
  return(descrtrn)
}

process.ftest3<-function(lm,bcoef,cv=0,chr=0,brsq=0,skip=0,y,x)
{
  lmat2<-as.matrix(lm)
  y<-as.matrix(y)
  x<-as.matrix(x)
  n<-nrow(y)
  if (skip==0)
  {
    lmat2<-as.matrix(diag(as.numeric(lm)))
    lmat3<-matrix(0,nrow(lmat2),1)
    for (flp in c(1:ncol(lmat2)))
    {
      if (sum(lmat2[,flp])==1)
      {lmat3<-cbind(as.matrix(lmat3),as.matrix(lmat2[,flp]))}
    }
    lmat2<-as.matrix(lmat3[,2:ncol(lmat3)])
  }
  fratio<-(t(t(lmat2)%*%bcoef)%*%solve(t(lmat2)%*%cv%*%lmat2)%*%((t(lmat2)%*%bcoef)))/ncol(lmat2)
  pfr<-(1-pf(fratio,ncol(lmat2),(n-nrow(bcoef))))
  fresult<-matrix(c(fratio,ncol(lmat2),(n-nrow(bcoef)),pfr),ncol=4)
  if (chr==1)
  {
    lmat3<-as.matrix(1-rowSums(lmat2))
    xfm<-matrix(0,n,sum(lmat3))
    flpc<-1
    for (flp in (1:nrow(lmat3)))
    {
      if (lmat3[flp,1]==1){xfm[,flpc]=x[,flp];flpc<-flpc+1}
    }
    bfm<-solve(t(xfm)%*%xfm)%*%t(xfm)%*%y
    resid<-y-(xfm%*%bfm)
    sstotal<-t(y-(sum(y)/n))%*%(y-(sum(y)/n))
    ssresid<-t(resid)%*%resid
    rsqch<-as.numeric(brsq)-((sstotal-ssresid)/sstotal)
    fresult<-matrix(c(rsqch,fresult),ncol=5)
  }
  ftestout<-as.matrix(fresult)
  return(ftestout)
}

#type1 for ols, type2 for logistic LLR, type3 for logistic bootstrapping
process.modelest<-function(y,x,type,full,xp2,hc=5,iterate=100,converge=.00001)
{
  if (type==1)
  {
    invxtx<-solve(t(x)%*%x)
    b<-invxtx%*%t(x)%*%y
    modres<-b
    if (full==1)
    {
      n1<-nrow(x)
      dfres<-(n1-(ncol(x)))
      sstotal<-t(y-(sum(y)/n1))%*%(y-(sum(y)/n1))
      resid=y-x%*%b
      ssresid<-sum(t(resid)%*%resid)       
      r2<-(sstotal-ssresid)/sstotal
      adjr2<-(1-((1-r2)*(n1-1)/(dfres)))
      mse<-ssresid/(n1-ncol(x))
      #HC covariance matrix
      varb<-mse*invxtx
      k3<-ncol(x)
      xhc<-0
      if (hc != 5)
      {
        xhc<-x
        hat<-matrix(xhc[,1])
        for (i3 in c(1:nrow(xhc)))
        {
          xhcm<-matrix(xhc[i3,])
          hat[i3,1]<-t(xhcm)%*%invxtx%*%xhcm
        }     
        if ((hc==0) | (hc==1))
        {
          for (i3 in c(1:k3)){xhc[,i3]<-xhc[,i3]*resid}
        }
        if ((hc==3) | (hc==2))
        {
          for (i3 in c(1:k3))
          {xhc[,i3]<-(resid/(1-hat)^(1/(4-hc)))*xhc[,i3]}
        }
        if (hc==4)
        {
          hcmn<-matrix(4,n1,2);hcmn[,2]<-(n1*hat)/k3
          minr<-apply(hcmn,1,FUN=min)
          for (i3 in c(1:k3))
          {
            xhc[,i3]<-(resid/(1-hat)^(minr/2))*xhc[,i3]
          }
        }
        varb<-(invxtx%*%t(xhc)%*%xhc%*%invxtx)
        if (hc==1){varb<-(n1/(n1-ncol(x)))*varb}
      }
      seb<-sqrt(diag(varb))
      trat<-b/seb
      p<-2*pt(-abs(trat),df=dfres)
      tval<-sqrt(dfres* (exp((dfres-(5/6))*((xp2/(dfres-(2/3)+(.11/dfres)))*(xp2/(dfres-(2/3)+(.11/dfres)))))-1))
      modres<-matrix(c(modres,seb, trat,p,(b-tval*seb),(b+tval*seb)),ncol=6)
      modresl<-t(matrix(c("coeff","hclab","t","p","LLCI","ULCI")))
      lmat<-diag(ncol(x));lmat<-lmat[,2:ncol(lmat)]
      fratio<-(t(t(lmat)%*%b)%*%solve(t(lmat)%*%varb%*%lmat)%*%((t(lmat)%*%b)))/(ncol(x)-1)
      pfr<-1-pf(fratio,(ncol(x)-1),dfres)
      modsum=matrix(c(sqrt(r2),r2,mse,fratio,(ncol(x)-1),dfres,pfr))
      modsuml=matrix(c("R","R-sq","MSE","hcflab","df1","df2", "p"))
      modretrn<-list(modres,modresl,modsum,modsuml,b,varb,tval,resid)
      return(modretrn)
    }
    if (full==0){return(modres)} 
  }
  #for logistic Y model
  if ((type==2) | (type==3))
  {
    xlp<-x;ylp<-as.matrix(y)
    pt2<-matrix((sum(ylp)/nrow(ylp)),nrow(ylp),1)
    if ((type==2)|(type==3)) {LL3<-(ylp*log(pt2)+(1-ylp)*log(1-pt2))}    
    LL3<-(-2*sum(LL3))
    bt1<-matrix(0,ncol(xlp),1);LL1<-0
    pt1<-matrix(0.5,nrow(ylp),1);pt1lp<-pt1
    for (jjj in (1:iterate))
    {
      xlptmp<-t(xlp)
      vecprb<-(pt1lp*(1-pt1lp))
      for (kkk in (1:ncol(xlp))){xlptmp[kkk,]<-xlptmp[kkk,]*t(vecprb)}
      b<-bt1+solve(xlptmp%*%xlp)%*%t(xlp)%*%(ylp-pt1lp)
      if ((type==2) | (type==3))
      {
        xlpb<-xlp%*%b
        xlpbt<-as.numeric(xlpb > -709.7)
        xlpb709<-((1-xlpbt)*(-709.7))
        xlpb<-((xlpb*xlpbt)+xlpb709)
        pt1lp<-1/(1+exp(-(xlpb)))
      }
      itprob<-sum((pt1lp < .00000001) | (pt1lp > .9999999))
      if (itprob > 0)
      {
        for (kkk in (1:nrow(pt1lp)))
        {
          if (pt1lp[kkk,1] > .9999999){pt1lp[kkk,1]<-.9999999}     
          if (pt1lp[kkk,1] < .00000001){pt1lp[kkk,1]<-.00000001}     
        }
        itprob<-0
      }
      if (itprob==0)
      {
        if ((type==2)|(type==3)){LL<-(ylp*log(pt1lp)+(1-ylp)*log(1-pt1lp))}
        LL2<-(-2*sum(LL))
      }
      if (abs(LL1-LL2) < converge)
      {
        if (full==1)
        {
          xlptmp<-t(xlp)
          vecprb<-(pt1lp*(1-pt1lp))   
          for (kkk in (1:ncol(xlp))){xlptmp[kkk,]<-xlptmp[kkk,]*t(vecprb)} 
          varb<-solve(xlptmp%*%xlp)
          seb<-matrix(sqrt(diag(varb)))
        }
        break
      }
      bt1<-b;LL1<-LL2
    }
    modres<-b
    if (jjj > iterate)
    {
      itprob<-2
      if (booting==0){iterrmod<-1}   
      if (booting==1){bootiter==1}   
      if (itprobtg==0)
      {   
        itprobtg<-1;errcode[errs,1]<-47;errs<-errs+1
        if ((booting==0) & (full==1))
        {
          pt1lpc<-(pt1lp*(1-pt1lp))
          vt1<-diag(pt1lpc)
          varb<-solve(t(xlp)%*%vt1%*%xlp)
          seb<-matrix(sqrt(diag(varb)))
        }
      }
    }
    if (full==1)
    {
      trat<-b/seb
      dfres<-nrow(xlp)   
      p<-2*(1-pnorm(abs(trat)))
      modres<-cbind(modres,seb,trat,p) 
      modres<-matrix(c(modres,(b-xp2*seb),(b+xp2*seb)),ncol=6)
      pvchi<-(1-pchisq((LL3-LL2),df=(nrow(modres)-1)))
      mcF<-(LL3-LL2)/LL3
      cox<-1-exp(-(LL3-LL2)/nrow(xlp))
      nagel<-cox/(1-exp(-(LL3)/nrow(xlp)))
      modsum<-matrix(c(LL2,(LL3-LL2),(nrow(modres)-1),pvchi, mcF,cox,nagel))
      modsuml<-matrix(c("-2LL","ModelLL", "df", "p", "McFadden", "CoxSnell", "Nagelkrk"))
      modresl<-t(matrix(c("coeff","se","Z","p","LLCI","ULCI")))
      modretrn<-list(modres,modresl,modsum,modsuml,b,varb,xp2)
      return(modretrn)
    }
    if ((full==0) & (type==2)){return(LL2)} 
    if ((full==0) & (type==3)){return(modres)}
  }
}

process.dummy3<-function(dd)
{
  uq<-unique(dd)
  uq<-matrix(unlist(uq[1]))
  dummy<-matrix(0,nrow(dd),nrow(uq))
  uq<-matrix(sort(uq))
  uq2<-matrix(seq(1:nrow(uq)))
  for (i in c(1:nrow(dd)))
  {for (j in c(1:nrow(uq)))
  {if (dd[i,1]==uq[j,1])
  {dummy[i,uq2[j,1]]<-1}  
  }
  }
  return(dummy)
}

process.makdummy<-function(dd,method,custcov=0,custcode=999)
{
  dd<-dd[order(dd[,2]),]
  newrow<-dd[,1]
  dd<-dd[,2]
  dd<-as.data.frame(dd)
  dummy<-process.dummy3(dd)
  uq<-unique(dd);uq<-matrix(unlist(uq[1]))
  criterrd<-0;errcode5<-0;errcode4<-0;errcode6<-0
  nvls<-ncol(dummy)
  nnvls<-colSums(dummy)
  mnvls<-min(nnvls)
  nnvls<-matrix(sort(uq))
  conmat1<-1
  if (mnvls < 2){errcode5<-1;criterrd<-1}
  if (nvls > 9){errcode4<-1;criterrd<-1}
  x<-0;dummat<-0
  if (criterrd==0)
  {
    dumok<-1
    if (method > 0)
    {
      x<-dummy[,2:ncol(dummy)]
      nx<-ncol(x)
      minus1<-matrix(-1,1,ncol(x))
      if (method==4)
      {
        for (k in (1:nrow(dd)))
        {shole<-sum(x[k,])
        if (shole==0)
        {x[k,]<-minus1}}
      }
      if ((method==2) | (method==3) | (method==5))
      {
        for (k in (1:nrow(dd)))
        {shole<-sum(x[k,])
        if (shole > 0)
        {
          for (i in (1:ncol(x)))
          {if (x[k,i]==0){x[k,i]<-1} else{break}
          }
        }
        }
        if (method==3)
        {
          conmat1=c(-8,1,1,1,1,1,1,1,1,0,-7,1,1,1,1,1,1,1,0,0,-6,1,1,1,1,1,1,0,0,0,-5,1,1,1,1,1,0,0,0,0,-4,1,1,1,1,0,0,0,0,0,-3,1,1,1,0,0,0,0,0,0,-2,1,1,0,0,0,0,0,0,0,-1,1)
          conmat1<-t(matrix(conmat1,9,8))
          for (i in (1:8)){conmat1[i,]=conmat1[i,]/(10-i)}
          conmat1<-t(conmat1[(10-nvls):8,(10-nvls):9])
          for (k in (1:nrow(dd)))
          {shole<-sum(x[k,]);x[k,]=conmat1[(shole+1),]}
        }
      }
      if (method==5)
      {
        if (ncol(custcode) != (nvls*(nvls-1)))
        {errcode6<-1}
        if (ncol(custcode)==(nvls*(nvls-1)))
        {
          conmat1<-matrix(0,nvls,(nvls-1));cnt<-1
          for (i in (1:nvls))
          {
            for (k in (1:(nvls-1)))
            {conmat1[i,k]<-custcode[1,cnt];cnt<-cnt+1}
          }
          for (k in (1:nrow(dd))){x[k,]<-conmat1[(sum(x[k,])+1),]}   #check this
        }
      }
      xskip<-1
      dummat<-matrix(0,(nx+1),nx)
      dummat[(2:nrow(dummat)),]<-diag(nx)
      if (method==4){dummat[1,]<-minus1}
      if (method==2)
      {for (i in 2:nrow(dummat))
      {for (j in (1:(i-1))){dummat[i,j]<-1}}
      }
      if (method==3){dummat<-conmat1}
      if ((method==5) & (errcode6==0)){dummat<-conmat1}
      dummat<-cbind(nnvls,dummat)
    } 
    x<-cbind(newrow,x)
    x<-x[order(x[,1]),]
  }
  dummrn<-list(x,dummat,nvls,nnvls,errcode4,errcode5,errcode6)
  return(dummrn)
}

process<-function(data,y="xxxxx",x="xxxxx",m="xxxxx",w="xxxxx",z="xxxxx",cov="xxxxx",
                  model=999,converge=.00001,iterate=100,hc=5,jn=0,effsize=0,stand=0,xmtest=0,normal=0,
                  mdichok=0,contrast=999,modelbt=0,matrices=0,covmy=0,covcoeff=0,boot=5000,mc=0,
                  intprobe=0.1,plot=0,total=0,save=0,mcx=0,mcw=0,mcz=0,moments=0,progress=1,
                  bmatrix=-999,wmatrix=-999,zmatrix=-999,wzmatrix=-999,cmatrix=-999,xcatcode=999,
                  wcatcode=999,zcatcode=999,wmodval=999,zmodval=999,center=0,conf=95,seed=-999,
                  decimals=9.4,maxboot=0,modelres=0,bc=0,outscreen=1,activate=0,describe=0,listmiss=0,
                  linsum=-999)
{

  #all this is initiation of variables and matrices
  ranseed<-0;
  if (seed != -999)
  {seed<-trunc(abs(seed));set.seed(seed)}
  wnames<-w;znames<-z;mcerpt<-0;wiscov<-0;ziscov<-0;itprobtg<-0;
  v2tag<-0;maxwwarn<-0;minwwarn<-0;maxzwarn<-0;minzwarn<-0
  toomany<-0;wdich<-0;zdich<-0;wnotev<-0;znotev<-0;singlr<-0
  nxpval<-1;nwpval<-1;nzpval<-1;errs<-1;notes<-1;criterr<-0
  novar<-0;adjust<-0;ncs<-0;serial<-0;sobelok<-0;hasw<-0;ydich<-0;
  hasz<-0;printw<-0;printz<-0;counterf<-0;wmodcust<-0;zmodcust<-0
  booting<-0;bootiter<-0;iterrmod<-0;model<-trunc(model)
  errcode<-matrix(0,100,1);notecode<-matrix(0,100,1)
  iterate<-abs(trunc(iterate));converge=abs(converge);badend<-0;booterr<-0;
  itprobtg<-0;v2tag<-0;maxwwarn<-0;minwwarn<-0;maxzwarn<-0
  minzwarn<-0;toomany<-0;wdich<-0;zdich<-0;wnotev<-0;znotev<-0
  nws<-0;nzs<-0;nms<-0;nys<-0;nxs<-0;maxresm<-9;bc<-as.numeric(bc==1);progress<-as.numeric(progress==1)
  mcxok<-0;mcwok<-0;mczok<-0;xprod<-0;zprod<-0;wprod<-0;modcok<-0;alttotal=0;bc<-as.numeric(bc==1)
  jn<-as.numeric(jn==1);effsize<-as.numeric(effsize==1);maxboots=abs(trunc(maxboot))
  normal<-as.numeric(normal==1);xmtest<-as.numeric(xmtest==1);modelres<-as.numeric(modelres==1)
  stand<-as.numeric(stand==1);outscreen<-as.numeric(outscreen==1);activate=as.numeric(activate==1)
  describe<-as.numeric(describe==1);listmiss<-as.numeric(listmiss==1);
  if (stand==1) {effsize<-1}
  pstog<-0;sobelok<-0;mdichok=as.numeric(mdichok==1)
  resultm<-matrix(99999,1,maxresm)

  linsum<-t(matrix(linsum));
  nlinsum<-ncol(linsum);
  if (linsum[1,1]==-999){nlinsum<-0}

  #contrast matrix
  contrast<-matrix(contrast);
  contrast<-t(contrast)
  ncontr<-ncol(contrast)
  if (contrast[1,1]==999) {ncontr<-1;contrast[1,1]<-0}
  if (ncontr==1)
  {
    contrast<-trunc(contrast)
    if ((contrast[1,1] > 2) | (contrast[1,1] < 0))
    {ncontr<-1;contrast[1,1]<-0}
  }
  if (ncontr > 1)
  {
    contvec<-contrast;contrast[1,1]<-3
    if (((model==2) | (model==3)) & (nms==0))
    {
      if (ncontr==4)
      {
        contvec<-t(matrix(contvec,2,2))
        contrast[1,1]<-0;modcok<-1;wcontval<-matrix(contvec[,1]);zcontval<-matrix(contvec[,2])
      }
      if (ncontr !=4)
      {
        notecode[notes,1]<-19;notes<-notes+1
        contrast[1,1]<-0
      }
    }
  }
  contrast<-contrast[1,1]
  modelbt<-as.numeric(modelbt==1);matrices<-as.numeric(matrices==1)
  covcoeff<-as.numeric(covcoeff==1);covmy<-trunc(covmy)
  if ((covmy < 0) | (covmy > 2)) {covmy<-0} 
  boot<-abs(trunc(boot));mc=abs(trunc(mc));hc=trunc(hc)
  if ((intprobe < 0) | (intprobe > 1))
  {intprobe<-0.10}
  plot<-trunc(plot)
  if ((plot < 0) | (plot > 2)) {plot<-0}
  total<-as.numeric(total==1)
  dototal<-0
  saveboot<-as.numeric((save==1)|(save==3))
  saveest<-as.numeric(save > 1)
  if (saveest==1){intprobe=1}
  if ((hc >=0) & (hc < 5)) {notecode[notes,1]=4;notes=notes+1}
  if ((hc > 5) | (hc < 0)) {hc=5}
  ####dont use these when y is dichotomous
  hclab<-matrix(c("se(HC0)","se(HC1)","se(HC2)","se(HC3)","se(HC4)","se"))
  hclab<-hclab[(hc+1),1]
  hcflab<-matrix(c("F(HC0)","F(HC1)","F(HC2)","F(HC3)","F(HC4)","F"))
  hcflab<-hcflab[(hc+1),1] 
  mcw=trunc(mcw);mcz=trunc(mcz);mcx=trunc(mcx);
  if ((mcx > 0) & (model==74)) {mcw=mcx}
  if ((mcw > 0) & (model==74)) {mcx=mcw}
  if ((mcx > 0) & (contrast > 0))
  {notecode[notes,1]=28;notes=notes+1;contrast=0};
  nxvls<-1;nmvls<-1;nwvls<-1;nzvls<-1
  paths<-matrix(999);pathsw<-matrix(999);pathsz<-matrix(999);
  pathswz<-matrix(999);pathsmod<-matrix(999);pathtype<-matrix(999);obscoeff<-999;
  pathsdv=matrix(" ");quantile<-1;moments=as.numeric(moments==1)
  if (moments==1){quantile<-0}
  bmatrix=t(matrix(bmatrix));wmatrix=t(matrix(wmatrix));zmatrix=t(matrix(zmatrix));
  wzmatrix=t(matrix(wzmatrix));cmatrix=t(matrix(cmatrix));xcatcode=t(matrix(xcatcode));
  wcatcode=t(matrix(wcatcode));zcatcode=t(matrix(zcatcode));
  needed<-0
  decimals<-paste("%",decimals,"f",sep='')
  if ((trunc(conf) >= 100) | (trunc(conf <= 50)))
  {conf=95;notecode[notes,1]=2;notes=notes+1}
  if ((model >= 0)  & (model < 4) & (modelbt==0)) {boot<-0;mc<-0;bc<-0;saveboot<-0}
  if ((boot > 0) & (mc > 0)) {boot<-0;bc<-0}
  if ((boot < 1000) & (boot > 0) & (mc==0)){boot=5000}
  if ((mc < 1000) & (mc > 0) & (boot==0)){mc=5000}

  #if ((model > 0) & (model < 4) & (boot = 98765) & (modelbt=1)) then;do;mc=0;boot=5000;end;
  #if ((model > 0) & (model < 4) & ((boot = 98765) | (mc > 0))) then;do;boot=0;mc=0;bc=0;end;
  #if ((boot = 98765) & (mc = 0)) then;do;boot=5000;end;

  p0<- -.322232431088;p1<- -1;p2<- -.342242088547;p3<- -.0204231210245;
  p4<- -.0000453642210148;q0<- .0993484626060;q1<- .588581570495;
  q2<- .531103462366;q3<- .103537752850;q4<- .0038560700634;
  priorlo<-(-9999999);priorhi<-9999999
  alpha2<-(1-(conf/100))/2;cilm<-alpha2*2;y5<-sqrt(-2*log(alpha2));
  xp2=(y5+((((y5*p4+p3)*y5+p2)*y5+p1)*y5+p0)/((((y5*q4+q3)*y5+q2)*y5+q1)*y5+q0))
  medlb=matrix(c("   M1 :","   M2 :","   M3 :","   M4 :","   M5 :","   M6 :","   M7 :","   M8 :","   M9 :","  M10 :"))
  medlb2=matrix(c("(M1)","(M2)","(M3)","(M4)","(M5)","(M6)","(M7)","(M8)","(M9)","(M10)"))
  xlb=matrix(c("X1 :","X2 :","X3 :","X4 :","X5 :","X6 :","X7 :","X8 :","X9 :","X10:"))
  highlbw=matrix(c("M1*W","M2*W","M3*W","M4*W","M5*W","M6*W","M7*W","M8*W","M9*W","M10*W"))
  highlbz=matrix(c("M1*Z","M2*Z","M3*Z","M4*Z","M5*Z","M6*Z","M7*Z","M8*Z","M9*Z","M10*Z"))
  highlbwz=matrix(c("M1*W*Z","M2*W*Z","M3*W*Z","M4*W*Z","M5*W*Z","M6*W*Z","M7*W*Z","M8*W*Z","M9*W*Z","M10*W*Z"))
  highlbx=matrix(c("M1*X","M2*X","M3*X","M4*X","M5*X","M6*X","M7*X","M8*X","M9*X","M10*X"))
  highlbbt=matrix(c("BOTH(M1)","BOTH(M2)","BOTH(M3)","BOTH(M4)","BOTH(M5)","BOTH(M6)","BOTH(M7)","BOTH(M8)","BOTH(M9)","BTH(M10)"))
  skipwz=0
  xnck=matrix(x);ynck=matrix(y);wnck=matrix(w);znck=matrix(z);mnck=matrix(m);cnck=matrix(cov)
  if ((nrow(xnck)>1) | (nrow(ynck)>1) | (nrow(wnck)>1) | (nrow(znck)>1))
  {errcode[errs,1]<-3;errs<-errs+1;criterr<-1}

  validm=matrix(c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0,0,1,1,0,0,0,0,
                  0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,0,1,1,0,0,0,1,1,1,1,1,1,1,1,1,1,1,1,1))
  dim(validm)<-c(1,92)

  # check for major errors in syntax
  if (activate==1)
  {errcode[errs,1]<-60;errs<-errs+1;criterr<-1}
  if (criterr==0)
  {
    if ((model > 0) & (model < 93))
    {
      if (validm[1,model]==0)
      {
        errcode[errs,1]<-6;errs<-errs+1;criterr<-1
      }
    }
    if (((model > 92) | (model < 0)) & (model != 999))
    {errcode[errs,1]<-7;errs<-errs+1;criterr<-1}
    if ((model==999) & (bmatrix[1,1]==-999))
    {errcode[errs,1]<-24;errs<-errs+1;criterr<-1}
    if ((model != 999) & (bmatrix[1,1] != -999))
    {errcode[errs,1]<-25;errs<-errs+1;criterr<-1}
    if (((model==74) | ((model > 0) & (model < 4))) & ((wmatrix[1,1] !=-999) | (zmatrix[1,1] != -999) | (wzmatrix[1,1] != -999)))
    {errcode[errs,1]<-41;errs<-errs+1;criterr<-1}
    if ((y=="xxxxx") | (x=="xxxxx"))
    {errcode[errs,1]<-1;errs<-errs+1;criterr<-1}
    if ((mnck[1,1]=="xxxxx") & (model > 3))
    {errcode[errs,1]<-8;errs<-errs+1;criterr<-1}
  } 
  #end of check for major errors in syntax

  #A loop reads data and does some other things

  if (criterr==0)
  {
    #read y data
    ytmp<-data[y];nys<-ncol(ytmp);needed<-nys
    ynames<-matrix(y);n<-nrow(ytmp);varnames<-matrix(ynames)
    dat<-ytmp
    modelvar<-matrix(c(model,ynames))
    if (model==999){modelvar[1,1]="CUSTOM"}

    #read x data
    xtmp<-data[x];
    nxs<-ncol(xtmp);needed<-needed+nxs;
    dat<-cbind(dat,xtmp)
    xnames<-matrix(x);n<-nrow(xtmp);xcatlab<-xnames;
    varnames<-matrix(c(varnames,xnames))
    modelvar<-matrix(c(modelvar,xnames))
    if (nxs==1){modelvlb<-matrix(c("Model :","    Y :","    X :"))}

    #read m data
    if (mnck[1,1] != "xxxxx")
    {
      mtmp<-data[m]
      nms<-ncol(mtmp)
      mnames<-matrix(m)
      mprod<-matrix(0,1,nms)
      n<-nrow(mtmp)
      needed<-needed+nms
      varnames<-matrix(c(varnames,mnames))
      dat<-cbind(dat,mtmp)
      modelvar<-matrix(c(modelvar,mnames))
      x2m<-matrix(0,99,nms)
      x2m<-matrix(0,99,nms)
      x2y<-matrix(0,99,nms)
      onem<-matrix(1,nms,1)
      if ((nms > 1) & (nms < 11))
      {modelvlb<-matrix(c(modelvlb,medlb[1:nms,1]))}
      if (nms==1)
      {modelvlb<-matrix(c(modelvlb,"    M :"))}
      if ((nms > 0) & (model < 4)){errcode[errs,1]<-9;errs<-errs+1;criterr<-1}
    }

    #read w data
    wlocatet<-0;wlocate<-0
    if (w != "xxxxx")
    {
      wtmp<-data[w];nws=ncol(wtmp);n<-nrow(wtmp)
      dat<-cbind(dat,wtmp)
      wnames<-matrix(w);wcatlab<-wnames;
      varnames<-matrix(c(varnames,wnames))
      modelvar<-matrix(c(modelvar,wnames))
      modelvlb<-matrix(c(modelvlb,"    W :"))
      wlocate<-nrow(varnames)
      if (model==74)
      {
        wlocatet<-1;
        if (xnames != wnames)
        {errcode[errs,1]<-45;errs<-errs+1;criterr<-1}
      }     
    }

    #read z data
    if (z != "xxxxx")
    {
      ztmp<-data[z];nzs<-ncol(ztmp);n<-nrow(ztmp);
      dat<-cbind(dat,ztmp)
      znames<-matrix(z);zcatlab<-znames;
      varnames<-matrix(c(varnames,znames))
      modelvar<-matrix(c(modelvar,znames))
      modelvlb<-matrix(c(modelvlb,"    Z :"))
    }

    #read cov data
    if (cnck[1,1] != "xxxxx")
    {
      ctmp<-data[cov];ncs<-ncol(ctmp);n<-nrow(ctmp);
      dat<-cbind(dat,ctmp)
      covnames<-matrix(cov);varnames<-matrix(c(varnames,covnames));  
    }

    if (((model==80) | (model==81)) & ((nms < 3) | (nms > 6)))
    {errcode[errs,1]<-32;errs<-errs+1;criterr<-1}
    if ((model==82) & (nms != 4))
    {errcode[errs,1]<-33;errs<-errs+1;criterr<-1}
    if (nms > 10)
    {errcode[errs,1]<-37;errs<-errs+1;criterr<-1}
    if (((model==6) | ((model > 82) & (model < 999))) & ((nms < 2) | (nms > 6)))
    {errcode[errs,1]<-34;errs<-errs+1;criterr<-1}

    #check for redundant variable names
    match<-0;match2<-0;mcwzcov<-0
    for (i in c(1:(nrow(varnames)-1)))
    {
      for (j in c((i+1):nrow(varnames)))
      {
        if (varnames[i]==varnames[j])
        {
          if (i < (nxs+nms+nys+1)){match2<-match2+1}
          if ((wlocate==1) & (i==2) & (j==wlocate)){match2<-match2-1}
          if ((wnames==znames) & ((nws > 0) | (nzs > 0))){match<-match2+1} 
          if ((i < (nrow(varnames)-ncs+1)) & (j > (ncol(varnames)-ncs)))
          {
            if ((varnames[j]==wnames) & (mcw==0))
            {match<-0;wiscov<-(j-(nrow(varnames)-ncs))}
            if ((varnames[j]==wnames) & (mcw !=0)){mcwzcov<=0}
            if ((varnames[j]==znames) & (mcz==0))
            {match<-0;ziscov<-(j-nrow(varnames)-ncs)}
            if ((varnames[j]==znames) & (mcz !=0)){mcwzcov<-0}
          }
        }
      }
    }  
    if ((match2 > 0) | (match==1))
    {errcode[errs,1]<-2;errs<-errs+1;criterr<-1}
    if (mcwzcov==1)
    {errcode[errs,1]<-50;errs<-errs+1;criterr<-1}

    #check for factors
    hello<-as.numeric(sapply(dat,is.factor))
    nfacs<-sum(hello)
    if (nfacs > 0){errcode[errs,1]<-53;errs<-errs+1;criterr<-1}

    if (criterr==0)
    {

      #listwise deletion
      ninit<-nrow(dat);rownum<-seq(1:nrow(dat));dat<-cbind(rownum,dat)
      datms<-is.na(dat)
      datms<-matrix(as.numeric(datms),nrow(datms))
      datms<-matrix(rowSums(datms))
      missrow<-0;
      for (i in 1:ninit)
      {if (datms[i,1] > 0){missrow<-cbind(missrow,i)}}
      missrow<-matrix(missrow)
      if (nrow(missrow) > 1)
      {
        missrow<-missrow[2:nrow(missrow)]
        notecode[notes,1]<-29;
        notes<-notes+1;
      }
      missrow<-t(matrix(missrow))
      dat<-na.omit(dat);n<-nrow(dat);nmiss=ninit-n;
      if (n < 5)
      {errcode[errs,1]<-62;errs<-errs+1;criterr<-1}
      if (criterr==0)
        #startit
      {
        rownum<-dat[,1];dat<-dat[,2:ncol(dat)]
        #extract the data back into vectors or matrices
        ytmp=as.data.frame(dat[,1:nys])
        desctmp2<-process.describ3(ytmp,0,quantile)
        desctmp<-matrix(unlist(desctmp2[1]))
        ysd<-desctmp[2,1];ovsd<-ysd;ydich<-0
        if (desctmp[8,1]==2){errcode[errs,1]<-15;errs<-errs+1;criterr<-1;novar<-1}
        if (desctmp[8,1]==1)
        {
          ydich<-1;
          if (total==1){total<-0;notecode[notes,1]<-24;notes<-notes+1}
          if (effsize==1){effsize<-0;notecode[notes,1]<-25;notes<-notes+1}
          omx<-max(ytmp);omn<-min(ytmp)
          ytmp<-matrix(as.numeric(ytmp==omx),ncol=nys)
          dat[,1:nys]<-matrix(as.numeric(dat[,1:nys]==omx),ncol=nys)
          rcd<-c(omn,omx,0,1);dim(rcd)<-c(2,2)
        }
        xtmp<-as.data.frame(dat[,(nys+1):(nys+nxs)])
        desctmp2<-process.describ3(xtmp,0,quantile)
        desctmp<-matrix(unlist(desctmp2[1]))
        xsd<-desctmp[2,1]
        if ((desctmp[8,1]==2) & (novar==0))
        {errcode[errs,1]<-15;errs<-errs+1;criterr<-1;novar<-1}
        xmodvals<-matrix(unlist(desctmp2[2]))
        nxpval<-nrow(xmodvals)
        xprobval<-as.matrix(xmodvals)
        xdich<-desctmp[8,1]
        if ((xdich==1) & (mcx > 0))
        {mcx<-0;errcode[errs,1]<-52;errs<-errs+1;criterr<-1}
        if ((model==74) & (xdich==1)){counterf==1}

        #need to do m
        if (nms > 0)
        {
          mtmp<-as.data.frame(dat[,(nys+nxs+1):(nys+nxs+nms)])
          desctmp2<-process.describ3(mtmp,0,quantile)
          desctmp<-matrix(unlist(desctmp2[1]),ncol=ncol(mtmp))
          ovsd<-matrix(c(desctmp[2,],ysd))
          zzzz<-rowSums(desctmp)
          if ((zzzz[8]>0) & (mdichok != 1))
          {errcode[errs,1]<-43;errs<-errs+1;criterr<-1}   
          mmodvals<-matrix(unlist(desctmp2[2]),ncol=ncol(mtmp))
          mprobval<-mmodvals

        }

        if (nws > 0)
        {
          wtmp<-as.data.frame(dat[,(nys+nxs+nms+1):(nys+nxs+nms+nws)])
          desctmp2<-process.describ3(wtmp,0,quantile)
          desctmp<-matrix(unlist(desctmp2[1]))
          if ((desctmp[8,1]==2) & (novar==0))
          {errcode[errs,1]<-15;errs<-errs+1;criterr<-1;novar<-1}
          wmodvals<-matrix(unlist(desctmp2[2]))
          wdich<-desctmp[8,1]
          if ((wdich==1) & (mcw > 0))
          {mcw<-0;errcode[errs,1]<-52;errs<-errs+1;criterr<-1}
          wmin<-desctmp[3,1]
          wmax<-desctmp[4,1] 
          minwwarn<-matrix(unlist(desctmp2[3]));maxwwarn<-matrix(unlist(desctmp2[4]))
          wnotev<-matrix(unlist(desctmp2[5]))
          wmodval<-t(matrix(wmodval))
          nwcontr<-ncol(wmodval)
          if (wmodval[1,1] != 999)
          {
            wmodvals<-matrix(wmodval[1,1]);wmodcust<-1
            if (nwcontr > 1){wmodvals<-t(wmodval)}
            minwwarn<-0;maxwwarn<-0;wnotev<-0
          }
          wprobval<-as.matrix(wmodvals)
          nwpval<-nrow(wmodvals)
        }

        if (nzs > 0)
        {
          ztmp<-as.data.frame(dat[,(nys+nxs+nms+nws+1):(nys+nxs+nms+nws+nzs)])
          desctmp2<-process.describ3(ztmp,0,quantile)
          desctmp<-matrix(unlist(desctmp2[1]))
          if ((desctmp[8,1]==2) & (novar==0))
          {errcode[errs,1]<-15;errs<-errs+1;criterr<-1;novar<-1}
          zmodvals<-matrix(unlist(desctmp2[2]))
          zdich<-desctmp[8,1]
          if ((zdich==1) & (mcz > 0))
          {mcz<-0;errcode[errs,1]<-52;errs<-errs+1;criterr<-1}
          zmin<-desctmp[3,1]
          zmax<-desctmp[4,1] 
          minzwarn<-matrix(unlist(desctmp2[3]));maxzwarn<-matrix(unlist(desctmp2[4]));
          znotev<-matrix(unlist(desctmp2[5]))
          zmodval<-t(matrix(zmodval))
          nzcontr<-ncol(zmodval)
          if (zmodval[1,1] != 999)
          {
            zmodvals<-matrix(zmodval[1,1]);zmodcust<-1
            if (nzcontr > 1){zmodvals<-t(zmodval)}
            minzwarn<-0;maxzwarn<-0;znotev<-0
          }
          zprobval<-as.matrix(zmodvals)
          nzpval<-nrow(zmodvals)
        }
        if (ncs > 0)
        {
          ctmp<-as.data.frame(dat[,(nys+nxs+nms+nws+nzs+1):(nys+nxs+nms+nws+nzs+ncs)])
          covmean2<-process.describ3(ctmp,1,quantile)
          covmeans<-matrix(unlist(covmean2[1]),ncol=ncs);covmeans<-matrix(covmeans[1,],nrow=1)
        }
        n<-nrow(ytmp)
        ones<-matrix(1,n,1)
        modresid<-matrix(9999,n,1)
        #creat codes for categorical variables
        if ((nws > 0) & (mcw > 0))
        {
          tmp<-cbind(rownum,wtmp[,1])
          dumtmp<-process.makdummy(tmp,mcw,2,wcatcode)
          wmodvals<-matrix(unlist(dumtmp[4]))
          nwpval<-nrow(wmodvals)
          errcode4<-unlist(dumtmp[5])
          errcode5<-unlist(dumtmp[6])
          errcode6<-unlist(dumtmp[7])
          if (errcode4==1){errcode[errs,1]<-4;errs<-errs+1;criterr<-1}
          if (errcode5==1){errcode[errs,1]<-5;errs<-errs+1;criterr<-1}
          if (errcode6==1){errcode[errs,1]<-39;errs<-errs+1;criterr<-1}   
          if (criterr==0)
          {
            nvls<-unlist(dumtmp[3])
            nwvls<-nvls-1
            minwwarn<-0;maxwwarn<-0;wnotev<-0
            wtmp<-matrix(unlist(dumtmp[1]),ncol=(nwvls+1))
            wtmp<-wtmp[,2:ncol(wtmp)];wtmp<-as.data.frame(wtmp)
            wcatlab<-c("W1","W2","W3","W4","W5","W6","W7","W8","W9");wcatlab<-matrix(wcatlab)
            mcwok<-1
            dummatw<-matrix(unlist(dumtmp[2]),ncol=(nwvls+1),nrow=(nwvls+1))
            wprobval<-dummatw[,2:ncol(dummatw)]
            if (modcok==1)
            {
              wcontval<-matrix(-999,2,ncol(wprobval))
              temp<-0
              for (i in (1:2))
              {
                for (j in (1:nrow(dummatw)))
                {
                  if (contvec[i,1]==dummatw[j,1])
                  {wcontval[i,]<-wprobval[j,];temp<-temp+1}
                }
              }
              if (temp < 2)
              {notecode[notes,1]<-20;notes<-notes+1;modcok<-0}
            }
            if (wmodval[1,1] != 999)
            {notecode[notes,1]<-9;notes<-notes+1}
          }
        }
        if ((nzs > 0) & (mcz > 0))
        {
          tmp<-cbind(rownum,ztmp[,1])
          dumtmp<-process.makdummy(tmp,mcz,3,zcatcode)
          zmodvals<-matrix(unlist(dumtmp[4]))
          nzpval<-nrow(zmodvals)
          errcode4<-unlist(dumtmp[5])
          errcode5<-unlist(dumtmp[6])
          errcode6<-unlist(dumtmp[7])
          if (errcode4==1){errcode[errs,1]<-4;errs<-errs+1;criterr<-1}
          if (errcode5==1){errcode[errs,1]<-5;errs<-errs+1;criterr<-1}
          if (errcode6==1){errcode[errs,1]<-40;errs<-errs+1;criterr<-1}
          if (criterr==0)
          {
            nvls<-unlist(dumtmp[3])
            nzvls<-nvls-1
            minzwarn<-0;maxzwarn<-0;znotev<-0
            ztmp<-matrix(unlist(dumtmp[1]),ncol=(nzvls+1))
            ztmp<-ztmp[,2:ncol(ztmp)];ztmp<-as.data.frame(ztmp)
            zcatlab<-c("Z1","Z2","Z3","Z4","Z5","Z6","Z7","Z8","Z9");zcatlab=matrix(zcatlab)
            mczok<-1
            dummatz<-matrix(unlist(dumtmp[2]),ncol=(nzvls+1),nrow=(nzvls+1))
            zprobval<-dummatz[,2:ncol(dummatz)]
            if (modcok==1)
            {
              zcontval<-matrix(-999,2,ncol(zprobval))
              temp<-0
              for (i in (1:2))
              {
                for (j in (1:nrow(dummatz)))
                {
                  if (contvec[i,2]==dummatz[j,1])
                  {zcontval[i,]<-zprobval[j,];temp<-temp+1}
                }
              }
              if (temp < 2)
              {notecode[notes,1]<-20;notes<-notes+1;modcok<-0}
            }
            if (zmodval[1,1] != 999)
            {notecode[notes,1]<-10;notes<-notes+1}
          }
        }
        if ((nxs > 0) & (mcx > 0))
        {
          tmp<-cbind(rownum,xtmp[,1])
          dumtmp<-process.makdummy(tmp,mcx,1,xcatcode)
          errcode4<-unlist(dumtmp[5])
          errcode5<-unlist(dumtmp[6])
          errcode6<-unlist(dumtmp[7])
          if (errcode4==1){errcode[errs,1]<-4;errs<-errs+1;criterr<-1}
          if (errcode5==1){errcode[errs,1]<-5;errs<-errs+1;criterr<-1}
          if (errcode6==1){errcode[errs,1]<-38;errs<-errs+1;criterr<-1}
          if (criterr==0)
          {
            nvls<-unlist(dumtmp[3])
            nxvls<-nvls-1
            xtmp<-matrix(unlist(dumtmp[1]),ncol=(nxvls+1))
            xtmp<-xtmp[,2:ncol(xtmp)];xtmp<-as.data.frame(xtmp)
            xcatlab<-c("X1","X2","X3","X4","X5","X6","X7","X8","X9");xcatlab=matrix(xcatlab)
            xdich<-as.numeric(nvls==2)
            mcxok<-1
            dummatx<-matrix(unlist(dumtmp[2]),ncol=(nxvls+1),nrow=(nxvls+1))
            xmodvals<-as.matrix(dummatx[,1])
            nxpval<-nrow(xmodvals)
          }
        }
        intlab<-matrix(" ",100,1)
        for (i in c(1:100)){intlab[i,1]<-paste("Int_",i,sep='')}
        bcmat<-matrix(0,needed,needed)
        wcmat<-matrix(0,needed,needed)
        zcmat<-matrix(0,needed,needed)
        wzcmat<-matrix(0,needed,needed)
        wsum<-0;zsum<-0;wzsum<-0
      }
    }

  }
  #endit

  # End A loop reads data and does some other things

  # Define model matrices for canned models
  if ((criterr==0) & (model != 999))
  { 
    # X->MW, X->MZ, X->MWZ, M->YW, M->YZ, M->YWZ, X->YW, X->YZ, X->YWZ
    modelmat<-
      matrix(c(1,0,0,0,0,0,0,1,0,0,2,0,0,0,0,0,0,1,1,0,3,0,0,0,0,0,0,1,1,1,4,0,0,0,0,0,0,0,0,0,
               5,0,0,0,0,0,0,1,0,0,6,0,0,0,0,0,0,0,0,0,7,1,0,0,0,0,0,0,0,0,8,1,0,0,0,0,0,1,0,0,
               9,1,1,0,0,0,0,0,0,0,10,1,1,0,0,0,0,1,1,0,11,1,1,1,0,0,0,0,0,0,12,1,1,1,0,0,0,1,1,1,
               13,1,1,1,0,0,0,1,0,0,14,0,0,0,1,0,0,0,0,0,15,0,0,0,1,0,0,1,0,0,16,0,0,0,1,1,0,0,0,0,
               17,0,0,0,1,1,0,1,1,0,18,0,0,0,1,1,1,0,0,0,19,0,0,0,1,1,1,1,1,1,20,0,0,0,1,1,1,1,0,0,
               21,1,0,0,0,1,0,0,0,0,22,1,0,0,0,1,0,1,0,0,23,0,0,0,0,0,0,0,0,0,24,0,0,0,0,0,0,0,0,0,
               25,0,0,0,0,0,0,0,0,0,26,0,0,0,0,0,0,0,0,0,27,0,0,0,0,0,0,0,0,0,28,1,0,0,0,1,0,0,1,0,
               29,1,0,0,0,1,0,1,1,0,30,0,0,0,0,0,0,0,0,0,31,0,0,0,0,0,0,0,0,0,32,0,0,0,0,0,0,0,0,0,
               33,0,0,0,0,0,0,0,0,0,34,0,0,0,0,0,0,0,0,0,35,0,0,0,0,0,0,0,0,0,36,0,0,0,0,0,0,0,0,0,
               37,0,0,0,0,0,0,0,0,0,38,0,0,0,0,0,0,0,0,0,39,0,0,0,0,0,0,0,0,0,40,0,0,0,0,0,0,0,0,0,
               41,0,0,0,0,0,0,0,0,0,42,0,0,0,0,0,0,0,0,0,43,0,0,0,0,0,0,0,0,0,44,0,0,0,0,0,0,0,0,0,
               45,0,0,0,0,0,0,0,0,0,46,0,0,0,0,0,0,0,0,0,47,0,0,0,0,0,0,0,0,0,48,0,0,0,0,0,0,0,0,0,
               49,0,0,0,0,0,0,0,0,0,50,0,0,0,0,0,0,0,0,0,51,0,0,0,0,0,0,0,0,0,52,0,0,0,0,0,0,0,0,0,
               53,0,0,0,0,0,0,0,0,0,54,0,0,0,0,0,0,0,0,0,55,0,0,0,0,0,0,0,0,0,56,0,0,0,0,0,0,0,0,0,
               57,0,0,0,0,0,0,0,0,0,58,1,0,0,1,0,0,0,0,0,59,1,0,0,1,0,0,1,0,0,60,1,1,0,1,0,0,0,0,0,
               61,1,1,0,1,0,0,1,0,0,62,1,1,0,1,0,0,0,1,0,63,1,1,0,1,0,0,1,1,0,64,1,0,0,1,1,0,0,0,0,
               65,1,0,0,1,1,0,1,0,0,66,1,0,0,1,1,0,0,1,0,67,1,0,0,1,1,0,1,1,0,68,1,1,1,1,0,0,0,0,0,
               69,1,1,1,1,0,0,1,1,1,70,1,0,0,1,1,1,0,0,0,71,1,0,0,1,1,1,1,1,1,72,1,1,1,1,1,1,0,0,0,
               73,1,1,1,1,1,1,1,1,1,74,0,0,0,1,0,0,0,0,0,75,1,1,0,1,1,0,0,0,0,76,1,1,0,1,1,0,1,1,0,
               77,0,0,0,0,0,0,0,0,0,78,0,0,0,0,0,0,0,0,0,79,0,0,0,0,0,0,0,0,0,80,0,0,0,0,0,0,0,0,0,
               81,0,0,0,0,0,0,0,0,0,82,0,0,0,0,0,0,0,0,0,83,1,0,0,0,0,0,0,0,0,84,1,0,0,0,0,0,0,0,0,
               85,1,0,0,0,0,0,1,0,0,86,1,0,0,0,0,0,1,0,0,87,0,0,0,1,0,0,0,0,0,88,0,0,0,1,0,0,0,0,0,
               89,0,0,0,1,0,0,1,0,0,90,0,0,0,1,0,0,1,0,0,91,0,0,0,0,0,0,0,0,0,92,1,0,0,1,0,0,1,0,0))
    dim(modelmat)<-c(10,92);modelmat=t(modelmat);
    if (model > 0){tmp<-modelmat[model,2:ncol(modelmat)];tmp=t(matrix(tmp))}
    if (model == 0){tmp<-matrix(0,1,9)}
    if (model < 4) {bcmat[(nxs+1),1]<-1}
    if ((model > 3) & (model != 6))
    {bcmat[(nxs+1):(nxs+nms),1]<-onem
    bcmat[nrow(bcmat),(nxs+1):(nxs+nms)]<-t(onem)
    bcmat[nrow(bcmat),1]<-1}
    if ((model==6) | ((model > 82) & (model < 93)))
    {
      for (j in c(2:nrow(bcmat)))
      {for (i in c(1:(j-1)))
      {bcmat[j,i]<-1}
      }
    } 
    if (model==80)
    {for (i in c(1:nms)){bcmat[(nrow(bcmat)-1),i]<-1}}
    if (model==81)
    {for (j in c(3:nrow(bcmat))){bcmat[j,2]<-1}}
    if (model==82)
    {bcmat[3,2]<-1;bcmat[5,4]<-1}
    if (tmp[1,1]==1)
    {
      wcmat[(nxs+1):(nxs+nms),1]=onem;wprod<-1;xprod<-1
      if ((model==83) | (model==86))
      {onemsx<-onem
      for (i in c(1:(nms-1))){onemsx[(i+1),1]<-0}
      wcmat[(nxs+1):(nxs+nms),1]<-onemsx
      } 
    }
    if (tmp[1,4]==1)
    {
      wcmat[nrow(wcmat),(nxs+1):(nxs+nms)]<-t(onem);wprod<-1;
      if ((model==87) | (model==90))
      {onemsx<-onem
      for (i in c(1:(nms-1))){onemsx[i,1]=0}
      wcmat[nrow(wcmat),(nxs+1):(nxs+nms)]=t(onemsx)
      }
    }
    if (tmp[1,7]==1){wcmat[nrow(wcmat),1]<-1;wprod<-1;xprod<-1}
    if (tmp[1,2]==1){zcmat[(nxs+1):(nxs+nms),1]<-onem;zprod<-1;xprod<-1}
    if (tmp[1,5]==1){zcmat[nrow(zcmat),(nxs+1):(nxs+nms)]=t(onem);zprod<-1}
    if (tmp[1,8]==1){zcmat[nrow(zcmat),1]<-1;zprod<-1;xprod<-1}
    if (tmp[1,3]==1){wzcmat[(nxs+1):(nxs+nms),1]<-onem;xprod<-1;wprod<-1;zprod<-1}
    if (tmp[1,6]==1){wzcmat[nrow(wzcmat),(nxs+1):(nxs+nms)]<-t(onem)}
    if (tmp[1,9]==1){wzcmat[nrow(wzcmat),1]<-1;xprod<-1;wprod<-1;zprod<-1}
    if ((model==91) | (model==92))
    {
      for (j in c(1:(nms-1)))
      {for (i in c(1:j)){wcmat[(nxs+1+j),(nxs+i)]<-1}
      }
    }
  }
  #define and check covariates matrix
  if (ncs > 0)
  {
    ccmat<-matrix(1,(nms+nys),ncs);ccmatoff<-ccmat
    if (covmy==1){ccmat[nrow(ccmat),]<-matrix(0,1,ncs)}
    if (covmy==2){ccmat[1:nms,]<-matrix(0,nms,ncs)}
    if (cmatrix[1,1] != -999)
    {if (ncol(cmatrix) != ((nms+nys)*ncs)){errcode[errs,1]<-29;errs<-errs+1;criterr<-1}
      if (criterr==0)
      {
        tmp<-1;for (i in c(1:(nms+nys)))
        {
          for (j in c(1:ncs))
          {ccmat[i,j]<-(1-as.numeric(cmatrix[1,tmp]==0));tmp<-tmp+1}
        }
        tmpcov<-colSums(ccmat);tmpcov<-as.numeric(tmpcov==0);tmpcov<-sum(tmpcov)
        if (tmpcov !=0)
        {errcode[errs,1]<-30;errs<-errs+1;criterr<-1}
      }
      if (covmy != 0){notecode[notes,1]<-1;notes<-notes+1}
    } 
  }

  #Define matrices for custom models and do some error checking
  if (criterr==0)
  {
    needed<-needed*(needed-1)/2;nopath<-0
    if (bmatrix[1,1] != -999)
    {tmp<-1
    if ((ncol(bmatrix) != needed) | (sum(bmatrix)==0))
    {errcode[errs,1]<-16;errs<-errs+1;criterr<-1} else {
      for (i in c(2:nrow(bcmat)))
      {for (j in c(1:(i-1)))
      {bcmat[i,j]<-(1-as.numeric(bmatrix[1,tmp]==0));tmp<-tmp+1}        
      } 
    }
    #check to make sure X affects something */
    tmpcov<-sum(bcmat[,1])
    if ((tmpcov==0) & (criterr==0))
    {errcode[errs,1]<-22;errs<-errs+1;criterr<-1}
    #check to make sure Y is affected by something */
    tmpcov<-sum(bcmat[nrow(bcmat),])
    if ((tmpcov==0) & (criterr==0))
    {errcode[errs,1]<-23;errs<-errs+1;criterr<-1}
    #check for dangling mediators
    dm<-0
    if (nms > 0)
    {
      for (i in c(1:nms))
        tmpcov<-sum(bcmat[(nxs+i),]);tmpcov2<-sum(bcmat[,(nxs+i)])
        {if (((tmpcov==0) | (tmpcov2==0)) & (dm==0) & (criterr==0))
        {errcode[errs,1]<-26;errs<-errs+1;criterr<-1;dm<-1}       
        }
    }
    }
  }

  #start b
  if (criterr==0)
  {
    if (wmatrix[1,1] != -999)
    {
      tmp<-1
      if (ncol(wmatrix) != needed)
      {errcode[errs,1]<-17;errs<-errs+1;criterr<-1} else { 
        modelvar[1,1]="CUSTOM"
        for (i in (2:nrow(wcmat)))
        {     
          for (j in (1:(i-1)))       
          {wcmat[i,j]<-(1-as.numeric(wmatrix[1,tmp]==0))
          #dont allow to specify moderation of a path that doesnt exist
          if ((wcmat[i,j]==1) & (bcmat[i,j]==0) & (nopath==0))
          {errcode[errs,1]<-20;errs<-errs+1;criterr<-1;nopath<-1}   
          tmp<-tmp+1
          }
        } 
      }
    }

    if (zmatrix[1,1] != -999)
    {
      tmp<-1
      if (ncol(zmatrix) != needed)
      {errcode[errs,1]<-18;errs<-errs+1;criterr<-1} else {
        modelvar[1,1]="CUSTOM"
        if ((sum(wcmat)==0) & (model==999))
        {errcode[errs,1]<-21;errs<-errs+1;criterr<-1}
        for (i in (2:nrow(zcmat)))
        {     
          for (j in (1:(i-1)))       
          {zcmat[i,j]<-(1-as.numeric(zmatrix[1,tmp]==0))
          #dont allow to specify moderation of a path that doesnt exist
          if ((zcmat[i,j]==1) & (bcmat[i,j]==0) & (nopath==0))
          {errcode[errs,1]<-20;errs<-errs+1;criterr<-1;nopath<-1}   
          tmp<-tmp+1
          }
        } 
      }
    }
    tmp<-1
    if (wzmatrix[1,1] != -999)
    {
      if (ncol(wzmatrix) != needed)
      {errcode[errs,1]<-19;errs<-errs+1;criterr<-1}
      modelvar[1,1]="CUSTOM"
    } 
    if (criterr==0)
    {
      for (i in (2:nrow(wzcmat)))
      {
        for (j in (1:(i-1)))
        {
          #set corresponding elements in W and Z for three way interaction
          if (wzmatrix[1,1] != -999){wzcmat[i,j]<-(1-as.numeric(wzmatrix[1,tmp]==0))}
          if (wzcmat[i,j]==1){wcmat[i,j]<-1;zcmat[i,j]<-1}       
          #dont allow to specify moderation of a path that doesnt exist
          if ((wzcmat[i,j]==1) & (bcmat[i,j]==0) & (nopath==0))
          {errcode[errs,1]<-20;errs<-errs+1;criterr<-1;nopath<-1}
          tmp<-tmp+1
        }
      }
    }
  }
  # end B

  if (criterr==0)
  {
    xprod<-(sum(wcmat[,1])+sum(zcmat[,1])+sum(wzcmat[,1]))
    xprod<-as.numeric(xprod > 0)
    wsum<-sum(wcmat)
    wprod<-as.numeric(wsum > 0)
    if (nms > 0)
    {
      for (i in c(1:nms))
      {
        tmp<-(sum(wcmat[,(1+i)])+sum(zcmat[,(1+i)])+sum(wzcmat[,(1+i)]))
        mprod[1,i]<-as.numeric(tmp>0)
      }
    }
    if ((wsum > 0) & (w=="xxxxx")){errcode[errs,1]<-11;errs<-errs+1;criterr<-1}
    if ((wsum==0) & (w != "xxxxx")){errcode[errs,1]<-10;errs<-errs+1;criterr<-1}
    zsum<-sum(zcmat)
    zprod<-as.numeric(zsum > 0)
    if ((zsum > 0) & (z=="xxxxx")){errcode[errs,1]<-13;errs<-errs+1;criterr<-1}
    if ((zsum==0) & (z != "xxxxx")){errcode[errs,1]<-12;errs<-errs+1;criterr<-1}
    if ((zsum > 0) & (wsum==0)){errcode[errs,1]<-35;errs<-errs+1;criterr<-1}
  }
  if ((criterr==0) & (nms > 1))
  {
    serchk<-bcmat[2:(nrow(bcmat)-1),2:ncol(bcmat)]
    if (sum(serchk) > 0)
    {serial<-1
    if (nms > 6){errcode[errs,1]<-36;errs<-errs+1;criterr<-1}
    }
  }
  #mean center if needed
  if ((center > 0) & (criterr==0))
  {
    centvar<-matrix(c(" "))
    if (criterr==0)
    {
      if ((center==1) | ((center==2) & (wdich==0)))
      {
        if ((wprod==1) & (mcwok==0) & (nwpval > 0))
        {
          for (i in c(1:nws))
          {
            wtmp[,i]<-wtmp[,i]-(sum(wtmp[,i])/n)
            centvar<-cbind(centvar,wnames[1,i])
          }
          desctmp2<-process.describ3(wtmp,wmodcust,quantile)
          desctmp<-matrix(unlist(desctmp2[1]))
          wmin<-desctmp[3,1];wmax<-desctmp[4,1]
          if (wmodcust==0)
          {modvals<-matrix(unlist(desctmp2[2]))
          wmodvals<-modvals;wprobval<-wmodvals}
        }
      }    
      if ((center==1) | ((center==2) & (zdich==0)))
      { 
        if ((zprod==1) & (mczok==0) & (nzpval > 0))
        {
          for (i in c(1:nzs))
          {
            ztmp[,i]<-ztmp[,i]-(sum(ztmp[,i])/n)
            centvar<-cbind(centvar,znames[1,i])
          }
          desctmp2<-process.describ3(ztmp,zmodcust,quantile)
          desctmp<-matrix(unlist(desctmp2[1]))
          zmin<-desctmp[3,1];zmax<-desctmp[4,1]
          if (zmodcust==0)
          {modvals<-matrix(unlist(desctmp2[2]))
          zmodvals<-modvals;zprobval<-zmodvals}
        }
      }
      if ((center==1) | ((center==2) & (xdich==0)))
      {
        if ((xprod==1) & (mcxok==0))
        {
          for (i in c(1:nxs))
          {
            xtmp[,i]<-xtmp[,i]-(sum(xtmp[,i])/n)
            centvar<-cbind(centvar,xnames[1,i])
          }
          desctmp2<-process.describ3(xtmp,0,quantile)
          modvals<-matrix(unlist(desctmp2[2]))
          xmodvals<-modvals;xprobval<-as.matrix(xmodvals)
        }
      }
      if (nms > 0)
      {
        for (i in c(1:nms))
        {
          if (mprod[1,i]==1)
          {
            mtmp[,i]<-mtmp[,i]-(sum(mtmp[,i])/n)
            centvar<-cbind(centvar,mnames[i,1])
          }
        }
        desctmp2<-process.describ3(mtmp,0,quantile)
        modvals<-matrix(unlist(desctmp2[2]),ncol=ncol(mtmp))
        mmodvals<-modvals;mprobval<-mmodvals;
      }
    }
    if (ncol(centvar) > 1)
    {notecode[notes,1]<-3;notes<-notes+1}
  }
  # Start D
  #CONSTRUCT THE DATA MATRICES FOR EACH OF THE MODELS
  if (criterr==0)
  {
    # The i loop is the dependent variable in the model matrices
    # The j loop is the predictor variables in the model matrices
    wsum<-sum(wcmat);zsum<-sum(zcmat);wzsum<-sum(wzcmat)
    nump<-matrix(-999,1,(nys+nms));numint<-matrix(0,1,(nys+nms))
    #DV
    datcount<-1;
    xtmpuse<-0;wtmpuse<-0;ztmpuse<-0;xwtmpus<-0;xztmpus<-0;wztmpus<-0;xwztmpu<-0;
    xtmploc<- matrix(-999);wtmploc<- -999;xwtmplo<- -999;ztmploc<- -999
    xztmplo<- -999;wztmplo<- -999;xwztmplo<- -999
    vlabs<-" "
    if (ncs > 0){ctmpuse<-matrix(0,1,ncs)}
    if (nms > 0)
    {
      mtmpuse<-matrix(0,1,nms)
      mwtmpus<-matrix(0,1,nms)
      mztmpus<-matrix(0,1,nms)
      mwztmpu<-matrix(0,1,nms)
      mtmploc<-matrix(0,1,nms)
      mwtmplo<-matrix(-999,nwvls,nms)
      mztmplo<-matrix(-999,nzvls,nms)
      mwztmplo<-matrix(-999,(nwvls*nzvls),nms)
    }
    if (ncs > 0){ctmploc<-matrix(0,1,ncs)}
    fulldat<-matrix(1,n,1)
    datindx<-matrix(-999,1000,(nms+nys))
    wherew<-matrix(-999,2,(nms+nys));wherex<-matrix(-999,2,(nms+nys))
    wherez<-matrix(-999,2,(nms+nys));wherexw<-matrix(-999,2,(nms+nys))
    wherexz<-matrix(-999,2,(nms+nys));wherewz<-matrix(-999,2,(nms+nys))
    wherexwz<-matrix(-999,2,(nms+nys))
    if (nms > 0)
    {
      wherem<-matrix(-999,nms,(nms+nys))
      wheremw<-matrix(-999,(nms*2),(nms+nys))
      wheremz<-matrix(-999,(nms*2),(nms+nys))
      wheremwz<-matrix(-999,(nms*2),(nms+nys))
    }  
    wzhigh<-matrix(0,1000,(((nms+1)*(nms+2))/2))
    whigh<-matrix(0,1000,(((nms+1)*(nms+2))/2))
    zhigh<-matrix(0,1000,(((nms+1)*(nms+2))/2))
    fochigh<-matrix(0,1000,(((nms+1)*(nms+2))/2))
    xcoefloc<-matrix(c(1,2,3,4,5,6,7,8,9))
    intkey<-t(matrix(c(" ", " ", " ", " ", " ", " ", " ")))
    wzhighct<-0;whighct<-0;zhighct<-0;foccnt<-0
    if (nms > 0){mnames<-t(mnames)}
    if (ncs > 0){covnames<-t(covnames)}
    xtmp<-data.matrix(xtmp)

    #START DV LOOP
    for (i in (2:nrow(bcmat)))
    {  
      wdid<-0;zdid<-0;wzdid<-0;cntmp<-1;start<-1;
      if (i < nrow(bcmat))
      {outv<-mtmp[,(i-1)]  
      modlabel<-matrix(c(mnames[1,(i-1)],"constant"))}
      if (i==nrow(bcmat))
      {outv<-ytmp;modlabel<-matrix(c(ynames,"constant"))}
      # The j loop is the mediator
      # START MED

      for (j in (1:(i-1)))
      {
        foccnt<-foccnt+1
        if ((j==1) & (bcmat[i,j]==1))
        {
          outv<-cbind(outv,xtmp)
          modlabel<-matrix(c(modlabel,xcatlab[1:nxvls,1]))
          if (xtmpuse==0)
          {
            fulldat<-matrix(c(fulldat,xtmp),nrow=n)
            xtmpuse<-1
            for (k4 in (datcount:(datcount+(nxvls-1))))
            {xtmploc<-matrix(c(xtmploc,k4))}
            xtmploc<-matrix(xtmploc[2:nrow(xtmploc),1])
            datcount<-datcount+nxvls
          }
          datindx[start:(start+nrow(xtmploc)-1),(i-1)]<-xtmploc
          wherex[1,(i-1)]<-start+1
          wherex[2,(i-1)]<-start+nrow(xtmploc)-1+1
          #do if (model = 74)
          #end if
          onebl<-matrix(1,nrow(xtmploc),1)
          fochigh[(start+1):(start+nrow(xtmploc)),foccnt]<-onebl
          start<-start+nrow(xtmploc)
        }  
        if ((j > 1) & (bcmat[i,j]==1))
        {
          outv<-cbind(outv,mtmp[,(j-1)])
          modlabel<-matrix(c(modlabel,mnames[1,(j-1)]))
          if (mtmpuse[1,(j-1)]==0)
          {
            fulldat<-matrix(c(fulldat,mtmp[,(j-1)]),nrow=n)
            mtmpuse[1,(j-1)]<-1
            mtmploc[1,(j-1)]<-datcount
            datcount<-datcount+1
          }
          datindx[start:(start+nrow(mtmploc)-1),(i-1)]<-mtmploc[1,(j-1)]
          wherem[(j-1),(i-1)]<-start+1
          #onebl<-matrix(1,(nrow(mtmploc[1,(j-1)])),1)
          #onebl<-matrix(1,mtmploc[1,(j-1)],1)
          onebl<-1
          #ttt<-nrow(mtmploc[1,(j-1)])+start-1
          #fochigh[(start+1):(start+nrow(mtmploc[1,(j-1)])),foccnt]<-onebl
          #fochigh[(start+1):(start+mtmploc[1,(j-1)]),foccnt]<-onebl
          fochigh[(start+1):(start+1),foccnt]<-onebl
          #start<-start+mtmploc[1,(j-1)]
          start<-start+1
        }
      }
      #END MED

      #START W
      if (wsum > 0)
        wtmp<-data.matrix(wtmp)
      {
        for (j in c(1:(i-1)))
        {
          whighct<-whighct+1
          if ((j==1) & (wcmat[i,j]==1))
          {
            if (wdid==0)
            {
              outv<-cbind(outv,wtmp)
              if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
              modlabel<-matrix(c(modlabel,wcatlab[1:nwvls,1]))
              wdid<-1
              if (wtmpuse==0)
              {
                fulldat<-matrix(c(fulldat,wtmp),nrow=n)
                if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}         
                wtmpuse<-1
                for (k4 in (datcount:(datcount+(nwvls-1))))
                {wtmploc<-matrix(c(wtmploc,k4))}    
                wtmploc<-matrix(wtmploc[2:nrow(wtmploc),1])
                datcount<-datcount+nwvls
              }
            }     
            datindx[start:(start+nrow(wtmploc)-1),(i-1)]<-wtmploc
            wherew[1,(i-1)]<-start+1
            wherew[2,(i-1)]<-start+nrow(wtmploc)-1+1
            start<-start+nrow(wtmploc)
            for (k1 in c(1:nxvls))
            {
              for (k2 in c(1:nwvls))
              {
                outv<-as.matrix(cbind(outv,(xtmp[,k1]*wtmp[,k2])))
                if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}          
                modlabel<-matrix(c(modlabel,intlab[cntmp,1]))
                intkeyt<-matrix(c(intlab[cntmp,1],":",xcatlab[k1,1],"x",wcatlab[k2,1]," "," "),ncol=7)
                intkey<-rbind(intkey,intkeyt)
                cntmp<-cntmp+1
              }
            }
            if (xwtmpus==0)
            {
              fulldat<-matrix(c(fulldat,outv[,(ncol(outv)-(nxvls*nwvls)+1):ncol(outv)]),nrow=n)
              xwtmpus<-1
              if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
              for (k4 in (datcount:(datcount+((nwvls*nxvls)-1))))
              {xwtmplo<-matrix(c(xwtmplo,k4))}
              xwtmplo<-matrix(xwtmplo[2:nrow(xwtmplo),1])
              datcount<-datcount+(nxvls*nwvls)
            }
            datindx[start:(start+nrow(xwtmplo)-1),(i-1)]<-xwtmplo
            wherexw[1,(i-1)]<-start+1
            wherexw[2,(i-1)]<-start+nrow(xwtmplo)-1+1
            onebl<-matrix(1,nrow(xwtmplo),1)
            whigh[(start+1):(start+nrow(xwtmplo)),whighct]<-onebl
            start<-start+nrow(xwtmplo)
          }
          if ((j > 1) & (wcmat[i,j]==1))
          {
            if ((wdid==0) & (model != 74))
            {
              outv<-cbind(outv,wtmp)
              if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
              modlabel<-matrix(c(modlabel,wcatlab[1:nwvls,1]))
              wdid<-1
              if (wtmpuse==0)
              {
                fulldat<-matrix(c(fulldat,wtmp),nrow=n)
                if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
                wtmpuse<-1
                for (k4 in (datcount:(datcount+(nwvls-1))))
                {wtmploc<-matrix(c(wtmploc,k4))}    
                wtmploc<-matrix(wtmploc[2:nrow(wtmploc),1])
                datcount<-datcount+nwvls
              }
              datindx[start:(start+nrow(wtmploc)-1),(i-1)]<-wtmploc
              wherew[1,(i-1)]<-start+1
              wherew[2,(i-1)]<-start+nrow(wtmploc)-1+1
              start<-start+nrow(wtmploc)
            }         
            for (k2 in c(1:nwvls))
            {
              outv<-as.matrix(cbind(outv,(mtmp[,(j-1)]*wtmp[,k2])))
              if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
              modlabel<-matrix(c(modlabel,intlab[cntmp,1]))
              intkeyt<-matrix(c(intlab[cntmp,1],":",mnames[1,(j-1)],"x",wcatlab[k2,1]," "," "),ncol=7)
              intkey<-rbind(intkey,intkeyt)
              cntmp<-cntmp+1
            }
            if (mwtmpus[1,(j-1)]==0)
            {
              fulldat<-matrix(c(fulldat,outv[,(ncol(outv)-nwvls+1):ncol(outv)]),nrow=n)
              if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
              mwtmpus[1,(j-1)]<-1
              mw22<- -999
              for (k4 in (datcount:(datcount+(nwvls-1))))
              {mw22<-matrix(c(mw22,k4))}
              mwtmplo[,(j-1)]=matrix(mw22[2:nrow(mw22),1])
              datcount<-datcount+nwvls
            }
            datindx[start:(start+nrow(mwtmplo)-1),(i-1)]<-mwtmplo[,(j-1)]
            wheremw[((2*j)-3),(i-1)]<-start+1
            wheremw[((2*j)-2),(i-1)]<-start+nrow(mwtmplo)-1+1
            onebl<-matrix(1,nrow(mwtmplo),1)
            whigh[(start+1):(start+nrow(mwtmplo)),whighct]<-onebl
            start<-start+nrow(mwtmplo)
          }
        }
        }
      #END W
      #START Z
      if (zsum > 0)
        ztmp<-data.matrix(ztmp)
      {
        for (j in c(1:(i-1)))    
        {
          zhighct<-zhighct+1
          if ((j==1) & (zcmat[i,j]==1))
          {
            if (zdid==0)
            {
              outv<-cbind(outv,ztmp)
              if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}         
              modlabel<-matrix(c(modlabel,zcatlab[1:nzvls,1]))
              zdid<-1
              if (ztmpuse==0)
              {
                fulldat<-matrix(c(fulldat,ztmp),nrow=n)
                if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0} 
                ztmpuse<-1
                for (k4 in (datcount:(datcount+(nzvls-1))))
                {ztmploc<-matrix(c(ztmploc,k4))}      
                ztmploc<-matrix(ztmploc[2:nrow(ztmploc),1])
                datcount<-datcount+nzvls
              }
            }   
            datindx[start:(start+nrow(ztmploc)-1),(i-1)]<-ztmploc
            wherez[1,(i-1)]<-start+1
            wherez[2,(i-1)]<-start+nrow(ztmploc)-1+1
            start<-start+nrow(ztmploc)
            for (k1 in c(1:nxvls))
            {for (k2 in c(1:nzvls))
            {        
              outv=as.matrix(cbind(outv,(xtmp[,k1]*ztmp[,k2])))
              if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}
              modlabel<-matrix(c(modlabel,intlab[cntmp,1]))
              intkeyt<-matrix(c(intlab[cntmp,1],":",xcatlab[k1,1],"x",zcatlab[k2,1]," "," "),ncol=7)
              intkey<-rbind(intkey,intkeyt)
              cntmp<-cntmp+1
            }
            }
            if (xztmpus==0)
            {
              fulldat<-matrix(c(fulldat,outv[,(ncol(outv)-(nxvls*nzvls)+1):ncol(outv)]),nrow=n)
              if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}
              xztmpus<-1
              for (k4 in (datcount:(datcount+((nzvls*nxvls)-1)))){xztmplo<-matrix(c(xztmplo,k4))}       
              xztmplo<-matrix(xztmplo[2:nrow(xztmplo),1])
              datcount<-datcount+(nxvls*nzvls)
            }
            datindx[start:(start+nrow(xztmplo)-1),(i-1)]<-xztmplo
            wherexz[1,(i-1)]<-start+1
            wherexz[2,(i-1)]<-start+nrow(xztmplo)-1+1
            onebl<-matrix(1,nrow(xztmplo),1)
            zhigh[(start+1):(start+nrow(xztmplo)),zhighct]<-onebl
            start<-start+nrow(xztmplo)
          }
          if ((j > 1) & (zcmat[i,j]==1))
          {
            if (zdid==0)
            {
              outv<-cbind(outv,ztmp)
              if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}
              modlabel<-matrix(c(modlabel,zcatlab[1:nzvls,1]))
              zdid<-1
              if (ztmpuse==0)
              {
                fulldat<-matrix(c(fulldat,ztmp),nrow=n)
                if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}
                ztmpuse<-1
                for (k4 in (datcount:(datcount+(nzvls-1))))
                {ztmploc<-matrix(c(ztmploc,k4))}
                ztmploc<-matrix(ztmploc[2:nrow(ztmploc),1])
                datcount<-datcount+nzvls
              }
              datindx[start:(start+nrow(ztmploc)-1),(i-1)]<-ztmploc
              wherez[1,(i-1)]<-start+1
              wherez[2,(i-1)]<-start+nrow(ztmploc)-1+1
              start<-start+nrow(ztmploc)
            }    
            for (k2 in (1:nzvls))
            {
              outv=as.matrix(cbind(outv,(mtmp[,(j-1)]*ztmp[,k2])))
              if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}
              modlabel<-matrix(c(modlabel,intlab[cntmp,1]))
              intkeyt<-matrix(c(intlab[cntmp,1],":", mnames[1,(j-1)],"x",zcatlab[k2,1]," "," "),ncol=7)
              intkey<-rbind(intkey,intkeyt)
              cntmp<-cntmp+1
            }
            if (mztmpus[1,(j-1)]==0)
            {
              fulldat<-matrix(c(fulldat,outv[,(ncol(outv)-nzvls+1):ncol(outv)]),nrow=n)
              if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}
              mztmpus[1,(j-1)]<-1
              mz22<- -999
              for (k4 in (datcount:(datcount+(nzvls-1)))){mz22<-matrix(c(mz22,k4))}
              mztmplo[,(j-1)]<-matrix(mz22[2:nrow(mz22),1])
              datcount<-datcount+nzvls
            }
            datindx[start:(start+nrow(mztmplo)-1),(i-1)]<-mztmplo[,(j-1)]
            wheremz[((2*j)-3),(i-1)]<-start+1
            wheremz[((2*j)-2),(i-1)]<-start+nrow(mztmplo)-1+1
            onebl<-matrix(1,nrow(mztmplo),1)
            zhigh[(start+1):(start+nrow(mztmplo)),zhighct]<-onebl
            start<-start+nrow(mztmplo)
          }
        }
        }
      #END Z

      #START WZ
      if (wzsum > 0)
      {  
        for (j in (1:(i-1)))
        {
          wzhighct<-wzhighct+1
          if ((j==1) & (wzcmat[i,j]==1))
          {
            if (wzdid==0)
            {
              for (k1 in (1:nwvls))
              {
                for (k2 in (1:nzvls))
                {
                  outv<-as.matrix(cbind(outv,(wtmp[,k1]*ztmp[,k2]))) 
                  if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}
                  if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
                  modlabel<-matrix(c(modlabel,intlab[cntmp,1]))
                  intkeyt<-matrix(c(intlab[cntmp,1],":",wcatlab[k1,1],"x",zcatlab[k2,1]," "," "),ncol=7)
                  intkey<-rbind(intkey,intkeyt)
                  cntmp<-cntmp+1
                }
              }
              if (wztmpus==0)
              {
                fulldat<-matrix(c(fulldat,outv[,(ncol(outv)-(nwvls*nzvls)+1):ncol(outv)]),nrow=n)
                if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}  
                if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
                wztmpus<-1
                for (k4 in (datcount:(datcount+((nwvls*nzvls)-1))))
                {wztmplo<-matrix(c(wztmplo,k4))}   
                wztmplo<-matrix(wztmplo[2:nrow(wztmplo),1])
                datcount<-datcount+(nzvls*nwvls)
              }
              wzdid<-1         
            }
            datindx[start:(start+nrow(wztmplo)-1),(i-1)]<-wztmplo
            wherewz[1,(i-1)]<-start+1
            wherewz[2,(i-1)]<-start+nrow(wztmplo)-1+1
            start<-start+nrow(wztmplo)
            for (k1 in (1:nxvls))
            {
              for (k2 in (1:nwvls))
              {
                for (k3 in (1:nzvls))
                {
                  outv<-as.matrix(cbind(outv,(xtmp[,k1]*wtmp[,k2]*ztmp[,k3]))) 
                  if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}  
                  if ((ncs > 0) & (wiscov > 0)){ccmatoff((i-1),wiscov)<-0}    
                  modlabel<-matrix(c(modlabel,intlab[cntmp,1]))
                  intkeyt<-matrix(c(intlab[cntmp,1],":",xcatlab[k1,1],"x",wcatlab[k2,1],"x",zcatlab[k3,1]),ncol=7)
                  intkey<-rbind(intkey,intkeyt) 
                  cntmp<-cntmp+1 
                }
              }
            }
            if (xwztmpu==0)
            {
              fulldat<-matrix(c(fulldat,outv[,(ncol(outv)-(nxvls*nwvls*nzvls)+1):ncol(outv)]),nrow=n)
              if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0} 
              if ((ncs > 0) & (wiscov > 0)){ccmatoff((i-1),wiscov)<-0}
              xwztmpu<-1
              for (k4 in (datcount:(datcount+((nzvls*nxvls*nwvls)-1))))
              {xwztmplo<-matrix(c(xwztmplo,k4))}  
              xwztmplo<-matrix(xwztmplo[2:nrow(xwztmplo),1])
              datcount<-datcount+(nxvls*nzvls*nwvls)
            }
            datindx[start:(start+nrow(xwztmplo)-1),(i-1)]<-xwztmplo
            wherexwz[1,(i-1)]<-start+1
            wherexwz[2,(i-1)]<-start+nrow(xwztmplo)-1+1
            onebl<-matrix(1,nrow(xwztmplo),1)
            wzhigh[(start+1):(start+nrow(xwztmplo)),wzhighct]<-onebl
            start<-start+nrow(xwztmplo)
          }
          if ((j > 1) & (wzcmat[i,j]==1))
          {
            if (wzdid==0)
            {
              for (k1 in (1:nwvls))
              {
                for (k2 in (1:nzvls))
                {
                  outv<-as.matrix(cbind(outv,(wtmp[,k1]*ztmp[,k2])))
                  if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}  
                  if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
                  modlabel<-matrix(c(modlabel,intlab[cntmp,1]))
                  intkeyt<-matrix(c(intlab[cntmp,1],":",wcatlab[k1,1],"x",zcatlab[k2,1]," "," "),ncol=7)
                  intkey<-rbind(intkey,intkeyt) 
                  cntmp<-cntmp+1
                }
              }
              if (wztmpus==0)
              {
                fulldat<-matrix(c(fulldat,outv[,(ncol(outv)-(nwvls*nzvls)+1):ncol(outv)]),nrow=n)
                if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}
                if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
                wztmpus<-1
                for (k4 in (datcount:(datcount+((nwvls*nzvls)-1))))
                {wztmplo<-matrix(c(wztmplo,k4))}    
                wztmplo<-matrix(wztmplo[2:nrow(wztmplo),1])
                datcount<-datcount+(nzvls*nwvls)
              }
              wzdid<-1
              datindx[start:(start+nrow(wztmplo)-1),(i-1)]<-wztmplo
              wherewz[1,(i-1)]<-start+1
              wherewz[2,(i-1)]<-start+nrow(wztmplo)-1+1
              start<-start+nrow(wztmplo)
            }
            for (k1 in (1:nwvls))
            {
              for (k2 in (1:nzvls))
              {
                outv<-as.matrix(cbind(outv,(mtmp[,(j-1)]*wtmp[,k1]*ztmp[,k2])))
                if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0}  
                if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0} 
                modlabel<-matrix(c(modlabel,intlab[cntmp,1]))
                intkeyt<-matrix(c(intlab[cntmp,1],":",mnames[1,(j-1)],"x",wcatlab[k1,1],"x",zcatlab[k2,1]),ncol=7)
                intkey<-rbind(intkey,intkeyt) 
                cntmp<-cntmp+1
              }
            }
            if (mwztmpu[1,(j-1)]==0)
            {
              fulldat<-matrix(c(fulldat,outv[,(ncol(outv)-(nwvls*nzvls)+1):ncol(outv)]),nrow=n)
              if ((ncs > 0) & (ziscov > 0)){ccmatoff[(i-1),ziscov]<-0} 
              if ((ncs > 0) & (wiscov > 0)){ccmatoff[(i-1),wiscov]<-0}
              mwztmpu[1,(j-1)]<-1
              mz22<- -999
              for (k4 in (datcount:(datcount+(nwvls*nzvls)-1)))
              {mz22<-matrix(c(mz22,k4))}
              mwztmplo[,(j-1)]<-matrix(mz22[2:nrow(mz22),1])
              datcount<-datcount+(nwvls*nzvls)
            }
            datindx[start:(start+nrow(mwztmplo)-1),(i-1)]<-mwztmplo[,(j-1)]
            wheremwz[((2*j)-3),(i-1)]<-start+1
            wheremwz[((2*j)-2),(i-1)]<-start+nrow(mwztmplo)-1+1
            onebl<-matrix(1,nrow(mwztmplo),1)
            wzhigh[(start+1):(start+nrow(mwztmplo)),wzhighct]<-onebl
            start<-start+nrow(mwztmplo)
          }
        }
      }
      #END WZ

      #START COV
      if (ncs > 0)
      {  
        ccmat<-ccmat*ccmatoff
        for (j in c(1:ncs))
        {
          if (ccmat[(i-1),j]==1)
          {
            if (j==wiscov){ctmp[,j]<-wtmp}
            if (j==ziscov){ctmp[,j]<-ztmp}
            outv<-cbind(outv,ctmp[,j])
            modlabel<-matrix(c(modlabel,covnames[1,j]))
            if (ctmpuse[1,j]==0)
            {
              fulldat<-matrix(c(fulldat,ctmp[,j]),nrow=n)
              ctmpuse[1,j]<-1
              ctmploc[1,j]<-datcount
              datcount<-datcount+1
            }
            datindx[start:(start+nrow(ctmploc)-1),(i-1)]<-ctmploc[1,j]
            start<-start+1
          }
        }
      }
      #END COV

      wdid<-0;zdid<-0;wzdid<-0
      vlabs<-matrix(c(vlabs,modlabel[2:nrow(modlabel),1]))
      numint[1,(i-1)]<-cntmp-1
      nump[1,(i-1)]<-nrow(modlabel)-1

    } 
    #END DV LOOP
    if ((modcok==1) & ((nms > 0) | (zcmat[2,1] != 1) | (mcx != 0)))
    {notecode[notes,1]<-19;notes<-notes + 1;modcok<-0}
    if (((serial==1) | (sum(numint)>0) | (nms==0)) & (mc > 0))
    {notecode[notes,1]<-15;notes<-notes+1;boot<-mc;mc<-0}

    if ((boot != 0) | (mc != 0))
    {
      bootsz<-boot
      if (mc > 0){bootsz<-mc;saveboot<-0}
      cilow<-0;cihigh<-bootsz+1
      while ((cilow <=0) | (cihigh > bootsz))
      {
        cilow<-round(bootsz*(1-(conf/100))/2)
        cihigh<-trunc((bootsz*(conf/100)+(bootsz*(1-(conf/100))/2)))+1
        if ((cilow < 1) | (cihigh > bootsz))
        {bootsz<-trunc((bootsz+1000)/1000)*1000;adjust<-1}
      }
      if (boot > 0){boot<-bootsz}
      if (mc > 0){mc<-bootsz}
      if ((adjust==1) & (boot > 0)){notecode[notes,1]<-8;notes<-notes+1}
      if ((adjust==1) & (mc > 0)){notecode[notes,1]<-16;notes<-notes+1}
    }
    maxboot<-trunc(2*boot)
    if (maxboots > maxboot){maxboot<-trunc(maxboots)}
    #insert seed here  
    if ((seed==-999) & (boot > 0) | (mc > 0))
    {seed<-trunc(runif(1,1,1000000));set.seed(seed);ranseed<-1}
    if (sum(numint) > 0){intkey<-matrix(intkey[2:nrow(intkey),],ncol=7)}
    vlabs<-matrix(vlabs[2:nrow(vlabs),1])
    fulldat<-as.data.frame(fulldat[,2:ncol(fulldat)])
    fulldat<-data.matrix(fulldat)
    fochigh<-fochigh[1:max(nump),]
    whigh<-as.matrix(whigh[1:max(nump),])
    zhigh<-as.matrix(zhigh[1:max(nump),])
    wzhigh<-as.matrix(wzhigh[1:max(nump),])
    coeffs<-fochigh+whigh+zhigh+wzhigh
    bootloc<-matrix(0,max(nump),ncol(nump))

    #Here is am deriving the locations in bootfile needed for indirect effects
    if (nms > 0)
    {
      cntmp<-1
      for (i in (1:ncol(nump)))
      {for (j in (1:nump[1,i]))
      {bootloc[j,i]<-cntmp;cntmp<-cntmp+1}
      }
      fochighb<-matrix(0,nrow(fochigh),ncol(fochigh))
      whighb<-fochighb
      zhighb<-fochighb
      wzhighb<-fochighb
      thetaxmb<-matrix(0,nrow(fochighb),nms)
      thetaxyb<-matrix(0,nrow(fochighb),1)
      pathsfoc<-matrix(0,nxvls,1)
      cntmp<-1
      for (i in (1:(nms+nys)))
      {
        for (j in (1:i))
        {
          fochighb[,cntmp]<-(fochigh[,cntmp]*bootloc[,i])
          whighb[,cntmp]<-(whigh[,cntmp]*bootloc[,i])
          zhighb[,cntmp]<-(zhigh[,cntmp]*bootloc[,i])
          wzhighb[,cntmp]<-(wzhigh[,cntmp]*bootloc[,i])
          coeffsb<-(fochighb+whighb+zhighb+wzhighb)
          if ((i < (nms+nys)) & (j==1)){thetaxmb[,i]<-coeffsb[,cntmp]}
          if ((i==(nms+nys)) & (j==1)){thetaxyb[,1]<-coeffsb[,cntmp]}
          cntmp<-cntmp+1
        }
      }
      thetamyb<-as.matrix(coeffsb[,(ncol(coeffsb)-nms+1):ncol(coeffsb)])
      if (serial==1)
      {thetammb<-matrix(0,nrow(coeffsb),((nms*(nms-1))/2))}
      cntmp<-1
      if ((nms > 1) & (serial==1))
      {
        for (i in (1:(nms-1)))
        {
          start<-((i+2)*(i+1))/2
          for (j in (2:(nms-i+1)))
          {
            thetammb[,cntmp]<-coeffsb[,start]
            start<-start+j+i-1
            cntmp<-cntmp+1
          }
        } 
      }
    }

    if ((total==1) & (sum(numint)==0))
    {
      dototal<-1
      if ((sum(bcmat[,1]) != (nms+nys)) | (sum(bcmat[nrow(bcmat),]) != (nms+nys)))
      {dototal<-0;alttotal<-1;notecode[notes,1]<-12;notes<-notes+1}
      if (ncs > 0)
      {
        if ((sum(ccmat)) < (nrow(ccmat)*ncol(ccmat)))
        {dototal<-0;alttotal<-1;notecode[notes,1]<-11;notes<-notes+1}
      }
    }
  }
  # End D
  if ((criterr==0) & (ncs > 0))
  {
    tmperr<-sum(as.numeric(colSums(ccmat)==0))
    if (sum(tmperr)!=0)
    {errcode[errs,1]<-51;errs<-errs+1;criterr<-1}
  }
  if (outscreen==1)
  {
    cat("\n")
    cat("********************* PROCESS for R Version 4.1.1 ********************* \n \n")
    cat("           Written by Andrew F. Hayes, Ph.D.  www.afhayes.com              \n")
    cat("   Documentation available in Hayes (2022). www.guilford.com/p/hayes3   \n \n")
  }
  if (criterr==0)
  {
    if ((stand==1) & (ydich==1)){stand<-0}
    anymod2<-sum(wcmat+zcmat+wzcmat)
    if ((anymod2 > 0) & (stand==1))
    {notecode[notes,1]<-27;notes<-notes+1;stand<-0}
    rownames(modelvar)<-modelvlb
    colnames(modelvar)<-" "
    funny<-1
    if (outscreen==1)
    {
      cat("*********************************************************************** \n")
      print(noquote(modelvar))
      cat("\n")
      if (ncs > 0)
      {
        cat("Covariates: \n")
        covname2<-c("      ", covnames)
        covname2<-t(noquote(covname2))  
        write.table(covname2,quote=FALSE,row.names=FALSE,col.names=FALSE)
        cat("\n")
      }
      cat("Sample size: ")
      write.table(n,quote=FALSE,row.names=FALSE,col.names=FALSE)
      cat("\n")
      if (seed != -999)    
      {if (ranseed==0)
      {cat("Custom seed: ")}
        if (ranseed==1)
        {cat("Random seed: ")}
        write.table(seed,quote=FALSE,row.names=FALSE,col.names=FALSE)
        cat("\n")
      }
      maxresm<-9;
      resultm<-matrix(99999,1,maxresm)  
      if (describe==1)
      {
        means=apply(dat,2,mean);sdvec=apply(dat,2,sd);corall=cor(dat);
        means<-rbind(t(matrix(means)),t(matrix(sdvec)))
        means2<-noquote(matrix(sprintf(decimals,means),nrow=nrow(means)))
        corall2<-noquote(matrix(sprintf(decimals,corall),nrow=nrow(corall)))
        if (ncol(means) > 9)
        {resultm<-matrix(99999,1,ncol(means));maxresm<-ncol(means)}
        resultm2<-matrix(99999,2,maxresm)
        resultm2[1:2,1:ncol(means)]<-means
        resultm<-rbind(resultm,resultm2)
        resultm2<-matrix(99999,ncol(corall),maxresm)
        resultm2[1:ncol(corall),1:ncol(corall)]<-corall
        resultm<-rbind(resultm,resultm2)
        colnames(means2)<-varnames
        colnames(corall2)<-varnames
        rownames(corall2)<-varnames
        rownames(means2)<-matrix(c("Mean","SD"))
        if (outscreen==1)
        {
          cat("Variable means and standard deviations:\n")
          print(means2,right=T)
          cat("\n")
          cat("Variable intercorrelations (Pearson's r):\n")
          print(corall2,right=T)
        }     
      }

      #resultm2<-matrix(99999,nrow(corall),maxresm)
      #resultm2[1:nrow(corall),1:ncol(corall)]<-corall
      #resultm<-rbind(resultm,resultm2)

      if (mcxok==1)
      {cat("Coding of categorical X variable for analysis: \n")
        dummatx2<-noquote(matrix(sprintf(decimals,dummatx),nrow=nrow(dummatx)))
        colnames(dummatx2)<-c(xnames,xcatlab[1:nxvls,1])
        rownames(dummatx2)<-t(matrix(replicate((nxvls+1)," ")))
        print(dummatx2,right=T)}
      if (mcwok==1)
      {cat("Coding of categorical W variable for analysis: \n")
        dummatw2<-noquote(matrix(sprintf(decimals,dummatw),nrow=nrow(dummatw)))
        colnames(dummatw2)<-c(wnames,t(wcatlab[1:nwvls,1]))
        rownames(dummatw2)<-t(matrix(replicate((nwvls+1)," ")))
        print(dummatw2,right=T)}
      if (mczok==1)
      {cat("Coding of categorical Z variable for analysis: \n")
        dummatz2<-noquote(matrix(sprintf(decimals,dummatz),nrow=nrow(dummatz)))
        colnames(dummatz2)<-c(znames,t(zcatlab[1:nzvls,1]))
        rownames(dummatz2)<-t(matrix(replicate((nzvls+1)," ")))
        print(dummatz2,right=T)}
    }
  } 
  #START cycle through the models  
  if (criterr==0)
  {
    outnames<-ynames
    outvars<-ytmp
    if (nms > 0)
    {
      outnames<-matrix(c(mnames,ynames))
      outvars<-cbind(mtmp,ytmp)
      indcov<-matrix(0,((nms*2)+(nms*(nxvls-1))),((nms*2)+(nms*(nxvls-1))))
      mcsopath<-matrix(0,((nms*2)+(nms*(nxvls-1))),1)
    }
    labstart<-1
    intstart<-1
    start<-1
    coeffmat<-matrix(0,1,6)
    conseq<-"        "
    dfmat<-0;coeffcol<-0;pathscnt<-1;pathscn2<-1

    # START G LOOP
    for (i in c(1:(nms+nys)))
    {
      if (outscreen==1)
      {cat("\n*********************************************************************** \n")}
      highf<-matrix(0,1,5);highf2<-highf
      if ((i==(nms+nys)) & (ydich==1)){highf=matrix(0,1,3);highf2<-highf}
      flabel<-" "   
      y<-outvars[,i]
      xindx<-datindx[1:(nump[1,i]-1),i]
      x<-fulldat[,xindx]
      x<-cbind(ones,x)
      xsq<-t(x)%*%x
      exsq<-eigen(xsq)
      exsq<-matrix(unlist(exsq[1]))
      zeroeig<-sum(as.numeric(exsq <= 0.000000000002))
      if (outscreen==1)
      {
        cat("Outcome Variable: ")
        write.table(outnames[i,1],quote=FALSE,row.names=FALSE,col.names=FALSE)
        cat("\n")
        if ((ydich==1) & (i==(nms+nys)))
        {
          cat("Coding of binary Y for logistic regression analysis:\n")
          rcd<-noquote(matrix(sprintf(decimals,rcd),nrow=nrow(rcd)))
          colnames(rcd)<-c(outnames[i,1], "Analysis")
          rownames(rcd)<-t(matrix(replicate(nrow(rcd)," ")))
          print(rcd,right=T)
          cat("\n")
        }
      }
      if (zeroeig > 0)
      {
        cat("\nSINGULAR OR NEAR SINGULAR DATA MATRIX.\n")
        criterr<-1;errcode[errs,1]<-31;errs<-errs+1;
      }    
      means<-colSums(x)/n
      vlabsm<-matrix(vlabs[labstart:(labstart+(nump[1,i]-1)),1])
      #START E
      if (criterr==0)
      {
        if ((ydich==0) | (i < (nms+nys)))
        {
          modoutz<-process.modelest(y,x,1,1,xp2,hc)
          modres<-matrix(unlist(modoutz[1]),nrow=nump[1,i])
          direff<-matrix(modres,nrow=ncol(x))
          #modresl<-matrix(unlist(modoutz[2]))  
          modsum<-matrix(unlist(modoutz[3]))
          tval<-matrix(unlist(modoutz[7]))
          resid<-matrix(unlist(modoutz[8]))
          modresid<-cbind(modresid,resid)
          brsq2<-modsum[2,1]
          dfres<-modsum[6,1]
          #for recording results
          resultm2<-matrix(99999,1,maxresm)
          resultm2[1,1:nrow(modsum)]<-t(modsum)
          resultm<-rbind(resultm,resultm2)
          resultm2<-matrix(99999,nrow(modres),maxresm)
          resultm2[1:nrow(modres),1:ncol(modres)]<-modres
          resultm<-rbind(resultm,resultm2)
          #end recording results
          modrest9<-modres
          modres<-noquote(matrix(sprintf(decimals,modres),nrow=nump[1,i]))
          modsum<-noquote(matrix(sprintf(decimals,modsum),nrow=1))
          b<-matrix(unlist(modoutz[5]))
          varb<-matrix(unlist(modoutz[6]),nrow=nump[1,i])
          dfmatt<-matrix(modsum[1,6],nrow(modres),1)
          modsuml<-matrix(c("R","R-sq","MSE",hcflab,"df1","df2", "p"))
          modresl<-t(matrix(c("coeff",hclab,"t","p","LLCI","ULCI")))
        }
        if ((ydich==1) & (i==(nms+nys)))
        {
          modoutz<-process.modelest(y,x,2,1,xp2,5,iterate,converge)
          modres<-matrix(unlist(modoutz[1]),nrow=nump[1,i])
          direff<-matrix(modres,nrow=ncol(x))
          modsum<-matrix(unlist(modoutz[3]))
          basemod<-modsum[1,1]
          basemodx<-basemod
          tval<-matrix(unlist(modoutz[7]))
          resultm2<-matrix(99999,1,maxresm)
          resultm2[1,1:nrow(modsum)]<-t(modsum)
          resultm<-rbind(resultm,resultm2)
          resultm2<-matrix(99999,nrow(modres),maxresm)
          resultm2[1:nrow(modres),1:ncol(modres)]<-modres
          resultm<-rbind(resultm,resultm2)
          modrest9<-modres
          modres<-noquote(matrix(sprintf(decimals,modres),nrow=nump[1,i]))
          modsum<-noquote(matrix(sprintf(decimals,modsum),nrow=1))
          b<-matrix(unlist(modoutz[5]))
          varb<-matrix(unlist(modoutz[6]),nrow=nump[1,i])
          modsuml<-matrix(c("-2LL","ModelLL", "df", "p", "McFadden", "CoxSnell", "Nagelkrk"))
          modresl<-t(matrix(c("coeff","se","Z","p","LLCI","ULCI")))
          #modretrn<-list(modres,modresl,modsum,modsuml,b,varb,xp2)
          dfmatt<-matrix(-999,nrow(modres),1)
        } 
        obscoeff<-cbind(obscoeff,t(b))
        if (outscreen==1)
        {
          cat("Model Summary: \n")
          colnames(modsum)<-modsuml;rownames(modsum)<-" "
          print(modsum,right=T)
          cat("\n")
          cat("Model: \n")
          rownames(modres)<-vlabsm;colnames(modres)<-modresl
          print(modres,right=T)   
          if ((ydich==1) & (i==(nms+nys)))
          {
            cat("\nThese results are expressed in a log-odds metric.\n")
            notecode[notes,1]<-26;notes<-notes+1
          }
        }
        coeffmat<-rbind(coeffmat,modres)
        conseqt<-matrix(outnames[i,1],nrow(modres),1)
        conseq<-matrix(c(conseq,conseqt))
        dfmat<-rbind(dfmat,dfmatt)
        labstart<-labstart+nump[1,i]

        if (stand==1)
        {
          predsd<-matrix(0,nrow(modres),1)
          stdmod<-as.numeric(modres[,1])/ovsd[i,1]
          for (jd in c(1:ncol(x)))
          {
            predsd[jd,1]<-sd(x[,jd])
          }
          if ((wherex[1,i] != -999) & ((nxvls > 1) | (xdich==1)))    
          {
            sdmsone<-matrix(1,nxvls,1)
            predsd[wherex[1,i]:wherex[2,i],1]<-sdmsone    
            pstog<-1    
          }
          predsd[1,1]<-1
          stdmod<-stdmod*predsd
          stdmod<-matrix(stdmod[2:nrow(stdmod),1])
          sdvlabs<-vlabsm[2:nrow(vlabsm),1]
          if (outscreen==1){cat("\nStandardized coefficients:\n")}
          resultm2<-matrix(99999,nrow(stdmod),maxresm)
          resultm2[1:nrow(stdmod),1:ncol(stdmod)]<-stdmod
          resultm<-rbind(resultm,resultm2)
          stdmod<-noquote(matrix(sprintf(decimals,stdmod),nrow=(nump[1,i]-1)))
          colnames(stdmod)<-"coeff"
          rownames(stdmod)<-sdvlabs
          if (outscreen==1)
          {print(stdmod,right=T)}
        }
        if ((nms > 0) & (serial==0) & (sum(numint)==0) & ((normal==1) | (mc > 0)))
        {
          if (i < (nms+nys))
          {
            indcov[(((i-1)*nxvls)+1):(i*nxvls),(((i-1)*nxvls)+1):(i*nxvls)]<-varb[2:(1+nxvls),2:(1+nxvls)]
            mcsopath[(((i-1)*nxvls)+1):(i*nxvls),1]<-matrix(modrest9[2:(1+nxvls),1])
          }
          if (i==(nms+nys))
          {
            atm<-ncol(wherem)
            indcov[((nms*nxvls)+1):nrow(mcsopath),((nms*nxvls)+1):nrow(mcsopath)]<-varb[wherem[1,atm]:(wherem[1,atm]+nms-1),wherem[1,atm]:(wherem[1,atm]+nms-1)]
            mcsopath[((nms*nxvls)+1):nrow(mcsopath),1]<-matrix(modrest9[wherem[1,atm]:(wherem[1,atm]+nms-1),1])
            sobelok<-1          
          }
        }
        obsdirfx<-matrix(0,1,nxvls);dirzes<-matrix(0,1,nxvls)    
        if ((i ==(nms+nys)) & (bcmat[nrow(bcmat),1]==1))
        {
          direff<-matrix(direff[2:(1+nxvls),],nrow=nxvls)
          obsdirfx=t(direff[,1])
          direfflb<-modresl
          direffl2<-vlabsm[2:(1+nxvls),]
          lmat<-matrix(0,nrow(b),1)
          lmat2<-matrix(1,nxvls,1)
          lmat[2:(1+nxvls),1]<-lmat2
          if (ydich != 1)
          {
            diromni<-process.ftest3(lmat,b,varb,1,brsq2,0,y,x)
          }
          if (ydich==1){diromni<-process.llrtest3(lmat,y,x,b,basemod,iterate,converge)}
        }

        if (numint[1,i] > 0)
        {
          if (outscreen==1)
          {
            cat("\nProduct terms key:\n")
            intkeym<-matrix(intkey[intstart:(intstart+numint[1,i]-1),],ncol=7)
            write.table(intkeym,quote=FALSE,row.names=FALSE,col.names=FALSE,sep = "  ")
          }
        }
        if (covcoeff==1)
        {
          if (outscreen==1){cat("\nCovariance matrix of regression parameter estimates:\n")}
          resultm2<-matrix(99999,nrow(varb),maxresm)
          if (ncol(varb) <= maxresm)
          {
            resultm2[1:nrow(varb),1:ncol(varb)]<-varb
            resultm<-rbind(resultm,resultm2)
          }
          if (ncol(varb) > maxresm)
          {
            resultmt<-matrix(99999,nrow(resultm),ncol(varb))
            resultmt[1:nrow(resultm),1:ncol(resultm)]<-resultm
            resultm<-resultmt
            resultm2<-matrix(99999,nrow(varb),ncol(resultm))
            resultm2[1:nrow(varb),1:ncol(varb)]<-varb
            resultm<-rbind(resultm,resultm2)
            maxresm<-ncol(resultm)
          }
          varbpr<-noquote(matrix(sprintf(decimals,varb),nrow=nump[1,i]))
          rownames(varbpr)<-vlabsm;colnames(varbpr)<-vlabsm
          if (outscreen==1)
          {print(varbpr,right=T)}
        }
        #start X by M interaction */
        if ((model != 74) & (xmtest==1) & (nms > 0))
        {
          r2tmp<-brsq2;btmp<-b;varbtmp<-varb
          dfrestmp<-dfres
          tvaltmp<-tval
          xmtst<-matrix(0,nms,4)
          xmtstlbc=matrix(c(hcflab,"df1","df2","p"))
          if ((i==(nms+nys)) & (ydich==1))
          {xmtst<-matrix(0,nms,3);xmtstlbc<-c("Chi-sq","df","p")
          }
          xmtstlb<-" ";xmtmat<-x;numxint<-0
          if (i > 1)
          {
            for (xmint in (2:i))
            {
              x<-xmtmat
              if ((bcmat[(i+1),xmint]==1) & (wzcmat[(i+1),xmint] != 1)) 
              {
                if (bcmat[(i+1),1]==0)
                {
                  x<-cbind(xmtmat,xtmp)
                  if ((ydich==1) & (i==(nms+nys)))
                  {
                    LL2<-process.modelest(y,x,2,0,xp2,5,iterate,converge)
                    basemodx<-LL2
                  }
                }
                for (xmtlp1 in (1:nxvls))
                {
                  mtmpmns<-sum(mtmp[,(xmint-1)])/nrow(mtmp)
                  x<-as.matrix(cbind(x,(xtmp[,xmtlp1]*((mtmp[,(xmint-1)])-mtmpmns))))
                }
                if ((i < (nms+nys)) | (ydich==0))
                {
                  modoutz<-process.modelest(y,x,1,1,xp2,hc)
                  b<-matrix(unlist(modoutz[5]))
                  varb<-matrix(unlist(modoutz[6]),nrow=nrow(b))
                  lmat<-matrix(0,nrow(b),nxvls)
                  lmattmp<-diag(nxvls)
                  lmat[(nrow(lmat)-nxvls+1):nrow(lmat),]<-lmattmp
                  fresult2<-process.ftest3(lmat,b,varb,0,0,1,y,x)
                  numxint<-numxint+1
                  xmtst[numxint,]<-fresult2
                }
                #right here
                if ((i==(nms+nys)) & (ydich==1))
                {
                  LL2<-process.modelest(y,x,2,0,xp2,5,iterate,converge)
                  chidfxm<-(basemodx-LL2)
                  numxint<-numxint+1
                  xmtst[numxint,1]<-chidfxm
                  xmtst[numxint,3]<-(1-pchisq(chidfxm,df=nxvls))
                }
                xmtstlbt<-matrix(c(highlbx[(xmint-1),1]))
                xmtstlb<-rbind(xmtstlb,xmtstlbt)     
              }      
            }
          }
          x<-xmtmat
          if (numxint > 0)
          {           
            xmtstlb<-matrix(xmtstlb[(2:(numxint+1)),])
            xmtst<-matrix(xmtst[1:numxint,],nrow=numxint)
            if (nms==1){xmtstlb<-" "}
            resultm2<-matrix(99999,nrow(xmtst),maxresm)
            resultm2[1:nrow(xmtst),1:ncol(xmtst)]<-xmtst
            resultm<-rbind(resultm,resultm2)
            if ((i < (nms+nys)) | (ydich==0))
            {
              xmtst<-noquote(matrix(sprintf(decimals,xmtst),ncol=4))
              colnames(xmtst)<-xmtstlbc
              rownames(xmtst)<-xmtstlb
              if (outscreen==1)
              {cat("\nTest(s) of X by M interaction:\n")
                print(xmtst,right=T)}
            } 
            if ((i==(nms+nys)) & (ydich==1))
            {
              xmtst<-noquote(matrix(sprintf(decimals,xmtst),ncol=3))
              colnames(xmtst)<-xmtstlbc
              rownames(xmtst)<-xmtstlb
              if (outscreen==1)
              {cat("\nLikelihood ratio test(s) of X by M interaction:\n")
                print(xmtst,right=T)}
            } 
          }   
          b<-btmp;varb<-varbtmp;dfres<-dfrestmp;tval<-tvaltmp;brsq2<-r2tmp
        }
        #end X by M interaction

        #here is where we do F tests
        #START F
        if (criterr==0)
        {
          jj<-0
          for (j in (start:((start+i)-1)))
          {
            dbint<-0
            lmat<-whigh[1:nump[1,i],j]
            lmat2<-wzhigh[1:nump[1,i],j]
            if ((sum(lmat) > 0) & (sum(lmat2)==0)) 
            {         
              if ((i < (nms+nys)) | (ydich != 1))
              {
                fresult2<-process.ftest3(lmat,b,varb,chr=1,brsq2,0,y,x)
                lmatdb<-lmat
                dbint=dbint<-1
              }
              if ((ydich==1) & (i==(nms+nys)))
              {                      
                fresult2<-process.llrtest3(lmat,y,x,b,basemod,iterate,converge)
                lmatdb<-lmat
                dbint<-dbint+1
              }
              highf<-rbind(highf,fresult2)
              highf2<-rbind(highf2,fresult2)
              if (j==start){flabel<-matrix(c(flabel,"X*W"))}      
              if (j > start)
              {
                if (nms > 1){flabel<-matrix(c(flabel,highlbw[jj,1]))}
                if (nms==1){flabel<-matrix(c(flabel,"M*W"))}
              }
            }     
            lmat<-zhigh[1:nump[1,i],j]
            lmat2<-wzhigh[1:nump[1,i],j]
            if ((sum(lmat) > 0) & (sum(lmat2)==0))
            {
              if ((i < (nms+nys)) | (ydich != 1))
              {
                fresult2<-process.ftest3(lmat,b,varb,chr=1,brsq2,0,y,x)
                dbint<-dbint+1             
              }
              if ((ydich==1) & (i==(nms+nys)))
              {
                fresult2<-process.llrtest3(lmat,y,x,b,basemod,iterate,converge)
                dbint<-dbint+1    
              }
              highf<-rbind(highf,fresult2)   
              highf2<-rbind(highf2,fresult2)                
              if (j==start){flabel<-matrix(c(flabel,"X*Z"))}
              if (j > start)
              {
                if (nms > 1){flabel<-matrix(c(flabel,highlbz[jj,1]))}
                if (nms==1){flabel<-matrix(c(flabel,"M*Z"))}
              }
            }
            if (dbint==2)
            {
              lmatdb<-(lmatdb+lmat)
              if ((ydich==1) & (i==(nms+nys)))
              {fresult2<-process.llrtest3(lmatdb,y,x,b,basemod,iterate,converge)}       
              if ((ydich != 1) | (i < (nms+nys)))
              {
                fresult2<-process.ftest3(lmatdb,b,varb,chr=1,brsq2,0,y,x)
              }
              dbint<-0 
              highf<-rbind(highf,fresult2)
              if ((jj==0) & (nms > 0)){flabel<-matrix(c(flabel,"BOTH(X)"))}  
              if ((jj==0) & (nms==0)){flabel<-matrix(c(flabel,"BOTH"))}  
              if ((jj > 0) & (nms==1)){flabel<-matrix(c(flabel,"BOTH(M)"))}  
              if ((nms > 1) & (jj > 0)){flabel<-matrix(c(flabel,highlbbt[jj,1]))}         
            } 
            lmat2<-wzhigh[1:nump[1,i],j]
            if (sum(lmat2) > 0)
            {
              if ((i < (nms+nys)) | (ydich != 1))
              {
                fresult2<-process.ftest3(lmat2,b,varb,chr=1,brsq2,0,y,x)
              }
              if ((ydich==1) & (i==(nms+nys)))
              {fresult2<-process.llrtest3(lmat2,y,x,b,basemod,iterate,converge)}
              highf<-rbind(highf,fresult2)
              highf2<-rbind(highf2,fresult2)
              if (j==start){flabel<-matrix(c(flabel,"X*W*Z"))}
              if (j > start)
              {
                if (nms > 1){flabel<-matrix(c(flabel,highlbwz[jj,1]))}
                if (nms==1){flabel<-matrix(c(flabel,"M*W*Z"))}
              }
            }
            jj<-jj+1
          }
          start<-start+i
          highfsz<-ncol(highf)
        }
        #END F

        if (nrow(highf) > 1)
        {
          highf<-matrix(highf[2:nrow(highf),],ncol=highfsz)
          highf2<-matrix(highf2[2:nrow(highf2),],ncol=highfsz)
          flabel<-as.matrix(flabel[2:nrow(flabel),])
          resultm2<-matrix(99999,nrow(highf),maxresm)
          resultm2[1:nrow(highf),1:ncol(highf)]<-highf
          resultm<-rbind(resultm,resultm2)
          highf<-noquote(matrix(sprintf(decimals,highf),ncol=highfsz))
          if (outscreen==1)
          {
            if ((i < (nms+nys)) | (ydich==0))
            {      
              cat("\nTest(s) of highest order unconditional interaction(s):\n")
              colnames(highf)<-matrix(c("R2-chng",hcflab,"df1","df2","p"))
              rownames(highf)<-flabel
              print(highf,right=T)      
            }
            if ((ydich==1) & (i==(nms+nys)))
            {
              cat("\nLikelihood ratio test of highest order\n")
              cat("unconditional interaction(s):\n")
              colnames(highf)<-matrix(c("Chi-sq", "df","p"))
              rownames(highf)<-flabel
              print(highf,right=T)
            }
          }      
          intpb<-as.matrix(highf2[,ncol(highf2)])
        }
        intstart<-intstart+numint[1,i]
      }
      #END E

      #Start PROBEandPLOT
      if (criterr==0)
      {
        threeway<-0;didprint<-0;didsome<-0;sigintct<-0
        #Start R
        for (jmed in (1:(nms+1)))
        {
          hasw<-0;hasz<-0;jnok<-0;nm1vls<-0;nm2vls<-0;panelgrp<-0
          focpred4<-matrix(" ")
          intprint<-0;modcat<-0    
          #Start A
          if (jmed <= i)
          {
            if ((jmed==1) & ((i+1)==nrow(bcmat)))
            {pathscnt<-pathscnt+1} else {
              paths<-cbind(paths,bcmat[(i+1),jmed])
              pathsw<-cbind(pathsw,wcmat[(i+1),jmed])
              pathsz<-cbind(pathsz,zcmat[(i+1),jmed])
              pathswz<-cbind(pathswz,wzcmat[(i+1),jmed])
              temp<-matrix(fochigh[,pathscnt]*bootloc[,i])
              pathsfoc<-as.matrix(cbind(pathsfoc,pathsfoc[,1]))
              if (jmed==1){pathtype<-cbind(pathtype,1)}
              if ((i+1)==nrow(bcmat)){pathtype<-cbind(pathtype,3)}
              if ((jmed > 1) & ((i+1) < nrow(bcmat))){pathtype<-cbind(pathtype,2)}
              if ((jmed==1) & (nxvls > 1) & (bcmat[(i+1),jmed]==1))
              {pathsfoc[,(pathscn2+1)]<-temp[2:(nxvls+1),1]}
              if ((jmed > 1) | ((jmed==1) & (nxvls==1)))
              {
                temp<-matrix(apply(temp,2,max))  #column max
                pathsfoc[1,(pathscn2+1)]<-temp
              }
              pathscnt<-pathscnt+1;pathscn2<-pathscn2+1
              if (i <= nms){pathsdv<-matrix(c(pathsdv,mnames[1,i]))}
              if (i > nms){pathsdv<-matrix(c(pathsdv,ynames))}
            }
            coeffcol<-coeffcol+1
            probettt<-matrix(coeffs[1:nrow(b),coeffcol])
            if ((jmed==1) & (bcmat[(i+1),jmed]==1))
            {
              omni<-matrix(0,nrow(probettt),nxvls)
              omnitmp<-diag(nxvls)
              omni[2:(1+nxvls),]<-omnitmp
            }
            if (sum(probettt) > 0)
            {
              probvarb<-matrix(999,sum(probettt),sum(probettt))
              probcoef<-matrix(999,sum(probettt),1)
              coefflp2<-1
              for (coefflp in (1:nrow(probettt)))
              {
                if (probettt[coefflp,1]==1)
                {
                  probcoef[coefflp2,1]<-b[coefflp,1]
                  coefflp2<-coefflp2+1
                }
              }
              coefflp<-0;coefflp2<-0
              for (iclp in (1:nrow(probettt)))
              {
                if (probettt[iclp,1]==1)
                {
                  coefflp<-coefflp+1
                  coefflp2<-coefflp
                  probvarb[coefflp,coefflp]<-varb[iclp,iclp]
                  if (iclp < nrow(probettt))
                  {
                    for (jclp in ((iclp+1):nrow(probettt)))
                    {       
                      if (probettt[jclp,1]==1)
                      {
                        coefflp2<-coefflp2+1
                        probvarb[coefflp,coefflp2]<-varb[iclp, jclp]
                        probvarb[coefflp2,coefflp]<-varb[iclp, jclp]
                      }
                    }
                  }
                }
              } 
            } 
          }
          #End A
          xprobval<-as.matrix(xmodvals)
          if ((nxvls > 1) | (mcx > 0)){xprobval<-as.matrix(dummatx[,2:ncol(dummatx)])}
          #Start B
          if ((wcmat[(i+1),jmed]==1) & (zcmat[(i+1),jmed]==0))
          {
            numplps<-1;modvals<-wmodvals;probeval<-wmodvals;wheremv1<-wherexw
            nm1vls<-nwvls;lpstsp<-t(matrix(c(1,1)));modcat<-0;jnmod<-wtmp;jnmodlab<-wnames
            jnok<-1;jnmin<-wmin;jnmax<-wmax;wherejn1<-2
            if (jmed==1)
            {
              wherejn3<-wherexw[1,i]
              if (nxvls > 1){jnok<-0}
            }
            if (jmed > 1)
            {
              wherejn1<-wherem[(jmed-1),i]
              wherejn3<-wheremw[((2*jmed)-3),i]
            }
            if (nwvls > 1)
            {
              probeval<-wprobval
              lpstsp[1,2]<-ncol(probeval)
              modcat<-1;jnok<-0
            }
            if (wdich==1){modcat<-1;jnok<-0}
            problabs<-wnames
            focpred3<-t(matrix(c(wnames,"(W)"))) 
            hasw<-1
            modgrph<-wnames;intprint<-1;sigintct<-sigintct+1;
            printpbe<-intpb[sigintct,1]
          }
          #end B
          #start C
          if ((wcmat[(i+1),jmed]==0) & (zcmat[(i+1),jmed]==1))
          {
            numplps<-1;modvals<-zmodvals;probeval<-zmodvals;wheremv1<-wherexz
            nm1vls<-nzvls;lpstsp<-t(matrix(c(1,1)));jnok<-1;jnmod<-ztmp;jnmin<-zmin;jnmax<-zmax
            jnmodlab<-znames;wherejn1<-2
            if (jmed==1)
            { 
              wherejn3<-wherexz[1,i]
              if (nxvls > 1){jnok<-0}
            }
            if (jmed > 1){wherejn1<-wherem[(jmed-1),i];wherejn3<-wheremz[((2*jmed)-3),i]}
            if (nzvls > 1)
            {
              probeval<-zprobval;lpstsp[1,2]<-ncol(probeval);modcat<-1;jnok<-0
            }
            if (zdich == 1){modcat<-1;jnok<-0}
            problabs<-znames
            focpred3<-t(matrix(c(znames,"(Z)"))) 
            modgrph<-znames;hasz<-1;intprint<-1;sigintct<-sigintct+1;
            printpbe<-intpb[sigintct,1]
          }
          # end C
          # start D
          if ((wzcmat[(i+1),jmed]==1) |  ((wcmat[(i+1),jmed]==1) & (zcmat[(i+1),jmed]==1)))
          {
            numplps<-2;probecnt<-1;intprint<-1
            if (wzcmat[(i+1),jmed]==1)
            {
              sigintct<-sigintct+1;printpbe<-intpb[sigintct,1]
            } else {
              sigintct<-sigintct+2;printpbe<-min(intpb[(sigintct-1):sigintct,1])
            } 
            panelgrp<-1;hasw<-1;hasz<-1
            modgrph<-wnames
            lpstsp<-matrix(1,2,2)
            wheremv1<-wherexw;nm1vls<-nwvls;wheremv2<-wherexz;nm2vls<-nzvls;jnok<-0
            if (wzcmat[(i+1),jmed]==1){jnok<-1}
            if (jmed > 1){mprobval<-mmodvals}
            if (jmed==1){if (nxvls > 1){jnok<-0}}
            if (nwvls > 1){lpstsp[1,2]<-ncol(wprobval);modcat<-1;jnok<-0}
            if (zdich==1){modcat<-1;jnok<-0}
            lpstsp[2,1]<-lpstsp[1,2]+1;lpstsp[2,2]<-lpstsp[1,2]+1
            if (nzvls > 1)
            {
              lpstsp[2,1]<-lpstsp[1,2]+1
              lpstsp[2,2]<-lpstsp[1,2]+ncol(zprobval)
              jnok<-0
            }
            if (zdich==1){jnok<-0}
            omni3<-matrix(0,nrow(b),(nxvls*nwvls),0)
            if (jmed > 1){omni3<-matrix(0,nrow(b),nwvls)}
            focpred3<-t(matrix(c(wnames,"(W)")))
            focpred4<-t(matrix(c("      Moderator:", znames, "(Z)"))) 
            modvals<-matrix(0,(nrow(wmodvals)*nrow(zmodvals)),2)
            probeval<-matrix(0,(nrow(wmodvals)*nrow(zmodvals)),(ncol(wprobval)+ncol(zprobval)))
            for (probei in (1:nrow(wmodvals)))
            {
              for (probej in (1:nrow(zmodvals)))
              {
                modvals[probecnt,1]<-wmodvals[probei,1]
                probeval[probecnt,1:nwvls]<-wprobval[probei,]
                modvals[probecnt,2]<-zmodvals[probej,1]
                probeval[probecnt,(nwvls+1):(nwvls+nzvls)]<-zprobval[probej,]
                probecnt<-probecnt+1
              }
            }
            if (wzcmat[(i+1),jmed]==1)
            {
              numplps<-numplps+1
              probprod<-matrix(0,1,(ncol(wprobval)*ncol(zprobval)))
              lpstsp2<-t(matrix(c(1,1)))
              lpstsp=rbind(lpstsp,lpstsp2)
              lpstsp[3,1]<-lpstsp[2,2]+1
              lpstsp[3,2]<-lpstsp[2,2]+ncol(probprod)
              jnmod<-ztmp;jnmin<-zmin;jnmax<-zmax;jnmodlab<-znames
              if (jmed == 1){wherejn1<-wherexw[1,i];wherejn3<-wherexwz[1,i]}
              if (jmed > 1)
              {
                wherejn1<-wheremw[((2*jmed)-3),i]
                wherejn3<-wheremwz[((2*jmed)-3),i]
              }
              for (probei in (1:nrow(wmodvals)))
              {
                for (probej in (1:nrow(zmodvals)))
                {
                  probtemp<-1
                  for (probek in (1:ncol(wprobval)))
                  {
                    #probtemp<-cbind(probtemp,(wprobval[probei,probek]*zprobval[probej,]))
                    probtemp<-cbind(probtemp,matrix((wprobval[probei,probek]*zprobval[probej,]),ncol=ncol(zprobval)))
                  }
                  probprod<-rbind(probprod,probtemp[1,2:ncol(probtemp)])
                }
              }
              probprod<-probprod[2:nrow(probprod),]
              probeval<-cbind(probeval,probprod)
            }
            problabs<-cbind(wnames,znames)
          }
          # end D
          # start E
          if (intprint==1)
          {
            focpred<-"Focal predictor:"
            if (jmed==1)
            {focpred<-cbind(focpred,xnames,"(X)");focplotv<-as.matrix(xmodvals)}
            if (jmed >1)
            {
              if (nms > 1){focpred<-cbind(focpred,mnames[1,(jmed-1)], medlb2[(jmed-1),1])}
              if (nms==1){focpred<-cbind(focpred,mnames[1,(jmed-1)], "(M)")}
              focplotv<-as.matrix(mmodvals[,(jmed-1)],nrow=nrow(mmodvals))
            }
            focpred2<-cbind("      Moderator:",focpred3)
            focpred<-rbind(focpred,focpred2)
            if (ncol(focpred4) > 1){focpred<-rbind(focpred,focpred4);focpred4<-matrix(c(" "))}
            if (outscreen==1)
            {
              if (((plot==1) | (plot==2)) | (printpbe <= intprobe))      
              {cat("----------\n");write.table(focpred,quote=FALSE,row.names=FALSE,col.names=FALSE)}
            }
            foctmp<-matrix(1,nrow(modvals),1)
            probexpl<-1
            probeva2<-cbind(foctmp,probeval)
            if ((jmed==1) & (nxs > 0) & (mcx > 0)){probexpl<-nxvls}
            foctmp<-matrix(1,nrow(modvals),1)
            modvals3<-matrix(0,1,(6+ncol(problabs)))
            probrown<-matrix(0,nrow(probeval),1)
            jtmp<-1
            for (probei in (1:nrow(probeval))){probrown[probei,1]<-jtmp;jtmp<-jtmp+nxvls}
            probrow<-999;modvarl<-problabs

            # start F */ 
            if ((plot==1) | (plot==2) | (nxvls > 1))
            {
              plotvals<-matrix(999,(nrow(modvals)*nrow(focplotv)),(ncol(modvals)+1))
              for (ploti in (1:nrow(modvals)))
              {
                for (plotj in (1:nrow(focplotv)))
                {
                  plotvals[(((ploti-1)*nrow(focplotv))+plotj),2:ncol(plotvals)]<-modvals[ploti,]
                  plotvals[(((ploti-1)*nrow(focplotv))+plotj),1]<-focplotv[plotj,1]
                }
              }
              focpredn<-3
              if (jmed==1)
              {
                if (nxvls > 1){focpredn<-(nxvls+1)}
                if ((nxvls==1) & (xdich==1)){focpredn<-2}
              }
              meanmat<-diag(c(means))
              onesmat<-matrix(1,nrow(meanmat),(nrow(probeval)*focpredn))
              probeplt<-t(diag(means)%*%onesmat)
              # start G
              if (jmed==1)
              {
                if ((wcmat[(i+1),1]==1) | (zcmat[(i+1),1]==1))
                {
                  plotcnt<-1;iloops<-nwpval*nzpval;plotmx<-nxpval*nzpval
                  if ((wcmat[(i+1),1]==1) & (zcmat[(i+1),1]==0))
                  {iloops<-nwpval;plotmx<-nxpval}
                  if ((wcmat[(i+1),1]==0) & (zcmat[(i+1),1]==1))
                  {iloops<-nzpval;plotmx<-nxpval}     
                  xestvals<-matrix(-999,(nxpval*iloops),ncol(xprobval))
                  if (wcmat[(i+1),1]==1){westvals<-matrix(-999,nrow(xestvals),ncol(wprobval))} 
                  if (zcmat[(i+1),1]==1){zestvals<-matrix(-999,nrow(xestvals),ncol(zprobval))}
                  for (ploti in (1:iloops))
                  {
                    for (plotj in (1:nxpval))
                    {xestvals[plotcnt,]<-xprobval[plotj,];plotcnt<-plotcnt+1}
                  }
                  plotcnt<-1;plotcnt1<-1;plotcnt2<-1;plotcntz<-1
                  for (ploti in (1:(iloops*nxpval)))
                  {
                    if (wcmat[(i+1),1]==1){westvals[ploti,]=wprobval[plotcnt1,]}        
                    if ((wcmat[(i+1),1]==0) & (zcmat[(i+1),1]==1))
                    {zestvals[ploti,]<-zprobval[plotcnt1,]}   
                    if ((wcmat[(i+1),1]==1) & (zcmat[(i+1),1]==1))
                    {zestvals[ploti,]<-zprobval[plotcnt2,];plotcntz=plotcntz+1}
                    plotcnt<-plotcnt+1
                    if (plotcnt > plotmx){plotcnt<-1;plotcnt1<-plotcnt1+1}
                    if (plotcntz > nxpval)
                    {
                      plotcnt2<-plotcnt2+1;plotcntz<-1
                      if (plotcnt2 > nzpval){plotcnt2<-1} 
                    }
                  }
                  probeplt[,2:(1+(ncol(xestvals)))]<-xestvals 
                  if (wcmat[(i+1),1]==1)
                  {probeplt[,wherew[1,i]:wherew[2,i]]<-westvals}            
                  if (zcmat[(i+1),1]==1)
                  {probeplt[,wherez[1,i]:wherez[2,i]]<-zestvals}  
                }
              }
              # end G
              # start H
              if (jmed > 1)
              {
                if ((wcmat[(i+1),jmed]==1) | (zcmat[(i+1),jmed]==1))
                {
                  plotcnt<-1;iloops<-(nwpval*nzpval);plotmx<-(3*nzpval)
                  if ((wcmat[(i+1),jmed]==1) & (zcmat[(i+1),jmed]==0))
                  {iloops<-nwpval;plotmx<-3}
                  if ((wcmat[(i+1),jmed]==0) & (zcmat[(i+1),jmed]==1))
                  {iloops<-nzpval;plotmx<-3}
                  mestvals<-matrix(-999,(3*iloops),1)
                  if (wcmat[(i+1),jmed]==1)
                  {westvals<-matrix(-999,nrow(mestvals),ncol(wprobval))}
                  if (zcmat[(i+1),jmed]==1)
                  {zestvals<-matrix(-999,nrow(mestvals),ncol(zprobval))}
                  for (ploti in (1:iloops))
                  {
                    for (plotj in (1:3))
                    {
                      mestvals[plotcnt,]<-mprobval[plotj,(jmed-1)]
                      plotcnt<-plotcnt+1
                    }
                  }
                  plotcnt<-1;plotcnt1<-1;plotcnt2<-1;plotcntz<-1
                  for (ploti in (1:(iloops*3)))
                  {
                    if (wcmat[(i+1),jmed]==1)
                    {westvals[ploti,]<-wprobval[plotcnt1,]}     
                    if ((wcmat[(i+1),jmed]==0) & (zcmat[(i+1),jmed]==1))
                    {zestvals[ploti,]<-zprobval[plotcnt1,]}   
                    if ((wcmat[(i+1),jmed]==1) & (zcmat[(i+1),jmed]==1))
                    {zestvals[ploti,]<-zprobval[plotcnt2,];plotcntz<-plotcntz+1}
                    plotcnt<-plotcnt+1
                    if (plotcnt > plotmx){plotcnt<-1;plotcnt1<-plotcnt1+1}        
                    if (plotcntz > 3)
                    {
                      plotcnt2<-plotcnt2+1;plotcntz<-1
                      if (plotcnt2 > nzpval){plotcnt2<-1} 
                    }
                  }
                  probeplt[,wherem[(jmed-1),i]]<-mestvals
                  if (wcmat[(i+1),jmed]==1)
                  {
                    if (model != 74)
                    {probeplt[,(wherew[1,i]):(wherew[2,i])]<-westvals}
                    if (model==74)
                    {probeplt[,(wherex[1,i]):(wherex[2,i])]<-westvals}
                  }    
                  if (zcmat[(i+1),jmed]==1)
                  {probeplt[,(wherez[1,i]):(wherez[2,i])]<-zestvals}  
                }
              } 
              # end H
              # Here I am doing the multiplications to produce data for the plot
              prodloop <-1
              if (jmed==1){prodloop<-ncol(xestvals)}
              if ((wcmat[(i+1),jmed])==1)
              {
                plotcnt<-0
                for (ploti in (1:prodloop))
                {
                  for (plotj in (1:ncol(westvals)))
                  {
                    if (jmed==1)
                    {probeplt[,(wherexw[1,i]+plotcnt)]<-(xestvals[,ploti]*westvals[,plotj])}
                    if (jmed > 1)
                    {probeplt[,(wheremw[((jmed*2)-3) ,i]+plotcnt)]<-(mestvals[,ploti]*westvals[,plotj])}
                    plotcnt<-plotcnt+1
                  }
                } 
              }
              if ((zcmat[(i+1),jmed])==1)
              {
                plotcnt<-0
                for (ploti in (1:prodloop))
                {
                  for (plotj in (1:ncol(zestvals)))
                  {   
                    if (jmed==1)
                    {probeplt[,(wherexz[1,i]+plotcnt)]<-(xestvals[,ploti]*zestvals[,plotj])}    
                    if (jmed > 1)
                    {probeplt[,(wheremz[((jmed*2)-3),i]+plotcnt)]<-(mestvals[,ploti]*zestvals[,plotj])}
                    plotcnt<-plotcnt+1
                  }
                } 
              }
              if ((wzcmat[(i+1),jmed])==1)
              {
                plotcnt<-0;threeway<-1
                for (ploti in (1:ncol(westvals)))
                {
                  for (plotj in (1:ncol(zestvals)))
                  {
                    probeplt[,(wherewz[1,i]+plotcnt)]<-(westvals[,ploti]*zestvals[,plotj])
                    plotcnt<-plotcnt+1
                  }
                }  
                plotcnt<-0
                for (plotk in (1:prodloop))
                {
                  for (ploti in (1:ncol(westvals)))
                  {
                    for (plotj in (1:ncol(zestvals)))
                    {
                      if (jmed==1)
                      {probeplt[,(wherexwz[1,i]+plotcnt)]<-(xestvals[,plotk]*westvals[,ploti]*zestvals[,plotj])}          
                      if (jmed > 1)
                      {probeplt[,(wheremwz[((jmed*2)-3),i]+plotcnt)]<-(mestvals[,plotk]*westvals[,ploti]*zestvals[,plotj])}          
                      plotcnt<-plotcnt+1
                    }
                  } 
                }
              }
              # here is where we add holding constant products as needed
              # start I
              for (newplp in (1:i))
              {  
                if (newplp != jmed)
                {  
                  if (wcmat[(i+1),newplp]==1)
                  {
                    prodloop<-1
                    if (newplp==1){prodloop<-nxvls}
                    plotcnt<-0
                    for (ploti in (1:prodloop))
                    {
                      for (plotj in (1:nwvls))
                      {
                        if (newplp==1)
                        {probeplt[,(wherexw[1,i]+plotcnt)]<-probeplt[,(1+ploti)]*probeplt[,(wherew[1,i]+plotj-1)]}
                        if (newplp > 1)
                        {
                          if (model != 74)
                          {probeplt[,(wheremw[((newplp*2)-3) ,i]+plotcnt)]<-probeplt[,wherem[(newplp-1),i]]*probeplt[,(wherew[1,i]+plotj-1)]}
                          if (model==74)
                          {probeplt[,(wheremw[((newplp*2)-3) ,i]+plotcnt)]<-probeplt[,wherem[(newplp-1),i]]*probeplt[,(wherex[1,i]+plotj-1)]}           
                        }
                        plotcnt<-plotcnt+1
                      }
                    } 
                  }  
                  if (zcmat[(i+1),newplp]==1)
                  {
                    prodloop<-1
                    if (newplp==1){prodloop<-nxvls}
                    plotcnt<-0
                    for (ploti in (1:prodloop))
                    {
                      for (plotj in (1:nzvls))
                      {
                        if (newplp==1)
                        {probeplt[,(wherexz[1,i]+plotcnt)]<-probeplt[,(1+ploti)]*probeplt[,(wherez[1,i]+plotj-1)]}
                        if (newplp > 1)
                        {probeplt[,(wheremz[((newplp*2)-3),i]+plotcnt)]<-probeplt[,wherem[(newplp-1),i]]*probeplt[,(wherez[1,i]+plotj-1)]}
                        plotcnt<-plotcnt+1
                      }
                    } 
                  }
                  if (wzcmat[(i+1),newplp]==1)
                  {
                    plotcnt<-0
                    if (threeway==0)
                    {
                      for (ploti in (1:nwvls))
                      {
                        for (plotj in (1:nzvls))
                        {
                          probeplt[,(wherewz[1,i]+plotcnt)]<-probeplt[,(wherew[1,i]+ploti-1)]*probeplt[,(wherez[1,i]+plotj-1)]
                          plotcnt<-plotcnt+1
                        }
                      }
                    } 
                    prodloop<-1
                    if (newplp==1){prodloop<-nxvls}
                    plotcnt<-0
                    for (plotk in (1:prodloop))
                    {
                      for (ploti in (1:nwvls))
                      {
                        for (plotj in (1:nzvls))
                        {
                          if (newplp==1)
                          {probeplt[,(wherexwz[1,i]+plotcnt)]<-probeplt[,(1+plotk)]*probeplt[,(wherew[1,i]+ploti-1)]*probeplt[,(wherez[1,i]+plotj-1)]}      
                          if (newplp > 1)
                          {probeplt[,(wheremwz[((newplp*2)-3),i]+plotcnt)]<-probeplt[,wherem[(newplp-1),i]]*probeplt[,(wherew[1,i]+ploti-1)]*probeplt[,(wherez[1,i]+plotj-1)]}      
                          plotcnt<-plotcnt+1
                        }
                      } 
                    }
                  }      
                }
              }
              #END I
              predvals<-probeplt%*%b

              if ((i==(nms+nys)) & (ydich==1))
              {
                predvalt<-as.numeric(predvals < 709.7)
                prevalt7<-(1-predvalt)*(709.7)
                predvals<-(predvals*predvalt)+prevalt7
                expyhat<-exp(predvals)/(1+exp(predvals))
              } 
              sepred<-matrix(999,nrow(plotvals),3)
              for (sei in (1:nrow(plotvals)))
              {
                ask<-as.matrix(probeplt[sei,],ncol=ncol(probeplt))       
                sepred[sei,1]<-sqrt(t(ask)%*%varb%*%ask)
                if ((i < (nms+nys)) | (ydich==0))
                {
                  sepred[sei,2]<-predvals[sei,1]-tval*sepred[sei,1]
                  sepred[sei,3]<-predvals[sei,1]+tval*sepred[sei,1]
                }
                if ((i==(nms+nys)) & (ydich==1))
                {
                  sepred[sei,2]<-predvals[sei,1]-xp2*sepred[sei,1]
                  sepred[sei,3]<-predvals[sei,1]+xp2*sepred[sei,1]
                }
              }
              prevloc<-ncol(plotvals)+1
              probeplt<-cbind(plotvals,predvals)
              if (plot==2){probeplt<-cbind(probeplt,sepred)}
              if ((i==(nms+nys)) & (ydich==1))
              {probeplt<-cbind(probeplt,expyhat)}
              didsome<-0
            }
            # END F

            # here is the loop that is printing the conditional effects
            # this does conditional two way interactions
            # START J
            if ((wzcmat[(i+1),jmed]==1) & (printpbe <= intprobe))
            {
              if (jmed==1)
              {omnilp2<-nxvls*nwvls;omnitmp<-diag(omnilp2);omni3[wherexw[1,i]:wherexw[2,i],]<-omnitmp}  
              if (jmed>1)
              {omnilp2<-nwvls;omnitmp<-diag(omnilp2);omni3[wheremw[((jmed*2)-3),i]:wheremw[((jmed*2)-2),i],]<-omnitmp}
              omnif<-matrix(0,1,4)
              if ((i==(nms+nys)) & (ydich==1))
              {omnif<-matrix(0,1,3)}
              condeff3<-0
              for (omnilp1 in (1:nrow(zprobval)))
              {
                for (omnilp in (1:omnilp2))
                {
                  if (jmed==1)
                  {
                    omni3[(wherexwz[1,i]+((omnilp-1)*nzvls)):(wherexwz[1,i]+((omnilp-1)*nzvls)+(nzvls-1)),omnilp]<-t(zprobval[omnilp1,])
                  }
                  if (jmed > 1)
                  {
                    omni3[(wheremwz[((jmed*2)-3),i]+((omnilp-1)*nzvls)):(wheremwz[((jmed*2)-3),i]+((omnilp-1)*nzvls)+(nzvls-1)),omnilp]<-t(zprobval[omnilp1,])
                  }
                }
                condeff<-t(omni3)%*%b
                condeff3<-rbind(condeff3,condeff)
                fresult2<-process.ftest3(omni3,b,varb,chr=0,brsq2,1,y,x)

                if ((i==(nms+nys)) & (ydich==1))
                {
                  fratio<-fresult2[1,1]*ncol(omni3)
                  pfr<-(1-pchisq(fratio,df=ncol(omni3)))
                  fresult2<-cbind(fratio,ncol(omni3),pfr)
                }
                omnif<-rbind(omnif,fresult2)
              }
              omnif<-matrix(omnif[2:nrow(omnif),],ncol=ncol(omnif))
              clabtmp<-znames
              condeff3<-matrix(condeff3[2:nrow(condeff3),],ncol=ncol(condeff3))
              if ((nxvls*nwvls)==1)
              {
                omnif<-cbind(condeff3,omnif)
                clabtmp<-cbind(clabtmp,"effect")
              } 
              omnif<-cbind(zmodvals,omnif)
              if ((i < nms+nys) | (ydich==0))
              {clabtmp<-c(clabtmp,hcflab,"df1","df2","p")}      
              if ((i==(nms+nys)) & (ydich==1))
              {clabtmp<-c(clabtmp,"Chi-sq","df","p")}
              resultm2<-matrix(99999,nrow(omnif),maxresm)
              resultm2[1:nrow(omnif),1:ncol(omnif)]<-omnif
              resultm<-rbind(resultm,resultm2)
              omnif2<-noquote(matrix(sprintf(decimals,omnif),nrow=nrow(omnif)))
              colnames(omnif2)<-clabtmp
              rownames(omnif2)<-t(matrix(replicate(nrow(omnif)," ")))
              if (outscreen==1)
              {
                if (jmed==1)
                {cat("\nTest of conditional X*W interaction at value(s) of Z:\n")
                  print(omnif2,right=T)}
                if (jmed > 1)
                {cat("\nTest of conditional M*W interaction at value(s) of Z:\n")
                  print(omnif2,right=T)}
              }
            }
            # END J

            # start O 
            for (probei in (1:probexpl))
            {
              if (probexpl > 1)
              {
                foctmp<-matrix(0,nrow(modvals),probexpl)
                foctmp[,probei]<-foctmp[,probei]+1
                probtemp<-matrix(0,nrow(modvals),1)
                for (probem in (1:numplps))
                {for (probek in (1:nxvls))
                {for (probej in (lpstsp[probem,1]:lpstsp[probem,2]))
                {probtemp<-cbind(probtemp,foctmp[,probek]*probeval[,probej])}  
                }
                }
                probeva2<-probtemp[,2:ncol(probtemp)]
                probeva2<-cbind(foctmp,probeva2)
              }    
              probres<-probeva2%*%probcoef
              probrese<-as.matrix(sqrt(diag(probeva2%*%probvarb%*%t(probeva2)))) 
              tratio<-probres/probrese
              if ((ydich==1) & (i==(nms+nys)))
              {
                p<-2*(1-pnorm(abs(tratio)))
              } else {
                p<-2*pt(-abs(tratio),df=dfres)
              }
              modvals2<-cbind(modvals,probres,probrese,tratio, p)
              if ((i < nms+nys) | (ydich==0))
              {
                modvals2<-cbind(modvals2,(probres-as.numeric(tval)*probrese),(probres+as.numeric(tval)*probrese))   
                problabs<-cbind(problabs,"effect",hclab,"t", "p", "LLCI", "ULCI")
              }
              if ((ydich==1) & (i==(nms+nys)))
              {
                modvals2<-cbind(modvals2,(probres-xp2*probrese),(probres+xp2*probrese))
                problabs<-cbind(problabs,"effect","se","Z", "p", "LLCI", "ULCI")
              }
              # start L
              if ((probexpl > 1) & (printpbe <= intprobe))
              {
                if (hasz==1){printz<-1}
                if (hasw==1){printw<-1}
                probrlab<-matrix(xcatlab[probei,1],nrow(modvals),1)
                modvals3<-rbind(modvals3,modvals2)
                probrow<-rbind(probrow,probrown)
                probrown<-probrown+1
                if (probei==probexpl)
                {
                  xproblab<-xcatlab[1:nxvls,1]
                  probrow<-matrix(probrow[2:nrow(probrow),1])
                  modvals3<-modvals3[2:nrow(modvals3),]
                  #modvals3<-modvals3[order(modvals3[,1]),]
                  modvals3<-modvals3[order(probrow[,1]),]
                  start2<-1
                  problabs<-problabs[1,(1+(ncol(modvarl))):ncol(problabs)]
                  pstart<-1
                  # start K
                  for (probek in (1:nrow(probeval)))
                  {
                    endstart<-start2+(nxvls-1)
                    temp<-modvals3[start2:endstart,(1+ncol(modvarl)):ncol(modvals3)]
                    temp2<-t(modvals3[start2:start2,1:ncol(modvarl)])
                    trnames<-t(modvarl)
                    if (outscreen==1)
                    {
                      if (probek > 1){cat("----------\n")}
                      if (probek==1)
                      {
                        cat("\nConditional effects of the focal predictor at values of the moderator(s):\n")
                        if ((jmed==1) & (i ==(nms+nys)) & (nms > 0))
                        {
                          if (nxvls==1)
                          {cat("\n(These are also the conditional direct effects of X on Y.)\n")}
                          if (nxvls != 1)
                          {cat("\n(These are also the relative conditional direct effects of X on Y.)\n")}
                        }
                        cat("\n")
                      }
                      cat("Moderator value(s):\n")
                    }
                    resultm2<-matrix(99999,nrow(temp2),maxresm)
                    resultm2[1:nrow(temp2),1:ncol(temp2)]<-temp2
                    resultm<-rbind(resultm,resultm2)
                    temp2<-noquote(matrix(sprintf(decimals,temp2),nrow=nrow(temp2)))
                    temp3<-matrix(temp2[1,1:ncol(temp2)])
                    rownames(temp3)<-trnames
                    colnames(temp3)<-" "
                    if (outscreen==1)
                    {print(temp3,right=TRUE,quote=FALSE)
                      cat("\n")}
                    #for (mdpntr in (1:ncol(temp2)))
                    #{
                    #write.table(temp2[1,mdpntr],quote=FALSE,row.names=trnames[mdpntr,1],col.names=FALSE)
                    #}          
                    resultm2<-matrix(99999,nrow(temp),maxresm)
                    resultm2[1:nrow(temp),1:ncol(temp)]<-temp
                    resultm<-rbind(resultm,resultm2)
                    temp<-noquote(matrix(sprintf(decimals,temp),nrow=nrow(temp)))
                    colnames(temp)<-problabs[1:ncol(temp)]
                    rownames(temp)<-xproblab
                    if (outscreen==1)
                    {print(temp,right=T)}
                    start2<-start2+nxvls
                    didsome<-1
                    if (jmed==1)
                    {
                      mod1val<-probeval[probek,1:nm1vls]
                      for (omnilp in (1:nxvls))
                      {
                        omni[(wheremv1[1,i]+((omnilp-1)*nm1vls)):(wheremv1[1,i]+((omnilp-1)*nm1vls)+(nm1vls-1)),omnilp]<-t(mod1val)
                        if (nm1vls < ncol(probeval))
                        {
                          mod2val<-probeval[probek,(nm1vls+1):(nm1vls+nm2vls)]
                          omni[(wheremv2[1,i]+((omnilp-1)*nm2vls)):(wheremv2[1,i]+((omnilp-1)*nm2vls)+(nm2vls-1)),omnilp]<-t(mod2val)
                          if ((nm1vls+nm2vls) < ncol(probeval))
                          {
                            intlen<-(nm1vls*nm2vls)
                            modintvl<-probeval[probek,(nm1vls+nm2vls+1):ncol(probeval)]
                            omni[(wherexwz[1,i]+((omnilp-1)*intlen)):(wherexwz[1,i]+((omnilp-1)*intlen)+(intlen-1)),omnilp]<-t(modintvl)
                          }
                        }
                      }
                      fresult2<-process.ftest3(omni,b,varb,0,brsq2,1,y,x)
                      if ((i == (nms+nys)) & (ydich==1))
                      {
                        fratio<-fresult2[1,1]*nxvls
                        pfr<-(1-pchisq(fratio,df=nxvls))
                        fresult2<-cbind(fratio,nxvls,pfr)
                      }
                      resultm2<-matrix(99999,nrow(fresult2),maxresm)
                      resultm2[1:nrow(fresult2),1:ncol(fresult2)]<-fresult2
                      resultm<-rbind(resultm,resultm2)
                      fresult2<-noquote(matrix(sprintf(decimals,fresult2),nrow=nrow(fresult2)))
                      if (outscreen==1)
                      {
                        if (i < (nms + nys) | (ydich == 0))
                        {
                          cat("\nTest of equality of conditional means\n")            
                          colnames(fresult2)<-c(hcflab,"df1","df2","p")
                          rownames(fresult2)<-" "
                          print(fresult2,right=T)
                        }
                        if ((i==(nms + nys)) & (ydich==1))
                        {
                          cat("\nTest of equality of conditional logits or probabilities\n")
                          rownames(fresult2)<-" "
                          colnames(fresult2)<-c("Chi-sq","df","p")
                          print(fresult2,right=T)            
                        }
                      }
                      probetmp<-probeplt[pstart:(pstart+nxvls),1]
                      probetmp<-cbind(probetmp,probeplt[pstart:(pstart+nxvls),prevloc:ncol(probeplt)])
                      pstart<-pstart+(nxvls+1)
                      resultm2<-matrix(99999,nrow(probetmp),maxresm)
                      resultm2[1:nrow(probetmp),1:ncol(probetmp)]<-probetmp
                      resultm<-rbind(resultm,resultm2)
                      probetmp<-noquote(matrix(sprintf(decimals,probetmp),nrow=nrow(probetmp)))
                      if (outscreen==1)
                      {
                        if (i < ((nms + nys)) | (ydich==0))
                        {   
                          clabtmp<-cbind(xnames, outnames[i,1], hclab, "LLCI", "ULCI")
                          cat("\nEstimated conditional means being compared:\n")
                          colnames(probetmp)<-clabtmp[1:ncol(probetmp)]
                          rownames(probetmp)<-t(matrix(replicate(nrow(probetmp)," ")))
                          print(probetmp,right=T)
                        }
                        if ((i==(nms+nys)) & (ydich==1))
                        {
                          cat("\nEstimated conditional logits and probabilities:\n")
                          clabtmp<-c(xnames, outnames[i,1], "prob")
                          probetm2<-matrix(probetmp[,1:2],ncol=2)
                          probetm2<-cbind(probetm2,probetmp[,ncol(probetmp)])
                          probetm2<-noquote(probetm2)
                          colnames(probetm2)<-clabtmp[1:ncol(probetm2)]
                          rownames(probetm2)<-t(matrix(replicate(nrow(probetm2)," ")))
                          print(probetm2,right=T)
                        }
                      }
                    }
                  }
                  # end K
                }
              }
              # end L
              # start N
              if ((probexpl==1) & (printpbe <= intprobe))
              {
                resultm2<-matrix(99999,nrow(modvals2),maxresm)
                resultm2[1:nrow(modvals2),1:ncol(modvals2)]<-modvals2
                resultm<-rbind(resultm,resultm2)
                modva2<-noquote(matrix(sprintf(decimals,modvals2),nrow=nrow(modvals2)))
                colnames(modva2)<-problabs[1:ncol(modva2)]
                rownames(modva2)<-t(matrix(replicate(nrow(modva2)," ")))
                if (outscreen==1)
                {cat("\nConditional effects of the focal predictor at values of the moderator(s):\n")
                  print(modva2,right=T)}
                didsome<-1
                if (hasz==1){printz<-1}
                if (hasw==1){printw<-1}
                # start M (JN method)
                if ((jn==1) & (jnok==1))
                {
                  if (criterr==0)
                  {
                    dfres<-(n-nrow(b));
                    roots<-as.matrix(99999)
                    jncrit<-(dfres* (exp((dfres-(5/6))*((xp2/(dfres-(2/3)+(.11/dfres)))*(xp2/(dfres-(2/3)+(.11/dfres)))))-1))
                    if ((i ==(nms+nys)) & (ydich==1)){jncrit<-xp2*xp2}
                    jnb1<-b[wherejn1,1]
                    jnb3<-b[wherejn3,1]
                    jnsb1<-varb[wherejn1,wherejn1]
                    jnsb3<-varb[wherejn3,wherejn3]
                    jnsb1b3<-varb[wherejn1,wherejn3]
                    ajn<-(jncrit*jnsb3)-(jnb3*jnb3)
                    bjn<-2*((jncrit*jnsb1b3)-(jnb1*jnb3))
                    cjn<-((jncrit*jnsb1)-(jnb1*jnb1))
                    radarg<-(bjn*bjn)-(4*ajn*cjn)
                    den<-2*ajn
                    nrts<-0
                    if ((radarg >= 0) & (den != 0))
                    {
                      x21<-(-bjn+sqrt(radarg))/den
                      x22<-(-bjn-sqrt(radarg))/den
                      if ((x21 >= jnmin) & (x21 <= jnmax))
                      {nrts<-1;roots<-rbind(roots,x21)}
                      if ((x22 >= jnmin) & (x22 <= jnmax))
                      {nrts<-nrts+1;roots<-rbind(roots,x22)}
                      roots<-cbind(roots,matrix(0,nrow(roots),2))
                    }
                    if (nrts > 0)
                    {	                          
                      roots<-matrix(roots[2:nrow(roots),1:3],ncol=3)
                      roots[1,2]<-sum((as.numeric(jnmod < roots[1,1]))/n)*100
                      roots[1,3]<-sum((as.numeric(jnmod > roots[1,1]))/n)*100
                      if (nrow(roots)==2)
                      {
                        roots[2,2]<-sum((as.numeric(jnmod < roots[2,1]))/n)*100
                        roots[2,3]<-sum((as.numeric(jnmod > roots[2,1]))/n)*100
                      }
                      roots2<-noquote(matrix(sprintf(decimals,roots),nrow=nrow(roots)))
                      colnames(roots2)<-c("Value", "% below", "% above")
                      rownames(roots2)<-t(matrix(replicate(nrow(roots)," ")))
                      if (nrts==1){tmprts=matrix(99999,1,3);roots<-rbind(roots,tmprts)}
                      if (outscreen==1)
                      {cat("\nModerator value(s) defining Johnson-Neyman significance region(s):\n")
                        print(roots2,right=T)}
                    }
                    if (nrts==0)
                    {
                      roots<-as.matrix(c(99999,99999))
                      if (outscreen==1)
                      {cat("\nThere are no statistical significance transition points within the observed\n")
                        cat("range of the moderator found using the Johnson-Neyman method.\n")}
                    }
                    resultm2<-matrix(99999,nrow(roots),maxresm)
                    resultm2[1:nrow(roots),1:ncol(roots)]<-roots
                    resultm<-rbind(resultm,resultm2)
                    jnvals<-matrix(0,23,7)
                    for (jni in (0:(21-nrts)))
                    {jnvals[(jni+1),1]<-jnmin+(jni*((jnmax-jnmin)/(21-nrts)))}
                    if (nrts > 0)
                    {
                      for (jni in (1:nrts))
                      {
                        for (jnj in (2:nrow(jnvals)))
                        {
                          if ((roots[jni,1] > jnvals[(jnj-1),1]) & (roots[jni,1] < jnvals[jnj,1]))
                          {
                            jnvals[(jnj+1):(21+jni),1]<-jnvals[jnj:(20+jni),1]
                            jnvals[jnj,1]<-roots[jni,1]
                          }
                        }
                      }
                    }
                    jnvals<-jnvals[1:22,] 
                    for (jni in (1:nrow(jnvals)))
                    {
                      jnvals[jni,2]<-jnb1+jnb3*jnvals[jni,1]
                      jnvals[jni,3]<-sqrt(jnsb1+2*jnvals[jni,1]*jnsb1b3+(jnvals[jni,1]*jnvals[jni,1])*jnsb3)
                      jnvals[jni,4]<-jnvals[jni,2]/jnvals[jni,3]
                      jnvals[jni,5]<-2*pt(-abs(jnvals[jni,4]), df=dfres)
                      jnvals[jni,6]<-jnvals[jni,2]-sqrt(jncrit)*jnvals[jni,3]
                      jnvals[jni,7]<-jnvals[jni,2]+sqrt(jncrit)*jnvals[jni,3]
                      if ((i==(nms + nys)) & (ydich==1))
                      {
                        jnvals[jni,5]<-2*(1-pnorm(abs(jnvals[jni,4])))
                        jnvals[jni,6]<-jnvals[jni,2]-xp2*jnvals[jni,3]
                        jnvals[jni,7]<-jnvals[jni,2]+xp2*jnvals[jni,3]
                      }
                    } 
                    resultm2<-matrix(99999,nrow(jnvals),maxresm)
                    resultm2[1:nrow(jnvals),1:ncol(jnvals)]<-jnvals
                    resultm<-rbind(resultm,resultm2)
                    jnvals2<-noquote(matrix(sprintf(decimals,jnvals),nrow=nrow(jnvals)))
                    rownames(jnvals2)<-t(matrix(replicate(nrow(jnvals)," ")))
                    if ((i < nms+nys) | (ydich==0))
                    {colnames(jnvals2)<-c(jnmodlab,"effect", hclab, "t", "p", "LLCI", "ULCI")}          
                    if ((i==(nms+nys)) & (ydich==1))
                    {colnames(jnvals2)<-c(jnmodlab,"effect","se","Z", "p", "LLCI", "ULCI")}
                    if (outscreen==1)
                    {
                      if (((wcmat[(i+1),jmed]==1) | (zcmat[(i+1),jmed]==1)) & (wzcmat[(i+1),jmed]==0))
                      {cat("\nConditional effect of focal predictor at values of the moderator:\n");print(jnvals2,right=T)}          
                      if ((jmed==1) & (wzcmat[(i+1),jmed]==1))
                      {cat("\nConditional X*W interaction at values of the moderator Z:\n");print(jnvals2,right=T)}   
                      if ((jmed > 1) & (wzcmat[(i+1),jmed]==1))
                      {cat("\nConditional M*W interaction at values of the moderator Z:\n");print(jnvals2,right=T)} 
                    }  
                  }
                }
                # END M (JN method)
              }
              if ((i==(nms+nys)) & (jmed==1) & (bcmat[nrow(bcmat),1]==1))
              {
                if (probei==1)
                {direfflb<-problabs;direff<-modvals2}
                if (probei>1)
                {direff<-rbind(direff,modvals2)}  
              }
              intprint<-0
              #  This does the contrast for conditional effects */
              if ((jmed==1) & (i==1) & (nms==0) & (modcok==1))
              {
                contvec2=matrix(1,2,1)
                contvec2<-cbind(contvec2,wcontval,zcontval)
                if (wzcmat[(i+1),jmed]==1)
                {
                  for (conti in (1:ncol(wcontval)))
                  {
                    for (contj in (1:ncol(zcontval)))
                    {contvec2<-cbind(contvec2,wcontval[,conti]*zcontval[,contj])}
                  }
                }
                conteff<-contvec2%*%probcoef
                contdiff<-matrix(contvec2[1,]-contvec2[2,])
                contse<-sqrt(t(contdiff)%*%probvarb%*%contdiff)
                conteffd<-conteff[1,1]-conteff[2,1]
                contvec=cbind(contvec,conteff)
                contvecm<-contvec
                resultm2<-matrix(99999,nrow(contvecm),maxresm)
                resultm2[1:nrow(contvecm),1:ncol(contvecm)]<-contvecm
                resultm<-rbind(resultm,resultm2)
                contvecm<-noquote(matrix(sprintf(decimals,contvecm),nrow=nrow(contvecm)))
                rownames(contvecm)<-c("Effect1:","Effect2:")
                colnames(contvecm)<-problabs[1:3]
                if (outscreen==1)
                {cat("\nContrast between conditional effects of X:\n")
                  print(contvecm,right=T)}
                if (ydich==0)
                {
                  p<-2*pt(-abs(conteffd/contse),df=dfres)
                  contvec<-cbind(conteffd,contse,conteffd/contse, p)
                  contvec<-cbind(contvec,(conteffd-(tval*contse)))
                  contvec<-cbind(contvec,(conteffd+(tval*contse)))
                  contlabs<-c("Contrast", hclab, "t", "p", "LLCI", "ULCI")
                }
                if (ydich==1)
                { 
                  p<-2*(1-pnorm(abs(conteffd/contse)))
                  contvec<-cbind(conteffd,contse,conteffd/contse, p)
                  contvec<-cbind(contvec,(conteffd-(xp2*contse)))
                  contvec<-cbind(contvec,(conteffd+(xp2*contse)))
                  contlabs<-c("Contrast", "se", "Z", "p", "LLCI", "ULCI")
                } 
                resultm2<-matrix(99999,nrow(contvec),maxresm)
                resultm2[1:nrow(contvec),1:ncol(contvec)]<-contvec
                resultm<-rbind(resultm,resultm2)
                contvec<-noquote(matrix(sprintf(decimals,contvec),nrow=nrow(contvec)))
                colnames(contvec)<-contlabs
                rownames(contvec)<-" "
                if (outscreen==1)
                {cat("\nTest of Effect1 minus Effect2\n")
                  print(contvec,right=T)}
              }
            }
            # end O

            if ((plot==1) | (plot==2))
            { 
              datalabs<-cbind(t(focpred[,2]),outnames[i,1])
              if (plot==2)
              {datalabs<-c(datalabs,"se", "LLCI", "ULCI")}
              if ((i==(nms+nys)) & (ydich==1))
              {datalabs<-c(datalabs,"prob")}
              resultm2<-matrix(99999,nrow(probeplt),maxresm)
              resultm2[1:nrow(probeplt),1:ncol(probeplt)]<-probeplt
              resultm<-rbind(resultm,resultm2)
              probepnt<-noquote(matrix(sprintf(decimals,probeplt),nrow=nrow(probeplt)))
              colnames(probepnt)<-datalabs
              rownames(probepnt)<-t(matrix(replicate(nrow(probepnt)," ")))
              if (outscreen==1)
              {cat("\nData for visualizing the conditional effect of the focal predictor:\n")
                print(probepnt,right=T)}
            }    
          }      
          #End E
        }
        #End R
      }
      #End PROBEandPLOT

      #linear sum
      if ((i==(nms+nys)) & (model >= 0) & (model < 4) & (linsum[1,1] != -999))
      {
        lhyprob<-1;meansub<-0
        if (((nlinsum==nrow(b)) | (nlinsum==(nrow(b)-ncs))))
        {
          if ((nlinsum ==(nrow(b)-ncs)) & (ncs > 0))
          {
            linsum<-cbind(linsum,covmeans)
            meansub<-1
          }
          lhyprob<-0;hypest<-linsum%*%b
          sehypest<-sqrt(linsum%*%varb%*%t(linsum))
          phypest<-2*(pt((-abs(hypest/sehypest)),dfres))
          hypest<-matrix(c(hypest,sehypest,(hypest/sehypest),phypest,(hypest-tval*sehypest),(hypest+tval*sehypest)),ncol=6)
          resultm2<-matrix(99999,ncol(linsum),maxresm)
          resultm2[,1]<-t(linsum)
          resultm<-rbind(resultm,resultm2)
          resultm2<-matrix(99999,1,maxresm)
          resultm2[1,1:ncol(hypest)]<-hypest
          resultm<-rbind(resultm,resultm2)
          if (outscreen==1)
          {
            hyplabs<-c("Estimate",hclab,"t","p","LLCI","ULCI")
            cat("\n----------\n")
            cat("Linear Combination Estimate and Hypothesis Test\n")
            linsumpt<-noquote(matrix(sprintf(decimals,linsum),ncol=1))
            colnames(linsumpt)<-"weight"
            rownames(linsumpt)<-vlabsm
            cat("\nWeight vector:\n")
            print(linsumpt,right=TRUE)
            hypestpt<-noquote(matrix(sprintf(decimals,hypest),ncol=6))
            colnames(hypestpt)<-hyplabs
            rownames(hypestpt)<-" "
            cat("\n")
            print(hypestpt,right=TRUE)
            if (meansub==1)
            {cat("\nCovariate weight(s) set to the sample mean.\n")}
          }
        }
        if (lhyprob==1){notecode[notes,1]<-30;notes<-notes + 1}
      }
    }
    #END G LOOP
    if ((criterr==0) & (dototal==1))
    {
      if (outscreen==1)
      {
        cat("\n************************ TOTAL EFFECT MODEL *************************** \n")
        cat("Outcome Variable: ")
        write.table(outnames[nrow(outnames),1],quote=FALSE,row.names=FALSE,col.names=FALSE)
        cat("\n")
      }
      x<-xtmp
      vlabsm<-matrix(c("constant",xcatlab[1:nxvls,1]))
      if (ncs > 0)
      {
        x<-cbind(as.matrix(x),as.matrix(ctmp))
        vlabsm<-matrix(c(vlabsm,t(covnames)))
      }
      x<-cbind(ones,x)
      modoutz<-process.modelest(y,x,1,1,xp2,hc)
      modsum<-matrix(unlist(modoutz[3]))
      resultm2<-matrix(99999,1,maxresm)
      resultm2[1,1:nrow(modsum)]<-t(modsum)
      resultm<-rbind(resultm,resultm2)
      modsum<-noquote(matrix(sprintf(decimals,modsum),nrow=1))
      modres<-matrix(unlist(modoutz[1]),nrow=ncol(x))
      toteff<-matrix(modres,nrow=ncol(x))
      toteff<-matrix(toteff[2:(1+nxvls),],nrow=nxvls)
      resultm2<-matrix(99999,nrow(modres),maxresm)
      resultm2[1:nrow(modres),1:ncol(modres)]<-modres
      resultm<-rbind(resultm,resultm2)
      modres<-noquote(matrix(sprintf(decimals,modres),nrow=ncol(x)))
      varb<-matrix(unlist(modoutz[6]),nrow=ncol(x))
      brsq2<-modsum[1,2]
      b<-matrix(unlist(modoutz[5]))
      colnames(modsum)<-modsuml;rownames(modsum)<-" "
      rownames(modres)<-vlabsm;colnames(modres)<-modresl
      if (outscreen==1)
      {
        cat("Model Summary: \n")
        print(modsum,right=T)
        cat("\n")
        cat("Model: \n")
        print(modres,right=T)
      }   
      totefflb<-modresl
      toteffl2<-vlabsm[2:(1+nxvls),]
      lmat<-matrix(0,nrow(b),1)
      lmat2<-matrix(1,nxvls,1)
      lmat[2:(1+nxvls),1]<-lmat2
      if (ydich != 1)
      {totomni<-process.ftest3(lmat,b,varb,1,brsq2,0,y,x)} 
      if (stand==1)
      {
        predsd<-matrix(0,nrow(modres),1)
        stdmod<-as.numeric(modres[,1])/ovsd[nrow(ovsd),1]
        for (jd in c(1:ncol(x)))
        {predsd[jd,1]<-sd(x[,jd])}   
        if ((wherex[1,ncol(wherex)] != -999) & ((nxvls > 1) | (xdich==1)))  
        {  
          sdmsone<-matrix(1,nxvls,1)
          predsd[wherex[1,ncol(wherex)]:wherex[2,ncol(wherex)],1]<-sdmsone    
          pstog<-1    
        }
        predsd[1,1]<-1
        stdmod<-stdmod*predsd
        stdmod<-matrix(stdmod[2:nrow(stdmod),1])
        sdvlabs<-vlabsm[2:nrow(vlabsm),1]
        resultm2<-matrix(99999,nrow(stdmod),maxresm)
        resultm2[1:nrow(stdmod),1:ncol(stdmod)]<-stdmod
        resultm<-rbind(resultm,resultm2)
        stdmod<-noquote(matrix(sprintf(decimals,stdmod),(nrow(b)-1)))
        colnames(stdmod)<-"coeff"
        rownames(stdmod)<-sdvlabs
        if (outscreen==1)
        {cat("\nStandardized coefficients:\n")
          print(stdmod,right=T)}
      }

      if (covcoeff==1)
      {
        varbpr<-noquote(matrix(sprintf(decimals,varb),nrow=nrow(b)))
        rownames(varbpr)<-vlabsm;colnames(varbpr)<-vlabsm
        if (outscreen==1)
        {
          cat("\nCovariance matrix of regression parameter estimates:\n")
          print(varbpr,right=T)
          resultm2<-matrix(99999,nrow(varb),maxresm)
          if (ncol(varb) <= maxresm)
          {
            resultm2[1:nrow(varb),1:ncol(varb)]<-varb
            resultm<-rbind(resultm,resultm2)
          }
          if (ncol(varb) > maxresm)
          {
            resultmt<-matrix(99999,nrow(resultm),ncol(varb))
            resultmt[1:nrow(resultm),1:ncol(resultm)]<-resultm
            resultm<-resultmt
            resultm2<-matrix(99999,nrow(varb),ncol(resultm))
            resultm2[1:nrow(varb),1:ncol(varb)]<-varb
            resultm<-rbind(resultm,resultm2)
            maxresm<-ncol(resultm)
          }
        }
      } 
    }
  }
  #END cycle through the models

  if ((criterr==0) & (nms > 0) & (ydich==0) & (modelres==1))
  {
    modresid<-matrix(modresid[,2:ncol(modresid)],ncol=(ncol(modresid)-1))
    corall<-cor(modresid)
    resultm2<-matrix(99999,nrow(corall),maxresm)
    resultm2[1:nrow(corall),1:ncol(corall)]<-corall
    resultm<-rbind(resultm,resultm2)
    corall2<-noquote(matrix(sprintf(decimals,corall),nrow=nrow(corall)))
    colnames(corall2)<-outnames
    rownames(corall2)<-outnames
    if (outscreen==1)
    {cat("\n*************** CORRELATIONS BETWEEN MODEL RESIDUALS **************** \n\n")
      print(corall2,right=T)}
  }

  #DO BOOTSTRAPPING
  if ((criterr==0) & (boot > 0))
  { 
    bootres<-matrix(-999,1,sum(nump))
    bootdir<-obsdirfx
    if (effsize==1){bootysd<-matrix(-999,1,1);bootxsd<-matrix(-999,1,1)}
    badboot<-0;goodboot<-0;smallest<-1;booting<-1
    j<-1
    if (outscreen==1)
    {
      cat("\n*********************************************************************** \n")
      if (progress==1)
      {cat("Bootstrapping progress:\n")
        bootprog<-txtProgressBar(min=0,max=boot,char=">",width=62,style=3)}
      if (progress !=1){cat("Bootstrapping in progress. Please wait.\n");flush.console()}
    }
    while ((goodboot < boot) & (j <= maxboot))
    {
      if ((outscreen==1) & (progress==1)){setTxtProgressBar(bootprog,j)}   
      nobootx<-1;modres2<-999;bad<-0;
      v<-as.matrix(trunc(runif(n)*n)+1)
      for (i in (1:(nms+nys)))
      {
        y<-as.matrix(outvars[v,i])
        ynovar<-sum((y-(sum(y)/nrow(y)))*(y-(sum(y)/nrow(y))))
        if (ynovar==0){bad<-1}
        xindx<-datindx[1:(nump[1,i]-1),i]
        hello<-0
        x<-as.matrix(fulldat[v,xindx])
        x<-cbind(ones,x)
        xsq<-t(x)%*%x
        exsq<-eigen(xsq)
        exsq<-matrix(unlist(exsq[1]))
        holymoly<-min(exsq)
        zeroeig<-sum(as.numeric(exsq <= 0.000000000002))
        bad<-bad+as.numeric(zeroeig > 0)
        bad=bad+as.numeric(sd(y)==0)
        if (bad==0)
        {
          if (holymoly < smallest){smallest<-holymoly}
          if ((ydich==0) | (i < (nms+nys)))
          {modrest<-process.modelest(y,x,type=1,full=0,xp2,hc)}
          if ((ydich==1) & (i==(nms+nys)))
          {modrest<-process.modelest(y,x,3,0,xp2,hc,iterate,converge)}
          modres2<-as.matrix(c(modres2,modrest))
          if (i==(nms+nys))
          {
            if (bcmat[(i+1),1]==1)
            {bootdir<-rbind(bootdir,t(modrest[wherex[1,i]:wherex[2,i],1]))}             
            if (bcmat[(i+1),1]==0)
            {bootdir<-rbind(bootdir,dirzes)}         
          }
          if ((bcmat[(i+1),1]==1) & (nobootx==1) & (effsize==1))
          {
            nobootx<-0
            xsdtemp<-sum((x[,2]-(sum(x[,2])/nrow(x)))*(x[,2]-(sum(x[,2])/nrow(x))))
            xsdtemp<-sqrt(xsdtemp/(nrow(x)-1))
          }
        }
      }
      if (bad==0)
      {
        modres2<-t(modres2[2:nrow(modres2),])
        #modres2<-matrix(modres2[,2:nrow(modres2),ncol=ncol(modres)-1)
        bootres<-rbind(bootres,modres2)
        if (effsize==1)
        {
          ysdtemp<-sd(y)
          bootysd<-rbind(bootysd,ysdtemp)
          bootxsd<-rbind(bootxsd,xsdtemp)
        }
        goodboot<-goodboot+1
      }
      if (bad != 0){badboot<-badboot+1}
      j<-j+1
    }
    bootres<-as.matrix(bootres[2:nrow(bootres),])
    if (effsize==1)
    {
      bootysd<-matrix(bootysd[2:nrow(bootysd),])
      if (nrow(bootxsd) > 1){bootxsd<-matrix(bootxsd[2:nrow(bootxsd)])}
    }
    if (goodboot < (boot))
    {boot<-0;modelbt<-0;notecode[notes,1]<-7;notes<-notes+1}
    if (boot > 0)
    {
      if (effsize==1){bootysd<-rbind(ysd,bootysd);bootxsd<-rbind(xsd,bootxsd)}
      if (saveboot==1)
      {
        savlabs<-matrix(" ",ncol(bootres))
        for (i in c(1:ncol(bootres))){savlabs[i,1]<-paste("col",i,sep='')}
        boots<-as.data.frame(bootres)
        colnames(boots)<-savlabs
        #assign("process.boots", boots, envir = .GlobalEnv)
      }
      if (modelbt==1)
      {   
        bootcim<-matrix(-99999,ncol(bootres),5)
        bootcim[,2]<-t(colSums(bootres)/nrow(bootres))
        bootcim[,1]<-as.numeric(coeffmat[2:nrow(coeffmat),1])
        for (i in (1:ncol(bootres)))
        { 
          if (bc==0){bootcim[i,3:5]<-t(process.pboot3(bootres[,i],cilow,cihigh))}
          if (bc==1)
          {
            bcbout<-process.bcboot3(bootres[,i],bootcim[i,1],xp2,badend,priorlo,priorhi)
            bootcim[i,3:5]<-t(matrix(unlist(bcbout[1])))
            badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
          }
        }
      } 
    }

    if (badboot > 0){notecode[notes,1]<-6;notes<-notes+1}
    if ((outscreen==1) & (progress==1)){close(bootprog)}
  }
  #end of bootstrapping

  #Here is for the indirect effects
  if ((criterr==0) & (nms > 0))
  {
    paths<-matrix(paths[,2:ncol(paths)],nrow=nrow(paths))
    pathsw<-matrix(pathsw[,2:ncol(pathsw)],nrow=nrow(pathsw))
    pathsz<-matrix(pathsz[,2:ncol(pathsz)],nrow=nrow(pathsz))
    pathswz<-matrix(pathswz[,2:ncol(pathswz)],nrow=nrow(pathswz))
    pathsmod<-pathsw+pathsz+pathswz
    pathsdv<-t(pathsdv[2:nrow(pathsdv),])
    pathsfoc<-matrix(pathsfoc[,2:ncol(pathsfoc)],nrow=nrow(pathsfoc))
    pathtype<-matrix(pathtype[,2:ncol(pathtype)],nrow=nrow(pathtype))
    anymod<-as.numeric(sum(pathsmod) > 0)
    obscoeff<-t(as.matrix(obscoeff[1,2:ncol(obscoeff)]))
    if (outscreen==1)
    {
      if ((dototal==0) & (alttotal==0))
      {cat("\n**************** DIRECT AND INDIRECT EFFECTS OF X ON Y ****************\n")}
      if (alttotal==1)
      {cat("\n************ TOTAL, DIRECT AND INDIRECT EFFECTS OF X ON Y ****************\n")}
    }
    if (dototal==1)
    {
      if (outscreen==1)
      {cat("\n************ TOTAL, DIRECT, AND INDIRECT EFFECTS OF X ON Y ************\n")}   
      totefflb[1,1]<-"effect"
      if (counterf != 1)
      {
        if (effsize==1)
        {
          toteffsz<-toteff[,1]/ysd 
          if ((xdich==1) | (mcx > 0)){totefflb<-cbind(totefflb,"c_ps")}
          if ((xdich==0) & (mcx==0))
          {toteffsz<-(toteffsz*xsd);totefflb<-cbind(totefflb,"c_cs")}     
          toteff<-cbind(toteff,toteffsz)
        }
        resultm2<-matrix(99999,nrow(toteff),maxresm)
        resultm2[1:nrow(toteff),1:ncol(toteff)]<-toteff
        resultm<-rbind(resultm,resultm2)
        toteff2<-noquote(matrix(sprintf(decimals,toteff),nrow=nrow(toteff)))
        colnames(toteff2)<-totefflb
        if (nxvls > 1)
        {
          rownames(toteff2)<-toteffl2
          if (outscreen==1)
          {cat("\nRelative total effects of X on Y:\n")
            print(toteff2,right=T)}
          clabtmp<-c("R2-chng", hcflab, "df1","df2","p")
          resultm2<-matrix(99999,nrow(totomni),maxresm)
          resultm2[1:nrow(totomni),1:ncol(totomni)]<-totomni
          resultm<-rbind(resultm,resultm2)
          totomni2<-noquote(matrix(sprintf(decimals,totomni),nrow=nrow(totomni)))
          colnames(totomni2)<-clabtmp
          rownames(totomni2)<-" "
          if (outscreen==1)
          {
            cat("\nOmnibus test of total effect of X on Y:\n")
            print(totomni2,right=T)
            cat("----------\n")
          }
        }
        if (nxvls < 2)
        {
          rownames(toteff2)<-" "
          if (outscreen==1)
          {cat("\nTotal effect of X on Y:\n")
            print(toteff2,right=T)}
        }
      }
    }
    moddir<-wcmat[nrow(bcmat),1]+zcmat[nrow(bcmat),1]
    if ((bcmat[nrow(bcmat),1]==1) & (counterf != 1))
    {
      if (ydich==1)
      {
        direfflb[,(ncol(direfflb)-5):ncol(direfflb)]<-t(matrix(c("Effect","se","Z","p","LLCI","ULCI")))
      } 
      if (moddir==0){direfflb[1,1]="effect"}
      if ((effsize==1) & (moddir==0) & (anymod == 0))
      {
        direffsz<-direff[,1]/ysd
        if ((xdich==1) | (mcx > 0)){direfflb<-cbind(direfflb,"c'_ps")}
        if ((xdich==0) & (mcx==0))
        {direffsz<-(direffsz*xsd);direfflb<-cbind(direfflb,"c'_cs")}    
        direff<-cbind(direff,direffsz)
      }
      resultm2<-matrix(99999,nrow(direff),maxresm)
      resultm2[1:nrow(direff),1:ncol(direff)]<-direff
      resultm<-rbind(resultm,resultm2)
      direff2<-noquote(matrix(sprintf(decimals,direff),nrow=nrow(direff)))
      colnames(direff2)<-direfflb
      if ((moddir==0) & (nxvls==1) & (outscreen==1))
      {
        rownames(direff2)<-" "
        cat("\nDirect effect of X on Y:\n")
        print(direff2,right=T)
      }
      if ((moddir==0) & (nxvls>1))  
      {
        rownames(direff2)<-direffl2
        if (outscreen==1)
        {cat("\nRelative direct effects of X on Y:\n")
          print(direff2,right=T)} 
        resultm2<-matrix(99999,nrow(diromni),maxresm)
        resultm2[1:nrow(diromni),1:ncol(diromni)]<-diromni
        resultm<-rbind(resultm,resultm2)
        diromni2<-noquote(matrix(sprintf(decimals,diromni),nrow=nrow(diromni)))
        rownames(diromni2)<-" "
        if (ydich==0)
        {  
          colnames(diromni2)<-c("R2-chng", hcflab, "df1","df2","p")
          if (outscreen==1)
          {cat("\nOmnibus test of direct effect of X on Y:\n")
            print(diromni2,right=T)}
        }
        if (ydich==1)
        {  
          colnames(diromni2)<-c("Chi-sq", "df", "p")
          if (outscreen==1)
          {cat("\nOmnibus likelihood ratio test of direct effect of X on Y:\n")
            print(diromni2,right=T)}
        }
        if (outscreen==1)
        {cat("\n----------\n")}
      }
      if ((moddir > 0) & (nxvls==1) & (outscreen==1))
      {cat("\nConditional direct effect(s) of X on Y:\n")
        rownames(direff2)<-t(matrix(replicate(nrow(direff2)," ")))
        print(direff2,right=T)}
      direffl4<-direffl2
      if ((moddir > 0) & (nxvls>1))
      {
        direffl2=" "
        for (i in (1:nxvls))
        {
          for (j in (1:(nrow(direff)/nxvls)))
          {direffl2<-rbind(direffl2,xcatlab[i,1])}     
        }
        direffl2<-direffl2[2:nrow(direffl2),1]
        rownames(direff2)<-direffl2
        if (outscreen==1)
        {cat("\nRelative conditional direct effect(s) of X on Y:\n")
          print(direff2,right=T)}
      }
      direffl2<-direffl4    
    }

    if ((bcmat[nrow(bcmat),1]==0) & (counterf != 1) & (outscreen==1))
    {cat("\nThe direct effect of X on Y is fixed to zero.\n")}

    #Here is the start of the indirect effects
    if (nms==1){indmark<-matrix(2);indsets<-t(matrix(c(1,2)))}
    if (nms==2){indmark<-t(matrix(c(2,2,3)));indsets<-t(matrix(c(1,4,2,5,1,3,5)));thetam<-matrix(1)}
    if (nms==3) 
    {indmark<-t(matrix(c(2,2,2,3,3,3,4)))
    indsets<-t(matrix(c(1,7,2,8,4,9,1,3,8,1,5,9,2,6,9,1,3,6,9)))
    thetam<-t(matrix(c(1,2,3)))}
    if (nms==4)
    {indmark<-t(matrix(c(2,2,2,2,3,3,3,3,3,3,4,4,4,4,5)))
    indsets<-t(matrix(c(1,11,2,12,4,13,7,14,1,3,12,1,5,13,1,8,14,2,6)))
    indsets<-cbind(indsets,t(matrix(c(13,2,9,14,4,10,14,1,3,6,13,1,3,9,14,1,5,10,14,2,6,10,14,1,3,6,10,14))))
    thetam<-t(matrix(c(1,2,4,3,5,6)))}
    if (nms==5)
    {indmark<-t(matrix(c(2,2,2,2,2,3,3,3,3,3,3,3,3,3,3,4,4,4,4,4,4,4,4,4,4,5,5,5,5,5,6)))
    indsets<-t(matrix(c(1,16,2,17,4,18,7,19,11,20,1,3,17,1,5,18,1,8,19,1,12,20,2,6,18,2,9,19,2)))
    indsets<-cbind(indsets,t(matrix(c(13,20,4,10,19,4,14,20,7,15,20,1,3,6,18,1,3,9,19,1,3,13,20,1,5,10))))
    indsets<-cbind(indsets,t(matrix(c(19,1,5,14,20,1,8,15,20,2,6,10,19,2,6,14,20,2,9,15,20,4,10,15,20,1))))
    indsets<-cbind(indsets,t(matrix(c(3,6,10,19,1,3,6,14,20,1,3,9,15,20,1,5,10,15,20,2,6,10,15,20,1,3))))
    indsets<-cbind(indsets,t(matrix(c(6,10,15,20))))
    thetam<-t(matrix(c(1,2,5,3,6,8,4,7,9,10)))}
    if (nms==6)
    {indmark<-t(matrix(c(2,2,2,2,2,2,3,3,3,3,3,3,3,3,3,3,3,3,3,3,3,4,4,4,4,4,4,4,4,4,4,4,4,4,4,4,4,4,4,4,4,5,5,5,5,5)))
    indmark<-cbind(indmark,t(matrix(c(5,5,5,5,5,5,5,5,5,5,6,6,6,6,6,6,7))))
    indsets<-t(matrix(c(1,22,2,23,4,24,7,25,11,26,16,27,1,3,23,1,5,24,1,8,25,1,12,26,1,17,27,2,6,24,2,9,25,2,13,26,2,18)))
    indsets<-cbind(indsets,t(matrix(c(27,4,10,25,4,14,26,4,19,27,7,15,26,7,20,27,11,21,27,1,3,6,24,1,3,9,25,1,3,13,26,1,3))))
    indsets<-cbind(indsets,t(matrix(c(18,27,1,5,10,25,1,5,14,26,1,5,19,27,1,8,15,26,1,8,20,27,1,12,21,27,2,6,10,25,2,6,14,26))))
    indsets<-cbind(indsets,t(matrix(c(2,6,19,27,2,9,15,26,2,9,20,27,2,13,21,27,4,10,15,26,4,10,20,27,4,14,21,27,7,15,21,27))))
    indsets<-cbind(indsets,t(matrix(c(1,3,6,10,25,1,3,6,14,26,1,3,6,19,27,1,3,9,15,26,1,3,9,20,27,1,3,13,21,27,1,5,10,15,26))))
    indsets<-cbind(indsets,t(matrix(c(1,5,10,20,27,1,5,14,21,27,1,8,15,21,27,2,6,10,15,26,2,6,10,20,27,2,6,14,21,27,2,9,15))))
    indsets<-cbind(indsets,t(matrix(c(21,27,4,10,15,21,27,1,3,6,10,15,26,1,3,6,10,20,27,1,3,6,14,21,27,1,3,9,15,21,27,1,5,10))))
    indsets<-cbind(indsets,t(matrix(c(15,21,27,2,6,10,15,21,27,1,3,6,10,15,21,27))))
    thetam<-t(matrix(c(1,2,6,3,7,10,4,8,11,13,5,9,12,14,15)))}
    if (nms==7)
    {indmark<-t(matrix(c(2,2,2,2,2,2,2)));indsets<-t(matrix(c(1,29,2,30,4,31,7,32,11,33,16,34,22,35)))}
    if (nms==8)
    {indmark<-t(matrix(c(2,2,2,2,2,2,2,2)));indsets<-t(matrix(c(1,37,2,38,4,39,7,40,11,41,16,42,22,43,29,44)))}
    if (nms==9)
    {indmark<-t(matrix(c(2,2,2,2,2,2,2,2,2)));indsets<-t(matrix(c(1,46,2,47,4,48,7,49,11,50,16,51,22,52,29,53,37,54)))}
    if (nms==10)
    {indmark<-t(matrix(c(2,2,2,2,2,2,2,2,2,2))) 
    indsets<-t(matrix(c(1,56,2,57,4,58,7,59,11,60,16,61,22,62,29,63,37,64,46,65)))}
    indlbl<-"Ind1"
    for (indb in (2:90))
    {indlbl<-cbind(indlbl,paste("Ind",indb,sep=''))}
    indlbl<-matrix(indlbl)
    cntname<-"(C1)"
    for (indb in (2:105))
    {cntname<-cbind(cntname,paste("(C",indb,")",sep=''))}
    cntname<-matrix(cntname)
    indmake<-matrix(0,ncol(indmark),(nms+2))
    indmod<-matrix(999,ncol(indmark),1)
    indmmm<-matrix(0,ncol(indmark),1)
    indmmmt<-matrix(0,ncol(indmark),1)
    start<-1;end<-0;nindfx<-0
    indlocs<-matrix(999,nrow(thetaxmb),ncol(paths))
    indkey<-matrix("      ",ncol(indmark),1+((max(indmark)*2)+1))
    c1<-1;c2<-1;c3<-1
    for (i in (1:ncol(paths)))
    {
      if (pathtype[1,i]==1){indlocs[,i]<-thetaxmb[,c1];c1<-(c1+1)}
      if (pathtype[1,i]==3){indlocs[,i]<-thetamyb[,c2];c2<-(c2+1)}
      if ((pathtype[1,i]==2) & (nms < 7) & (serial==1)){indlocs[,i]<-thetammb[,thetam[1,c3]];c3<-(c3+1)}
    }
    for (i in (1:ncol(indlocs)))
    {
      c1<-2
      for (j in (2:nrow(indlocs)))
      {
        if (indlocs[j,i] != 0){indlocs[c1,i]<-indlocs[j,i];c1<-(c1+1)}
      }
      indlocs[1,i]<-(c1-2)
    }
    indlocs<-indlocs[1:max((indlocs[1,])+1),]
    for (i in (1:ncol(indmark)))
    { 
      numget<-indmark[1,i];end<-end+numget;gotcha<-t(matrix(indsets[1,start:end]))
      start<-end+1;ok<-1;temp<-0;repoman<-matrix(0,4,1)
      for (j in (1:ncol(gotcha)))
      {
        if (paths[1,gotcha[1,j]]==0){ok<-0}
        if (pathsmod[1,gotcha[1,j]] > 0)
        {
          temp<-1
          temp2<-rbind(pathsw[1,gotcha[1,j]],pathsz[1,gotcha[1,j]],pathswz[1,gotcha[1,j]],0)
          temp2<-matrix(temp2)
          if ((temp2[1,1]==1) & (temp2[2,1]==1) & (temp2[3,1]==0)){temp2[4,1]<-1}
          repoman<-repoman+temp2
        }
      }
      temp<-0;tempmmm<-0;typemmm<-0
      if ((repoman[1,1] > 0) & (repoman[2,1]==0))
      {
        temp<-1
        if (repoman[1,1]==1){tempmmm<-1} 
        if ((repoman[1,1] > 1) & ((wdich==1) | (mcw > 0)))
        {
          tempmmm<-12;typemmm<-mcw
          if (wdich==1){typemmm<-1}
        }
        if ((repoman[1,1] > 1) & ((wdich==0) & (mcw==0))){tempmmm<-101} 
      }
      if ((repoman[1,1]==0) & (repoman[2,1] > 0))
      {
        temp<-2
        if (repoman[2,1]==1){tempmmm<-2}
        if ((repoman[2,1] > 1) & ((zdich==1) | (mcz > 0)))
        {
          tempmmm<-22;typemmm<-mcz
          if (zdich==1){typemmm<-1}
        }
        if ((repoman[2,1] > 1) & ((zdich==0) & (mcw==0))){tempmmm<-102}
      }
      if ((repoman[1,1] > 0) & (repoman[2,1] > 0))
      {
        temp<-3
        if ((repoman[1,1]==1) & (repoman[2,1]==1))
        {
          if (repoman[4,1]==1){tempmmm<-31}
          if (repoman[3,1]==1){tempmmm<-41}
        }
      }
      if ((repoman[1,1]==1) & (repoman[2,1]==1) & (repoman[3,1]==0) & (repoman[4,1]==0)){tempmmm<-51}
      if (ok==1)
      {
        nindfx<-nindfx+1
        indmake[nindfx,1]<-numget;indmod[nindfx,1]<-temp;indmmm[nindfx,1]<-tempmmm
        indmmmt[nindfx,1]<-typemmm;indmake[nindfx,2:(1+numget)]<-gotcha;indkey[nindfx,1]<-xnames
        for (j in (1:numget))
        {
          indkey[nindfx,(j*2+1)]<-pathsdv[1,gotcha[1,j]]
          indkey[nindfx,(j*2)]<-"   ->   "
        }
      }
    }
    indkey<-matrix(indkey[1:nindfx,1:((max(indmake[,1])*2)+1)],nrow=nindfx)
    indmake<-matrix(indmake[1:nindfx,1:(max(indmake[,1])+1)],ncol=(max(indmake[,1])+1))
    indmod<-as.matrix(indmod[1:nrow(indmake),1])
    indmmm<-as.matrix(indmmm[1:nrow(indmake),1])
    indmmmt<-as.matrix(indmmmt[1:nrow(indmake),1])
    ncpairs<-(((nindfx)*(nindfx-1))/2)
    if (((contrast==1) | (contrast==2)) & (ncpairs > 105))
    {contrast<-0;notecode[notes,1]<-13;notes<-notes+1}
    if (contrast==3)
    {
      if (ncol(contvec) != nindfx)
      {contrast<-0;notecode[notes,1]<-14;notes<-notes+1}
    }

    #This is for models with no moderator */
    if (anymod==0)
    {
      if ((nms==1) & (contrast > 0)){contrast<-0}
      efloop<-(((1-as.numeric(effsize==0))*2)+1)-((((mcx>0) | (xdich==1)))*(1-as.numeric(effsize==0)))
      for (kk in (1:efloop))
      {
        if (boot==0)
        {
          bootres<-obscoeff
          if (kk==1){totbtvec<-matrix(0,1,nxvls)}
          indtab<-matrix(999);inddiff<-matrix(999);bootysd<-matrix(ysd);bootxsd<-matrix(xsd)
        }    
        if (boot > 0)
        {
          bootres<-rbind(obscoeff,bootres)
          if (kk==1){totbtvec<-matrix(0,nrow(bootres),nxvls)}
          indtab<-matrix(999,1,4);inddiff<-matrix(999,nrow(bootres))
        }
        indtotal<-matrix(0,nrow(bootres),1)
        for (i in (1:nrow(indmake)))
        {
          for (j in (1:nxvls))
          {
            indtemp<-matrix(1,nrow(bootres),1)
            for (k in (1:indmake[i,1]))
            {
              jtemp<-1
              if ((j > 1) & (k==1)){jtemp<-j}
              indtemp<-indtemp*bootres[,pathsfoc[jtemp,indmake[i,(k+1)]]]
            }
            if (kk==2){indtemp<-indtemp/bootysd}
            if (kk==3){indtemp<-((bootxsd*indtemp)/bootysd)}
            if (contrast != 0){inddiff<-cbind(inddiff,indtemp)}
            if (nxvls==1){indtotal<-(indtotal+indtemp)}
            indeff<-indtemp[1,1]
            if (kk==1){totbtvec[,j]<-totbtvec[,j]+indtemp}
            if (boot > 0)
            {
              if (bc==0){bcitmp<-process.pboot3(indtemp[2:nrow(indtemp),1],cilow,cihigh)}
              if (bc==1)
              {
                bcbout<-process.bcboot3(indtemp[2:nrow(indtemp),1],indtemp[1,1],xp2,badend,priorlo,priorhi)
                bcitmp<-matrix(unlist(bcbout[1]))
                badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
              }
              indeff<-cbind(indeff,t(bcitmp))
            }
            if (kk==1){indtabn<-rbind(indtab,indeff)}
            indtab<-rbind(indtab,indeff)
          }
        }
        indtab<-matrix(indtab[2:nrow(indtab),],nrow=(nrow(indtab)-1))
        if (kk==1){indtabn<-matrix(indtabn[2:nrow(indtabn),],nrow=(nrow(indtabn)-1))}
        rowlbs<-matrix(indlbl[1:nrow(indtab),1])
        rowbls3<-rowlbs
        if (mc > 0)
        {
          inddiff<-matrix(-999,mc,1)
          indtab2<-matrix(-999,nrow(indtab),4)
          indtab2[,1]<-indtab;indtab<-indtab2;mcct<-0
          indtotal<-matrix(0,mc,1)
          if (kk==1)
          {
            x1<-sqrt(-2*log(matrix(runif(mc*nrow(mcsopath)),mc,nrow(mcsopath))))*cos((2*3.14159265358979)*matrix(runif(mc*nrow(mcsopath)),mc,nrow(mcsopath)))
            x1<-x1%*%chol(indcov)
            for (ii in (1:nrow(x1))){x1[ii,]<-(x1[ii,]+t(mcsopath))}
          }
          for (ii in (1:nms))
          {
            tmpb<-x1[,((nms*nxvls)+ii)];tmpb2<-tmpb
            if (nxvls > 1)
            {
              for (jj in (1:(nxvls-1)))
              {tmpb2<-cbind(tmpb2,tmpb)}        
            }
            indtemp<-as.matrix(x1[,(((ii-1)*nxvls)+1):(ii*nxvls)]*tmpb2)
            for (jj in (1:ncol(indtemp)))
            {
              if (kk==2){indtemp[,jj]<-indtemp[,jj]/ysd}
              if (kk==3){indtemp[,jj]<-(xsd*indtemp[,jj])/ysd}
              mcicon<-process.pboot3(indtemp[,jj],cilow,cihigh)
              mcct<-mcct+1
              indtab[mcct,2:4]<-t(mcicon)
            }
            if (nxvls==1)
            {
              indtotal<-indtotal+indtemp
              if (contrast != 0){inddiff=cbind(inddiff,indtemp)}
            }
          }
        }
        if ((normal==1) & (sobelok==1))
        {
          sobelmat<-matrix(indtab[,1])
          sobelmat<-cbind(sobelmat,(sobelmat/2),sobelmat,sobelmat)
          for (ii in (1:nms))
          {
            se2b<-(indcov[((nms*nxvls)+ii),((nms*nxvls)+ii)])
            bpath2<-(mcsopath[((nms*nxvls)+ii),1])^2
            se2a<-matrix(diag(matrix(indcov[(((ii-1)*nxvls)+1):(ii*nxvls),(((ii-1)*nxvls)+1):(ii*nxvls)],nrow=nxvls)))
            apath2<-matrix(mcsopath[(((ii-1)*nxvls)+1):(ii*nxvls),1])^2
            sesobel<-sqrt(apath2*se2b+bpath2*se2a+se2a*se2b)
            sobelmat[(((ii-1)*nxvls)+1):(ii*nxvls),2]<-sesobel
          }
          sobelmat[,3]<-sobelmat[,1]/sobelmat[,2]
          sobelmat[,4]<-2*(1-pnorm(abs(sobelmat[,3])))
        }
        if (serial==0){rowlbs<-t(mnames)}
        if ((nxvls==1) & (nms > 1))
        {
          rowlbs<-rbind("TOTAL",rowlbs)
          indtemp<-indtotal[1,1]
          if ((boot > 0) & (nxvls==1))
          {
            if (bc==0){bcitmp<-process.pboot3(indtotal[2:nrow(indtotal),1],cilow,cihigh)}
            if (bc==1)
            {
              bcbout<-process.bcboot3(indtotal[2:nrow(indtotal),1],indtotal[1,1],xp2,badend,priorlo,priorhi)
              bcitmp<-matrix(unlist(bcbout[1]))
              badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
            }
            indtemp<-cbind(indtemp,t(bcitmp))
          }
          if (mc > 0)
          {
            obtmc<-indtab[,1];indtemp<-sum(obtmc)
            mcicon2<-process.pboot3(indtotal[,1],cilow,cihigh)
            indtemp<-cbind(indtemp,t(mcicon2))
          }
          indtab<-rbind(indtemp,indtab)
        }
        bootlbs<-"Effect"
        if (boot > 0){bootlbs<-c("Effect","BootSE","BootLLCI","BootULCI")}
        if (mc > 0){bootlbs<-c("Effect","MC SE","MC LLCI","MC ULCI")}
        if (nxvls==1)
        {
          if (contrast != 0)
          {
            inddiff<-matrix(inddiff[,2:ncol(inddiff)],nrow=nrow(inddiff))
            if (mc > 0){inddiff=rbind(t(obtmc),inddiff)}
            if (contrast==3)
            {
              inddifft<-matrix(inddiff%*%t(contvec))
              indtemp<-inddifft[1,1]
              if ((boot > 0) | (mc > 0))
              {
                if ((mc > 0) | ((boot > 0) & (bc==0)))
                {bcicon<-process.pboot3(inddifft[2:nrow(inddifft),1],cilow,cihigh)}
                if ((boot > 0) & (bc==1))
                {
                  bcbout<-process.bcboot3(inddifft[2:nrow(inddifft),1],inddifft[1,1],xp2,badend,priorlo,priorhi)
                  bcicon<-matrix(unlist(bcbout[1]))
                  badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
                }
                indtemp<-cbind(indtemp,t(bcicon))
              }
              indtab<-rbind(indtab,indtemp)
            }
            if ((contrast==1) | (contrast==2))
            {
              conkey<-matrix(" ",1,4)
              for (i in (1:(ncol(inddiff)-1)))
              {     
                for (j in ((i+1):ncol(inddiff)))
                {           
                  inddifft<-matrix(inddiff[,i]-inddiff[,j])
                  if (contrast==2){inddifft<-matrix(abs(inddiff[,i])-abs(inddiff[,j]))}
                  indtemp<-inddifft[1,1]
                  conkeyt<-cbind(" ", rowlbs[(i+1),1]," minus  ",rowlbs[(j+1),1])
                  conkey<-rbind(conkey,conkeyt)
                  if ((boot > 0) | (mc > 0))
                  {
                    if ((mc > 0) | ((boot > 0) & (bc==0)))
                    {bcitmp2<-process.pboot3(inddifft[2:nrow(inddifft),1],cilow,cihigh)}
                    if ((boot > 0) & (bc==1))
                    {
                      bcbout<-process.bcboot3(inddifft[2:nrow(inddifft),1],inddifft[1,1],xp2,badend,priorlo,priorhi)
                      bcitmp2<-matrix(unlist(bcbout[1]))
                      badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
                    }
                    indtemp<-cbind(indtemp,t(bcitmp2)) 
                  }
                  indtab<-rbind(indtab,indtemp)
                }
              }
            }
            if (contrast != 3)
            {contlbs<-matrix(cntname[1:(((nindfx)*(nindfx-1))/2),1])}
            if (contrast==3)
            {contlbs<-matrix("(C1)")}    
            rowlbs<-rbind(rowlbs,contlbs)
          }
          resultm2<-matrix(99999,nrow(indtab),maxresm)
          resultm2[1:nrow(indtab),1:ncol(indtab)]<-indtab
          resultm<-rbind(resultm,resultm2)
          indtab5<-noquote(matrix(sprintf(decimals,indtab),nrow=nrow(indtab)))
          colnames(indtab5)<-bootlbs
          rownames(indtab5)<-rowlbs
          if (outscreen==1)
          {
            if (kk==1)
            {
              cat("\nIndirect effect(s) of X on Y:\n")
              print(indtab5,right=T)
            }

            if ((kk==2) & ((xdich==1) | (mcx > 0)))
            {
              cat("\nPartially standardized indirect effect(s) of X on Y:\n")
              print(indtab5,right=T)
            }
            if (kk==3)
            {
              cat("\nCompletely standardized indirect effect(s) of X on Y:\n")
              print(indtab5,right=T)
            }
          }
          if ((normal==1) & (sobelok==1) & (kk==1))
          {
            sobellab<-c("Effect",hclab,"Z","p")
            sobelrlb<-rowlbs
            if (nms > 1){sobelrlb<-matrix(rowlbs[2:(1+nms),1])}
            resultm2<-matrix(99999,nrow(sobelmat),maxresm)
            resultm2[1:nrow(sobelmat),1:ncol(sobelmat)]<-sobelmat
            resultm<-rbind(resultm,resultm2)
            sobelmt2<-noquote(matrix(sprintf(decimals,sobelmat),nrow=nrow(sobelmat)))
            rownames(sobelmt2)<-sobelrlb
            colnames(sobelmt2)<-sobellab
            if (outscreen==1)
            {cat("\nNormal theory test for indirect effect(s):\n")
              print(sobelmt2,right=T)}
          }
          if (contrast != 0)
          {
            if (((contrast==1) | (contrast==2)) & (kk==efloop))
            {
              conkey<-matrix(conkey[2:nrow(conkey),],nrow=(nrow(conkey)-1))
              if (outscreen==1)
              {cat("\nSpecific indirect effect contrast definition(s):\n")
                write.table(conkey,quote=FALSE,row.names=contlbs,col.names=FALSE,sep = "  ")}
            }
            if ((contrast==3) & (kk=efloop))
            {
              crowlbs<-rowlbs[2:(nindfx+1),1]
              contvect<-noquote(matrix(sprintf(decimals,contvec),nrow=1))
              colnames(contvect)<-crowlbs
              rownames(contvect)<-"(C1)"
              if (outscreen==1)
              {cat("\nSpecific indirect effect contrast weights:\n")
                print(contvect,right=T)}
            }
            if ((contrast==2) & (kk==efloop) & (outscreen==1))
            {cat("\nContrasts are differences between absolute values of indirect effects\n")}
          }
          if ((serial==1) & (kk==efloop))
          {rowlbst<-matrix(rowlbs[2:nrow(rowlbs),1],nrow<-(nrow(rowlbs)-1))
          if (outscreen==1)
          {cat("\nIndirect effect key:\n")
            write.table(indkey,quote=FALSE,row.names=rowbls3,col.names=FALSE,sep=" ")}}
        }
        if (nxvls > 1)
        {
          if (outscreen==1)
          { 
            if (kk==1){cat("\nRelative indirect effects of X on Y:\n")}
            if (kk==2){cat("\nPartially standardized relative indirect effects of X on Y:\n")}
            if (kk==3){cat("\nCompletely standardized relative indirect effects of X on Y:\n")}
          }
          for (i in (1:nrow(indmake)))
          {
            indtabsm<-indtab[(((i-1)*nxvls)+1):(nxvls*i),]
            indkeyt<-indkey[i,]
            if (outscreen==1)
            {
              cat("\n")
              write.table(t(indkeyt),quote=FALSE,row.names=FALSE,col.names=FALSE,sep=" ")
              cat("\n")
            }
            if (bcmat[nrow(bcmat),1]==0)
            {direffl2<-xcatlab[1:nxvls,1]}
            resultm2<-matrix(99999,nrow(indtabsm),maxresm)
            resultm2[1:nrow(indtabsm),1:ncol(indtabsm)]<-indtabsm
            resultm<-rbind(resultm,resultm2)
            indtasm2<-noquote(matrix(sprintf(decimals,indtabsm),nrow=nrow(indtabsm)))
            colnames(indtasm2)<-bootlbs
            rownames(indtasm2)<-direffl2
            if (outscreen==1)
            {print(indtasm2,right=T)}
            if ((normal==1) & (sobelok==1) & (kk==1))
            {
              sobelsm=sobelmat[(((i-1)*nxvls)+1):(nxvls*i),]
              sobellab=c("Effect",hclab,"Z","p")
              resultm2<-matrix(99999,nrow(sobelsm),maxresm)
              resultm2[1:nrow(sobelsm),1:ncol(sobelsm)]<-sobelsm
              resultm<-rbind(resultm,resultm2)
              sobelsm2<-noquote(matrix(sprintf(decimals,sobelsm),nrow=nrow(sobelsm)))
              colnames(sobelsm2)<-sobellab
              rownames(sobelsm2)<-direffl2
              if (outscreen==1)
              {cat("\n   Normal theory test for relative indirect effects:\n")
                print(sobelsm2,right=T)}
            }
          }
        }
        if ((effsize==1) & (boot > 0)){bootres=bootres[2:nrow(bootres),]}
      }
      if (alttotal==1)
      {
        altcnms<-"Effect"
        totbtvec<-totbtvec+bootdir
        alttotfx<-t(totbtvec[1,])
        if (boot > 0)
        {
          alttotfx<-matrix(0,ncol(totbtvec),4)
          alttotfx[,1]<-t(totbtvec[1,])
          for (cec in (1:ncol(totbtvec)))
          {
            if (bc==0){bcitmp<-process.pboot3(totbtvec[2:nrow(totbtvec),cec],cilow,cihigh)}
            if (bc==1)
            {
              bcbout<-process.bcboot3(totbtvec[2:nrow(totbtvec),cec],totbtvec[1,cec],xp2,badend,priorlo,priorhi)
              bcitmp<-matrix(unlist(bcbout[1]))
              badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
            }
            alttotfx[cec,2:4]<-t(bcitmp)
          }
          altcnms<-c(altcnms,"BootSE","BootLLCI","BootULCI")
        }
        resultm2<-matrix(99999,nrow(alttotfx),maxresm)
        resultm2[1:nrow(alttotfx),1:ncol(alttotfx)]<-alttotfx
        resultm<-rbind(resultm,resultm2)
        if (outscreen==1)
        {
          if (nxvls > 1)
          {cat("----------\n")
            cat("\nRelative total effects of X on Y (sum of direct and indirect effects):\n")
            alttotfp<-noquote(matrix(sprintf(decimals,alttotfx),nrow=nrow(alttotfx)))
            colnames(alttotfp)<-altcnms
            rownames(alttotfp)<-direffl2
            print(alttotfp,right=T)
          }
          if (nxvls==1)
          {
            cat("\nTotal effects of X on Y (sum of direct and indirect effects):\n")
            alttotfp<-noquote(matrix(sprintf(decimals,alttotfx),nrow=nrow(alttotfx)))
            colnames(alttotfp)<-altcnms
            rownames(alttotfp)<-" "
            print(alttotfp,right=T)
          }
        }
      }

    }
    #this is the end of the no moderators loop */

    if (anymod > 0)
    {
      if (boot==0){bootres<-obscoeff;indtab<-999}
      if (boot > 0){bootres<-rbind(obscoeff,bootres);indtab<-matrix(999,1,4)}
      if (sum(as.numeric(indmod > 0))==nrow(indmod))
      {if (outscreen==1)
      {
        if (nxvls > 1)
        {cat("\nRelative conditional indirect effects of X on Y:\n")}
        if (nxvls==1)
        {cat("\nConditional indirect effects of X on Y:\n")}
      }
      }
      if (sum(as.numeric(indmod > 0)) < nrow(indmod))
      {if (outscreen==1)
      {
        if (nxvls > 1)
        {cat("\nRelative conditional and unconditional indirect effects of X on Y:\n")}
        if (nxvls==1)
        {cat("\nConditional and unconditional indirect effects of X on Y:\n")}
      }
      }
      for (i in (1:nrow(indmake)))
      {
        indtab<-matrix(0,1,4)
        indkeyt<-indkey[i,]
        if (outscreen==1)
        {
          cat("\nINDIRECT EFFECT:\n\n")
          write.table(t(indkeyt),quote=FALSE,row.names=FALSE,col.names=FALSE,sep=" ")
          cat("\n")
        }
        if (indmod[i,1]==0)
        {
          for (j in (1:nxvls))
          {
            indtemp<-matrix(1,nrow(bootres),1)
            for (k in (1:indmake[i,1]))
            {
              jtemp<-1
              if ((j > 1) & (k==1)){jtemp<-j}       
              indtemp<-indtemp*bootres[,pathsfoc[jtemp,indmake[i,(k+1)]]]
            }
            indeff<-indtemp[1,1]
            if (boot > 0)
            {
              if (bc==0){bcitmp3<-process.pboot3(indtemp[2:nrow(indtemp),1],cilow,cihigh)}
              if (bc==1)
              {
                bcbout<-process.bcboot3(indtemp[2:nrow(indtemp),1],indtemp[1,1],xp2,badend,priorlo,priorhi)
                bcitmp3<-matrix(unlist(bcbout[1]))
                badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
              }
              indeff<-cbind(indeff,t(bcitmp3))
            }  
            indtab<-rbind(indtab,indeff)
          }
          indtab<-matrix(indtab[2:nrow(indtab),],ncol=ncol(indtab))
          resultm2<-matrix(99999,nrow(indtab),maxresm)
          resultm2[1:nrow(indtab),1:ncol(indtab)]<-indtab
          resultm<-rbind(resultm,resultm2)
          indtab10<-noquote(matrix(sprintf(decimals,indtab),nrow=nrow(indtab)))
          if (nxvls > 1)
          {rownames(indtab10)<-xcatlab[1:nxvls,1]}
          if (nxvls==1)
          {rownames(indtab10)<-" "}
          colnames(indtab10)<-c("Effect","BootSE","BootLLCI","BootULCI")
          if (outscreen==1)
          {print(indtab10,right=T)}
        }
        # end of unmoderated
        # start of moderated
        if (indmod[i,1] > 0)
        {
          if (indmod[i,1]==1)
          {indmodva<-wmodvals;indprova<-wprobval;condlbs<-wnames;printw<-1}
          if (indmod[i,1]==2)
          {indmodva<-zmodvals;indprova<-zprobval;condlbs<-znames;printz<-1}
          if (indmod[i,1]==3)
          {
            cntmp<-1;printz<-1;printw<-1
            indmodva<-matrix(999,(nrow(wmodvals)*nrow(zmodvals)),2)
            for (k7 in (1:nrow(wmodvals)))
            {
              for (k8 in (1:nrow(zmodvals)))
              {indmodva[cntmp,]<-cbind(wmodvals[k7,1],zmodvals[k8,1]);cntmp<-cntmp+1}
            }
            condlbs<-cbind(wnames,znames)
          }
          condres<-matrix(999,nrow(indmodva),1)
          if (boot > 0)
          {condres=matrix(999,nrow(indmodva),4)}  
          condres<-cbind(indmodva,condres)

          # Here is where the computations start
          for (k4 in (1:nxvls))
          {
            imm3<-matrix(1,nrow(bootres),1);imm4<-matrix(1,nrow(bootres),1);indcontr<-0  
            if (indmod[i,1]==3){tihsw<-wprobval;tihsz<-zprobval}
            for (k1 in (1:nrow(indmodva)))
            {
              tucker2<-matrix(1,nrow(bootres),1)
              imm2<-matrix(1,nrow(bootres),1)
              wfirst<-0;zfirst<-0;immset<-0
              for (k2 in (1:indmake[i,1]))
              { 
                colnumb<-indmake[i,(k2+1)]
                if (k2==1)
                {
                  wbb<-matrix(0,nrow(bootres),(nwvls*nxvls))
                  zbb<-matrix(0,nrow(bootres),(nzvls*nxvls))
                  wzbb<-matrix(0,nrow(bootres),(nwvls*nzvls*nxvls))
                }
                if (k2 != 1)
                {
                  wbb<-matrix(0,nrow(bootres),nwvls)
                  zbb<-matrix(0,nrow(bootres),nzvls)
                  wzbb<-matrix(0,nrow(bootres),(nwvls*nzvls))
                }
                cnt<-1;tihs<-matrix(indlocs[2:((indlocs[1,colnumb])+1),colnumb])
                if (k2==1)
                {
                  focbb<-matrix(tihs[1:nxvls,1]);focbb<-matrix(bootres[,focbb],ncol=nxvls)
                  if (indmmm[i,1] > 0)
                  {imm<-matrix(focbb[,k4]);condbb<-matrix(imm,ncol=ncol(imm))}
                  focaddon<-matrix(0,1,nxvls);focaddon[1,k4]<-1;cnt<-cnt+nxvls;placeh<-nxvls
                  if (indmod[i,1]==1)
                  {
                    tihsz<-matrix(0,nrow(wprobval),(nzvls*nxvls))
                    tihswz<-matrix(0,nrow(wprobval),(nwvls*nzvls*nxvls))
                    if (pathsw[1,colnumb]==1)
                    {
                      temp<-matrix(0,nrow(wprobval),(nxvls*nwvls))
                      for (k5 in (1:nrow(wprobval)))
                      {
                        for (k6 in (1:nwvls))
                        {temp[k5, (((k4-1)*nwvls)+k6)]<-wprobval[k5,k6]}
                      }
                      indprova<-cbind(temp,tihsz,tihswz)
                    } else {
                      indprova<-cbind(wprobval,tihsz,tihswz)}
                  }
                  if (indmod[i,1]==2)
                  {
                    tihsw<-matrix(0,nrow(zprobval),(nwvls*nxvls))
                    tihswz<-matrix(0,nrow(zprobval),(nwvls*nzvls*nxvls))
                    if (pathsz[1,colnumb]==1)
                    {
                      temp<-matrix(0,nrow(zprobval),(nxvls*nzvls))
                      for (k5 in (1:nrow(zprobval)))
                      {
                        for (k6 in (1:nzvls))
                        {temp[k5,(((k4-1)*nzvls)+k6)]<-zprobval[k5,k6]}  
                      }
                      indprova<-cbind(tihsw,temp,tihswz)
                    } else {
                      indprova<-cbind(tihsw,zprobval,tihswz)}
                  }
                  if (indmod[i,1]==3)
                  {
                    indprova<-matrix(0,(nrow(wprobval)*nrow(zprobval)),((ncol(wprobval)*nxvls)+(ncol(zprobval)*nxvls)+(nwvls*nzvls*nxvls)))
                    cntemp<-1           
                    for (k7 in (1:nrow(wprobval)))
                    {
                      for (k8 in (1:nrow(zprobval)))
                      {
                        temp<-(wprobval[k7,]*focaddon[1,k4])
                        indprova[cntemp,(((k4-1)*nwvls)+1):(k4*(nwvls))]<-temp
                        temp<-zprobval[k8,]*focaddon[1,k4]
                        indprova[cntemp,((((k4-1)*nzvls)+1)+(nxvls*nwvls)):((((k4-1)*nzvls)+1)+(nxvls*nwvls)+(nzvls-1))]<-temp
                        cntemp<-cntemp+1
                      }
                    }
                    if (pathsz[1,colnumb]==0)
                    {
                      temp<-matrix(0,nrow(indprova),(ncol(zprobval)*nxvls))
                      indprova[,((ncol(wprobval)*nxvls)+1):((ncol(wprobval)+ncol(zprobval))*nxvls)]<-temp
                    }
                    if (pathsw[1,colnumb]==0)
                    {
                      temp<-matrix(0,nrow(indprova),(ncol(wprobval)*nxvls))
                      indprova[,1:(ncol(wprobval)*nxvls)]<-temp   
                    }
                    if (pathswz[1,colnumb]==1)
                    {
                      cntemp<-(ncol(wprobval)*nxvls)+(ncol(zprobval)*nxvls)+((k4-1)*ncol(wprobval)*ncol(zprobval))+1
                      for (k7 in (1:ncol(wprobval)))
                      {
                        for (k8 in (1:ncol(zprobval)))
                        {
                          indprova[,cntemp]<-matrix(indprova[,((ncol(wprobval)*(k4-1))+k7)])*matrix(indprova[,((((k4-1)*ncol(zprobval))+k8)+(nxvls*ncol(wprobval)))])
                          cntemp<-cntemp+1
                        }
                      }
                    }
                  }
                }
                if (k2 > 1)
                {
                  focbb<-tihs[1,1];focbb=matrix(bootres[,focbb])
                  if (indmmm[i,1] > 0){imm<-matrix(focbb[,1]);condbb<-matrix(imm)}     
                  focaddon<-matrix(1);cnt<-cnt+1;placeh<-1
                  if (indmod[i,1]==1)
                  {
                    tihsz<-matrix(0,nrow(wprobval),nzvls)
                    tihswz<-matrix(0,nrow(wprobval),(nwvls*nzvls))
                    indprova<-cbind(wprobval,tihsz,tihswz)
                  }
                  if (indmod[i,1]==2)
                  {
                    tihsw<-matrix(0,nrow(zprobval),nwvls)
                    tihswz<-matrix(0,nrow(zprobval),(nwvls*nzvls))
                    indprova<-cbind(tihsw,zprobval,tihswz)
                  }
                  if (indmod[i,1]==3)
                  {
                    indprova<-matrix(0,(nrow(wprobval)*nrow(zprobval)),((ncol(wprobval)+ncol(zprobval))+(nwvls*nzvls)))
                    cntemp<-1           
                    for (k7 in (1:nrow(wprobval)))
                    {
                      for (k8 in (1:nrow(zprobval)))
                      {
                        indprova[cntemp,1:(ncol(wprobval)+ncol(zprobval))]<-cbind(t(matrix(wprobval[k7,])),t(matrix(zprobval[k8,])))
                        cntemp<-cntemp+1
                      }
                    }
                    if (pathsz[1,colnumb]==0)
                    {
                      temp<-matrix(0,nrow(indprova),ncol(zprobval))
                      indprova[,(ncol(wprobval)+1):(ncol(wprobval)+ncol(zprobval))]<-temp
                    }
                    if (pathsw[1,colnumb]==0)
                    {
                      temp<-matrix(0,nrow(indprova),ncol(wprobval))
                      indprova[,1:ncol(wprobval)]<-temp   
                    }
                    if (pathswz[1,colnumb]==1)
                    {
                      cntemp<-(ncol(wprobval)+ncol(zprobval)+1)
                      for (k7 in (1:ncol(wprobval)))
                      {
                        for (k8 in (1:ncol(zprobval)))
                        {
                          indprova[,cntemp]<-matrix(indprova[,k7])*matrix(indprova[,(ncol(wprobval)+k8)])
                          cntemp<-cntemp+1
                        }
                      }
                    }
                  }
                }  
                if (pathsw[1,colnumb]==1)
                {
                  wbb<-matrix(tihs[cnt:(cnt+(placeh*nwvls)-1),1])
                  wbb<-matrix(bootres[,wbb],ncol=nrow(wbb))
                  immlbs2<-matrix(wcatlab[1:nwvls,1])
                  if (zfirst==0){wfirst<-1}              
                  if ((indmmm[i,1]==1) | (indmmm[i,1]==31) | (indmmm[i,1]==51))
                  {
                    imm<-matrix(wbb[,1])
                    for (k7 in (1:nwvls))
                    {imm<-cbind(imm,wbb[,(((k4-1)*nwvls*(as.numeric(k2==1)))+k7)])}
                    imm<-matrix(imm[,2:ncol(imm)],ncol=(ncol(imm)-1))               
                  }
                  if ((indmmm[i,1]==41) | (indmmm[i,1]==51))
                  {
                    condbb<-matrix(0,nrow(bootres),1)
                    for (k7 in (1:nwvls))        
                    {condbb<-cbind(condbb,wbb[,(((k4-1)*nwvls*(as.numeric(k2==1)))+k7)])}          
                    condbb<-matrix(condbb[,2:ncol(condbb)],ncol=(ncol(condbb)-1))
                  }    
                  cnt<-cnt+(placeh*nwvls)
                }
                if (pathsz[1,colnumb]==1)
                {
                  zbb<-matrix(tihs[cnt:(cnt+(placeh*nzvls)-1),1])
                  zbb<-matrix(bootres[,zbb],ncol=nrow(zbb))
                  if (wfirst==0){zfirst<-1}
                  if (indmmm[i,1] != 31){immlbs2<-matrix(zcatlab[1:nzvls,1])}
                  if ((indmmm[i,1]==2) | (indmmm[i,1]==31) | (indmmm[i,1]==51))
                  {
                    if (indmmm[i,1]==2){imm<-matrix(zbb[,1])}
                    for (k7 in (1:nzvls))
                    {imm<-cbind(imm,zbb[,(((k4-1)*nzvls*(as.numeric(k2==1)))+k7)])}
                    if ((indmmm[i,1]==2) | (indmmm[i,1]==51))
                    {
                      imm<-matrix(imm[,2:ncol(imm)],ncol=(ncol(imm)-1))  
                      if (indmmm[i,1]==51){condbb<-cbind(condbb,imm)}
                    }         
                  }
                  cnt<-cnt+(placeh*nzvls)
                }
                if (pathswz[1,colnumb]==1)
                {
                  wzbb<-matrix(tihs[cnt:(cnt+(placeh*nwvls*nzvls)-1),1])
                  wzbb<-matrix(bootres[,wzbb],ncol=nrow(wzbb))
                  if (indmmm[i,1]==41)
                  {
                    imm<-matrix(wzbb[,1])              
                    for (k7 in (1:(nwvls*nzvls)))
                    {imm<-cbind(imm,wzbb[,(((k4-1)*nzvls*nwvls*(as.numeric(k2==1)))+k7)])}
                  }
                  if (indmmm[i,1]==41)
                  {
                    imm<-matrix(imm[,2:ncol(imm)],ncol=(ncol(imm)-1))
                    condbb<-cbind(condbb,imm[,(ncol(imm)-(nwvls*nzvls)+1):ncol(imm)]) 
                  }
                  cnt<-cnt+(placeh*nzvls*nwvls)
                }
                indprobe<-cbind(focaddon,t(matrix(indprova[k1,])))
                tucker<-cbind(focbb,wbb,zbb,wzbb)
                for (k3 in (1:ncol(indprobe)))
                {tucker[,k3]<-matrix(tucker[,k3]*indprobe[1,k3])}
                tucker2<-tucker2*matrix(rowSums(tucker))
                if ((indmmm[i,1]==1) | (indmmm[i,1]==2) | (indmmm[i,1]==31) | (indmmm[i,1]==41) | (indmmm[i,1]==51))
                {
                  if (immset==1)
                  {
                    if ((ncol(imm2)==1) & (ncol(imm)==1)){imm2<-(matrix(imm2)*matrix(imm))}
                    if ((indmmm[i,1]==41) | (indmmm[i,1]==51))
                    {
                      if ((ncol(condbb2) > 1) & (ncol(condbb) > 1))
                      {
                        condbb2t<-matrix(-999999,nrow(condbb2),(ncol(condbb2)*ncol(condbb)))
                        k9<-1
                        if (wfirst==1)
                        {
                          for (k7 in (1:ncol(condbb2)))
                          {
                            for (k8 in (1:ncol(condbb)))
                            {condbb2t[,k9]<-(matrix(condbb2[,k7])*matrix(condbb[,k8]));k9<-k9+1}
                          }
                        }
                        if (zfirst==1)
                        {
                          for (k7 in (1:ncol(condbb)))
                          {
                            for (k8 in (1:ncol(condbb2)))
                            {condbb2t[,k9]<-(matrix(condbb[,k7])*matrix(condbb2[,k8]));k9<-k9+1}             
                          }
                        }
                        condbb2<-matrix(condbb2t,ncol=ncol(condbb2t))            
                      }
                      if ((ncol(condbb2) > 1) & (ncol(condbb)==1))
                      {
                        for (k7 in (1:ncol(condbb2)))
                        {condbb2[,k7]<-(matrix(condbb2[,k7])*matrix(condbb))}
                      }
                      if ((ncol(condbb2)==1) & (ncol(condbb) > 1))
                      {
                        for (k7 in (1:ncol(condbb)))
                        {condbb[,k7]<-(matrix(condbb2)*matrix(condbb[,k7]))}
                        condbb2<-matrix(condbb,ncol=ncol(condbb))
                      }                 
                    }
                    if ((ncol(imm2) != 1) & (ncol(imm) != 1))
                    {
                      imm2t<-matrix(-999999,nrow(imm2),(ncol(imm2)*ncol(imm)))
                      k9<-1
                      if (wfirst==1)
                      {
                        for (k7 in (1:ncol(imm2)))
                        {
                          for (k8 in (1:ncol(imm)))
                          {imm2t[,k9]<-(matrix(imm2[,k7])*matrix(imm[,k8]));k9<-k9+1}
                        }
                      }
                      if (zfirst==1)
                      {
                        for (k7 in (1:ncol(imm)))
                        {
                          for (k8 in (1:ncol(imm2)))
                          {imm2t[,k9]<-(matrix(imm[,k7])*matrix(imm2[,k8]));k9<-k9+1}
                        }
                      }
                      imm2<-matrix(imm2t,ncol=ncol(imm2t))
                    }
                    if ((ncol(imm2) > 1) & (ncol(imm)==1))
                    {
                      for (k7 in (1:ncol(imm2)))
                      {imm2[,k7]<-(matrix(imm2[,k7])*matrix(imm))}
                    }
                    if ((ncol(imm2)==1) & (ncol(imm) > 1))
                    {
                      for (k7 in (1:ncol(imm)))
                      {imm[,k7]<-(matrix(imm2)*matrix(imm[,k7]))}
                      imm2<-matrix(imm,ncol=ncol(imm))
                    }
                  }
                  if (immset==0)
                  {
                    imm2<-matrix(imm,ncol=ncol(imm))
                    if ((indmmm[i,1]==41) | (indmmm[i,1]==51))
                    {condbb2<-matrix(condbb,ncol=ncol(condbb))}
                    immset<-1
                  }
                }
                # that is it for mmm loop */
              }
              # end of looping through paths: k2
              indtemp<-tucker2[1,1]
              if ((indmmm[i,1]==12) | (indmmm[i,1]==22))
              {
                imm3<-cbind(imm3,tucker2)
                if (k1==nrow(indmodva))
                {
                  imm3<-matrix(imm3[,2:ncol(imm3)],ncol=(ncol(imm3)-1))
                  immstop<-ncol(imm3)
                  for (k8 in (2:immstop))
                  {
                    if (indmmmt[i,1]==1){imm3<-cbind(imm3,(imm3[,k8]-imm3[,1]))}
                    if (indmmmt[i,1]==2){imm3=cbind(imm3,(imm3[,k8]-imm3[,(k8-1)]))}
                    if (indmmmt[i,1]==3)
                    {
                      imm3<-cbind(imm3,((rowSums(imm3[,(k8:immstop)])/(immstop-k8+1))-imm3[,(k8-1)]))
                    }
                    if (indmmmt[i,1]==4)
                    {imm3<-cbind(imm3,(imm3[,k8]-(rowSums(imm3[,1:immstop])/immstop)))}
                  }
                  if (indmmmt[i,1] < 5)
                  {imm2<-matrix(imm3[,(immstop+1):ncol(imm3)],ncol=(ncol(imm3)-immstop))}
                }
              }      
              if ((indmmm[i,1]>-1) & ((contrast==1) | (contrast==2)))
              {
                imm4<-cbind(imm4,tucker2)
                if ((k1==nrow(indmodva)) & (k1 > 1))
                {
                  imm4<-matrix(imm4[,2:ncol(imm4)],ncol=(ncol(imm4)-1))
                  immstop<-ncol(imm4)
                  condcont<-matrix(-999,(immstop*(immstop-1)/2),6)              
                  for (k8 in (1:(immstop-1)))
                  {
                    for (k9 in ((k8+1):immstop))
                    {                   
                      if (contrast==1){imm4<-cbind(imm4,matrix(imm4[,k9]-imm4[,k8]))}
                      if (contrast==2){imm4<-cbind(imm4,matrix(abs(imm4[,k9])-abs(imm4[,k8])))}
                      condcont[(ncol(imm4)-immstop),1]<-imm4[1,k9]
                      condcont[(ncol(imm4)-immstop),2]<-imm4[1,k8]
                    }
                  }
                  imm4<-matrix(imm4[,(immstop+1):ncol(imm4)],ncol=(ncol(imm4)-immstop))   
                  for (k8 in (1:ncol(imm4)))
                  {
                    condcont[k8,3]<-imm4[1,k8]
                    if (boot > 0)
                    {
                      if (bc==0){condcon3<-process.pboot3(imm4[2:nrow(imm4),k8],cilow,cihigh)}
                      if (bc==1)
                      {
                        bcbout<-process.bcboot3(imm4[2:nrow(imm4),k8],imm4[1,k8],xp2,badend,priorlo,priorhi)
                        condcon3<-matrix(unlist(bcbout[1]))
                        badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
                      }
                      condcont[k8,4:6]<-t(condcon3)  
                    }
                  }
                  if (boot==0){condcont<-matrix(condcont[,1:3],ncol=3)}
                  indcontr<-1
                }
              }
              if (boot > 0)
              {
                if (bc==0){bcitmp11<-process.pboot3(tucker2[2:nrow(tucker2),1],cilow,cihigh)}
                if (bc==1)
                {
                  bcbout<-process.bcboot3(tucker2[2:nrow(tucker2),1],tucker2[1,1],xp2,badend,priorlo,priorhi)
                  bcitmp11<-matrix(unlist(bcbout[1]))
                  badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
                }
                indtemp<-cbind(indtemp,t(bcitmp11))  
              }
              condres[k1,(ncol(indmodva)+1):ncol(condres)]<-indtemp
            }
            # end of looping through indirect effects: k1
            # Here is where the computations end
            resultm2<-matrix(99999,nrow(condres),maxresm)
            resultm2[1:nrow(condres),1:ncol(condres)]<-condres
            resultm<-rbind(resultm,resultm2)
            condresp<-noquote(matrix(sprintf(decimals,condres),nrow=nrow(condres)))
            if (k4==1)
            {
              condlbs<-cbind(condlbs,"Effect")
              if (boot > 0)
              {condlbs<-cbind(condlbs,"BootSE","BootLLCI","BootULCI")}
              condrlb<-t(matrix(replicate(nrow(condres)," ")))
            }
            if (nxvls > 1)
            {condrlb<-matrix(xcatlab[k4,1],nrow(condres),1)}
            colnames(condresp)<-condlbs
            rownames(condresp)<-condrlb
            if (outscreen==1)
            {print(condresp,right=T)}
            if (indmmm[i,1]==0){cat("\n")}
            dichadj<-0;immcat<-0
            if (indmmm[i,1] > 0)
            {
              if ((indmmm[i,1]==1) | (indmmm[i,1]==12) | (indmmm[i,1]==31))
              {
                if ((wdich==1) & (mcw==0))
                {
                  if (indmmm[i,1] != 12){imm2[,1]<-imm2[,1]*(wmax-wmin)}
                  if (indmmm[i,1] != 31){dichadj<-1}
                }           
                if (((mcw==1) | (mcw==2)) & (indmmm[i,1] != 31)){immcat<-1}
              }
              if ((indmmm[i,1]==2) | (indmmm[i,1]==22) | (indmmm[i,1]==31))
              {
                if ((zdich==1) & (mcz==0))
                {
                  if (indmmm[i,1]==31)
                  {
                    imm2[,(nwvls+1):ncol(imm2)]<-matrix(imm2[,(nwvls+1):ncol(imm2)],ncol=(ncol(imm2)-nwvls))*(zmax-zmin)
                  }
                  if (indmmm[i,1]==2){imm2[,1]<-imm2[,1]*(zmax-zmin)}
                  if (indmmm[i,1] != 31){dichadj<-1}
                }
                if (((mcz==1) | (mcz==2)) & (indmmm[i,1] != 31)){immcat<-1}
              }
              immtemp2<-t(matrix(imm2[1,],ncol=ncol(imm2)))
              immtemp<-matrix(immtemp2,ncol=ncol(immtemp2))
              immlbs<-"Index"
              if (boot > 0)
              {
                immtemp<-matrix(0,1,3)
                for (k7 in (1:ncol(imm2)))
                {
                  if (bc==0){immbtci<-process.pboot3(imm2[2:nrow(imm2),k7],cilow,cihigh)} 
                  if (bc==1)
                  {
                    bcbout<-process.bcboot3(imm2[2:nrow(imm2),k7],imm2[1,k7],xp2,badend,priorlo,priorhi)
                    immbtci<-matrix(unlist(bcbout[1]))
                    badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
                  }
                  immtemp<-rbind(immtemp,t(immbtci))
                } 
                immtemp<-matrix(immtemp[2:nrow(immtemp),],ncol=ncol(immtemp))
                immtemp<-cbind(immtemp2,immtemp)
                immlbs<-c(immlbs,"BootSE","BootLLCI","BootULCI")
              }
              resultm2<-matrix(99999,nrow(immtemp),maxresm)
              resultm2[1:nrow(immtemp),1:ncol(immtemp)]<-immtemp
              resultm<-rbind(resultm,resultm2)
              immtempr<-noquote(matrix(sprintf(decimals,immtemp),nrow=nrow(immtemp)))
              colnames(immtempr)<-immlbs        
              if ((dichadj==0) & (immcat==0) & (indmmmt[i,1] != 5) & (indmmm[i,1] < 100))
              {
                if (indmmm[i,1] < 30)
                {
                  rownames(immtempr)<-immlbs2
                  if (outscreen==1)
                  {cat("\n     Index of moderated mediation:\n")
                    print(immtempr,right=T)}
                  if ((nxvls > 1) & (k4 < nxvls)){cat("\n")}
                }
                if (indmmm[i,1]==31)
                {
                  immlbs2<-rbind(immlbs2,matrix(zcatlab[1:nzvls,1]))
                  rownames(immtempr)<-immlbs2
                  if (outscreen==1)
                  {cat("\n     Indices of partial moderated mediation:\n")
                    print(immtempr,right=T)}
                  if ((nxvls > 1) & (k4 < nxvls)){cat("\n")}
                }
                if ((nzvls==1) & (nwvls==1))
                {
                  if ((indmmm[i,1]==41) | (indmmm[i,1]==51))
                  {
                    for (k7 in (1:nwvls))
                    {
                      immlbs2<-zcatlab[1:nzvls,1]
                      immtemp2<-matrix(immtemp[(((k7-1)*nzvls)+1):(((k7-1)*nzvls)+nzvls),],ncol=ncol(immtemp))
                      #resultm2<-matrix(99999,nrow(immtemp2),maxresm)
                      #resultm2[1:nrow(immtemp2),1:ncol(immtemp2)]<-immtemp2
                      #print(resultm)
                      #print(resultm2)
                      #resultm<-rbind(resultm,resultm2)
                      #print(resultm)     
                      immtem2r<-noquote(matrix(sprintf(decimals,immtemp2),nrow=nrow(immtemp2)))
                      colnames(immtem2r)<-immlbs
                      if (nwvls > 1)
                      {
                        primodv<-cbind("Primary moderator:", wcatlab[k7,1])
                        if (outscreen==1)
                        {write.table(primodv,quote=FALSE,row.names=FALSE,col.names=FALSE)}
                      }
                      if (nzvls==1)
                      {
                        rownames(immtem2r)<-" "
                        if (outscreen==1)
                        {cat("\n     Index of moderated moderated mediation:\n")
                          print(immtem2r,right=T)}
                      } else {
                        rownames(immtem2r)<-immlbs2
                        if (outscreen==1)            
                        {cat("\n     Indices of moderated moderated mediation:\n")
                          print(immtem2r,right=T)}
                      }
                      cmmtemp<-matrix(0,nrow(zprobval),4)        
                      for (k8 in (1:nrow(zprobval)))
                      {
                        bbst<-((nwvls+1)+((k7-1)*nzvls));bben<-((nwvls+1)+((k7-1)*nzvls)+(nzvls-1))
                        condbb3<-matrix(condbb2[,bbst:bben],ncol=(bben-bbst+1))
                        if (ncol(zprobval) > 1)
                        {
                          condbb3<-condbb3*diag(c(zprobval[k8,]))
                        } else {
                          condbb3<-condbb3*zprobval[k8,]
                        }
                        condbb3<-cbind(condbb2[,k7],condbb3)
                        icmm<-matrix(rowSums(condbb3))
                        cmmtemp[k8,1]<-icmm[1,1]
                        if (boot > 0)
                        {
                          if (bc==0){cmmt3<-process.pboot3(icmm[2:nrow(icmm),1],cilow,cihigh)}
                          if (bc==1)
                          {
                            bcbout<-process.bcboot3(icmm[2:nrow(icmm),1],icmm[1,1],xp2,badend,priorlo,priorhi)
                            cmmt3<-matrix(unlist(bcbout[1]))
                            badend<-unlist(bcbout[2]);priorlo<-unlist(bcbout[3]);priorhi<-unlist(bcbout[4])
                          }
                          cmmtemp[k8,2:4]<-t(cmmt3)
                        }
                        cmmlbs<-cbind(znames,t(matrix(immlbs)))
                      }
                      if (boot==0){cmmtemp<-matrix(cmmtemp[,1:1],ncol=1);cmmlbs<-cbind(znames,"Index")}                   
                      cmmtemp<-cbind(zmodvals,cmmtemp)
                      resultm2<-matrix(99999,nrow(cmmtemp),maxresm)
                      resultm2[1:nrow(cmmtemp),1:ncol(cmmtemp)]<-cmmtemp
                      resultm<-rbind(resultm,resultm2)
                      cmmtempr<-noquote(matrix(sprintf(decimals,cmmtemp),nrow=nrow(cmmtemp)))
                      #if (boot==0){cmmtempr<-matrix(cmmtemp[,1:2],ncol=2)}                   
                      #cmmlbs<-cbind(znames,t(matrix(immlbs)))
                      rownames(cmmtempr)<-t(matrix(replicate(nrow(cmmtempr)," ")))
                      colnames(cmmtempr)<-cmmlbs
                      if (outscreen==1)
                      {cat("\n     Indices of conditional moderated mediation by W:\n")
                        print(cmmtempr,right=T)}
                    }
                  }
                }
              }
              if (((dichadj==1) | (immcat==1)) & (indmmm[i,1] < 30))
              {
                if (outscreen==1)
                {
                  cat("\n     Index of moderated mediation\n")
                  cat("     (differences beween conditional indirect effects):\n")
                  rownames(immtempr)<-immlbs2
                  print(immtempr,right=T)
                }
              }
            }
            if (indcontr==1)
            {
              resultm2<-matrix(99999,nrow(condcont),maxresm)
              resultm2[1:nrow(condcont),1:ncol(condcont)]<-condcont
              resultm<-rbind(resultm,resultm2)
              condcnt3<-noquote(matrix(sprintf(decimals,condcont),nrow=nrow(condcont))) 
              condctlb<-c("Effect1","Effect2","Contrast","BootSE", "BootLLCI","BootULCI")
              colnames(condcnt3)<-condctlb
              rownames(condcnt3)<-t(matrix(replicate(nrow(condcnt3)," ")))
              if (outscreen==1)
              {
                cat("\n     Pairwise contrasts between conditional indirect effects\n")
                cat("     (Effect1 minus Effect2):\n")
                print(condcnt3,right=T)
              }
            }
            if (outscreen==1)
            {cat("\n---\n")}
          }
        }
        #end of moderated */
      }
    }
    # This is the end of the moderated loop */
  }

  if ((criterr==0) & (saveboot==1))
  {
    if (boot > 0)
    {
      if (outscreen==1)
      {
        cat("\n*********************************************************************** \n")
        conseq<-conseq[2:nrow(conseq),]
        #cat("Bootstrap estimates were saved in a global dataframe named process.boots\n")
        colslab<-cbind(savlabs,conseq,vlabs)
        cat("\nMap of column names to model coefficients in bootstrap matrix:\n")
        rownames(colslab)<-t(matrix(replicate(nrow(colslab)," ")))
        colnames(colslab)<-c("name","Conseqnt","Antecdnt")
        print(noquote(colslab),right=T)
      }
    }
  }

  #PRINT BOOTSTRAP RESULTS FOR MODEL PARAMETERS
  if ((criterr==0) & (boot > 0) & (modelbt==1))
  {
    labstart<-1
    if (outscreen==1)
    {cat("\n********** BOOTSTRAP RESULTS FOR REGRESSION MODEL PARAMETERS **********\n\n")}
    for (iboot in (1:(nms+nys)))
    {
      vlabsm<-matrix(vlabs[labstart:(labstart+(nump[1,iboot]-1)),1])
      bootcimt<-matrix(bootcim[labstart:(labstart+(nump[1,iboot]-1)),],ncol=5)
      resultm2<-matrix(99999,nrow(bootcimt),maxresm)
      resultm2[1:nrow(bootcimt),1:ncol(bootcimt)]<-bootcimt
      resultm<-rbind(resultm,resultm2)
      bootcimt<-noquote(matrix(sprintf(decimals,bootcimt),ncol=5))
      colnames(bootcimt)<-matrix(c("Coeff","BootMean","BootSE","BootLLCI","BootULCI"))
      rownames(bootcimt)<-vlabsm
      if (outscreen==1)
      {
        cat("Outcome variable: ")
        write.table(outnames[iboot,1],quote=FALSE,row.names=FALSE,col.names=FALSE)
        cat("\n")
        print(bootcimt,right=T)
      }
      labstart<-labstart+nump[1,iboot]
      if ((iboot < (nms+nys)) & (outscreen==1)){cat("----------\n")}
    }
  }

  #print model matrices
  if ((criterr==0) & (matrices==1) & (outscreen==1))
  {
    cat("\n********************* MODEL DEFINITION MATRICES *********************** \n \n")
    cat("FROM variables are columns, TO variables are rows. \n \n")
    temp2<-matrix("0",nrow(bcmat),ncol(bcmat))
    for (i in c(2:nrow(bcmat)))
    {for (j in c(1:(ncol(bcmat)-1)))
    {if (bcmat[i,j]==1){temp2[i,j]<-"1"}
      if (j >= i){temp2[i,j]<-" "}}
    }
    temp2<-as.matrix(noquote(temp2[2:nrow(bcmat),(1:(ncol(bcmat)-1))]))
    if (nms > 0){cmatlabs<-c(xnames,mnames);rmatlabs<-c(mnames,ynames)}
    if (nms==0){cmatlabs<-xnames;rmatlabs<-ynames}
    colnames(temp2)<-cmatlabs
    rownames(temp2)<-rmatlabs
    cat("BMATRIX: Paths freely estimated (1) and fixed to zero (0): \n") 
    print(temp2)
    z<-0
    if (sum(wcmat) !=0)
    {
      temp2<-matrix("0",nrow(wcmat),ncol(wcmat))
      for (i in c(2:nrow(wcmat)))
      {for (j in c(1:(ncol(wcmat)-1)))
      {if (wcmat[i,j]==1){temp2[i,j]<-"1"}
        if (j >= i){temp2[i,j]<-" "}}
      }
      temp2<-as.matrix(noquote(temp2[2:nrow(wcmat),(1:(ncol(wcmat)-1))]))
      colnames(temp2)<-cmatlabs
      rownames(temp2)<-rmatlabs
      cat("\nWMATRIX: Paths moderated (1) and not moderated (0) by W: \n") 
      print(temp2)
    }
    if (sum(zcmat) !=0)
    {
      temp2<-matrix("0",nrow(zcmat),ncol(zcmat))
      for (i in c(2:nrow(zcmat)))
      {for (j in c(1:(ncol(zcmat)-1)))
      {if (zcmat[i,j]==1){temp2[i,j]<-"1"}
        if (j >= i){temp2[i,j]<-" "}}
      }
      temp2<-as.matrix(noquote(temp2[2:nrow(zcmat),(1:(ncol(zcmat)-1))]))
      colnames(temp2)<-cmatlabs
      rownames(temp2)<-rmatlabs
      cat("\nZMATRIX: Paths moderated (1) and not moderated (0) by Z: \n") 
      print(temp2)
    }
    if (sum(wzcmat) !=0)
    {
      temp2<-matrix("0",nrow(wzcmat),ncol(wzcmat))
      for (i in c(2:nrow(wzcmat)))
      {for (j in c(1:(ncol(wzcmat)-1)))
      {if (wzcmat[i,j]==1){temp2[i,j]<-"1"}
        if (j >= i){temp2[i,j]<-" "}}
      }
      temp2<-as.matrix(noquote(temp2[2:nrow(wzcmat),(1:(ncol(wzcmat)-1))]))
      colnames(temp2)<-cmatlabs
      rownames(temp2)<-rmatlabs
      cat("\nWZMATRIX: W moderated paths moderated (1) and not moderated (0) by Z: \n") 
      print(temp2)
    }
    if (ncs > 0)
    {colnames(ccmat)<-covnames
    rownames(ccmat)<-rmatlabs
    cat("\nCMATRIX: Covariates (columns) in (1) and not in (0) the models of M and Y (rows): \n")
    print(ccmat)
    }  
  }
  #end print model matrices

  if (criterr==0)
  {
    resultm<-resultm[2:nrow(resultm),]
    bocaj<-matrix(as.numeric(resultm==99999),nrow=nrow(resultm))
    bocaj<-(colSums(bocaj)==nrow(resultm))
    bocaj<-matrix(1-as.numeric(bocaj))
    j<-1
    for (i in (1:ncol(resultm)))
    {if (bocaj[i,1]==1){j=j+1}}
    resultm<-resultm[,1:(j-1)]
    resultm[resultm==99999]<-NA
  }
  if ((outscreen==1) & (activate==0))
  {cat("\n******************** ANALYSIS NOTES AND ERRORS ************************ \n")}
  if (activate==1)
  {cat("*********************************************************************** \n")}
  if ((criterr==0) & (outscreen==1))
  {
    cat("\nLevel of confidence for all confidence intervals in output: ")
    write.table(conf,quote=FALSE,row.names=FALSE,col.names=FALSE)
    if (boot>0)
    { 
      if ((goodboot==boot) & (bc==0))
      {
        cat("\nNumber of bootstraps for percentile bootstrap confidence intervals: ")
        write.table(boot,quote=FALSE,row.names=FALSE,col.names=FALSE)  
      }
      if ((goodboot==boot) & (bc==1))
      {
        cat("\nNumber of bootstraps for bias-corrected bootstrap confidence intervals: ")
        write.table(boot,quote=FALSE,row.names=FALSE,col.names=FALSE)  
      }
      if (length(badend) > 1)
      {
        cat("\nWARNING: Bootstrap CI endpoints below are not trustworthy. Decrease confidence\n")
        cat("or increase the number of bootstrap samples.\n")
        badend<-matrix(badend);badend<-t(badend[2:nrow(badend),1])
        badend<-noquote(matrix(sprintf(decimals,badend),nrow=nrow(badend)))
        write.table(badend,quote=FALSE,row.names=FALSE,col.names=FALSE)
      }
    }
    if (mc > 0)
    {
      cat("\nNumber of samples for Monte Carlo confidence intervals: ")
      write.table(mc,quote=FALSE,row.names=FALSE,col.names=FALSE)  
    }
    if ((wnotev > 0) & (printw==1))
    {
      if (wnotev==1)
      {cat("\nW values in conditional tables are the 16th, 50th, and 84th percentiles.\n")}
      if ((wnotev==2) & (minwwarn==0) & (maxwwarn==0))
      {cat("\nW values in conditional tables are the mean and +/- SD from the mean.\n")}  
      if (minwwarn==1)
      {cat("\nW values in conditional tables are the minimum, the mean, and 1 SD above the mean.\n")} 
      if (maxwwarn==1)
      {cat("\nW values in conditional tables are 1 SD below the mean, the mean, and the maximum.\n")}
    }
    if ((znotev > 0) & (printz==1))
    {
      if (znotev==1)
      {cat("\nZ values in conditional tables are the 16th, 50th, and 84th percentiles.\n")}
      if ((znotev==2) & (minzwarn==0) & (maxzwarn==0))
      {cat("\nZ values in conditional tables are the mean and +/- SD from the mean.\n")}  
      if (minzwarn==1)
      {cat("\nZ values in conditional tables are the minimum, the mean, and 1 SD above the mean.\n")} 
      if (maxzwarn==1)
      {cat("\nZ values in conditional tables are 1 SD below the mean, the mean, and the maximum.\n")}
    }
    if (minwwarn > 0)
    {cat("\nNOTE: One SD below the mean is below the minimum observed in the data for W,\n")
      cat("      so the minimum measurement on W is used for conditioning instead.\n")}
    if (maxwwarn > 0)
    {cat("\nNOTE: One SD above the mean is above the maximum observed in the data for W,\n")
      cat("      so the maximum measurement on W is used for conditioning instead.\n")}
    if (minzwarn > 0)
    {cat("\nNOTE: One SD below the mean is below the minimum observed in the data for Z,\n")
      cat("      so the minimum measurement on Z is used for conditioning instead.\n")}
    if (maxzwarn > 0)
    {cat("\nNOTE: One SD above the mean is above the maximum observed in the data for Z,\n")
      cat("      so the maximum measurement on Z is used for conditioning instead.\n")}
    if (pstog==1)
    {cat("\nNOTE: Standardized coefficients for dichotomous or multicategorical X are\n")
      cat("      in partially standardized form.\n")}
    for (i in c(1:100)) 
    {
      if (notecode[i,1]==1)
      {cat(" \n")
        cat("NOTE: COVMY is ignored when using CMATRIX option. \n")}
      if (notecode[i,1]==2)
      {cat(" \n")
        cat("NOTE: Confidence level restricted to between 50 and 99.9999%. 95% confidence is provided in output. \n")}
      if (notecode[i,1]==3)
      {cat(" \n")
        cat("NOTE: The following variables were mean centered prior to analysis: \n")
        centvar<-c("      ", centvar)
        centvar<-t(noquote(centvar))  
        write.table(centvar,quote=FALSE,row.names=FALSE,col.names=FALSE)
      }
      if ((notecode[i,1]==4) & (ydich != 1))
      {cat(" \n")
        cat("NOTE: A heteroscedasticity consistent standard error and covariance matrix estimator was used. \n")}  
      if (notecode[i,1]==6)
      {cat(" \n")
        cat("NOTE: Due to estimation problems, some bootstrap samples had to be replaced. \n")
        cat("      The number of times this happened was: ")
        write.table(badboot,quote=FALSE,row.names=FALSE,col.names=FALSE)}  
      if (notecode[i,1]==7)
      {cat(" \n")
        cat("NOTE: The bootstrapping was not completed due to problematic bootstrap samples. \n")
        cat("      Bootstrap confidence intervals are therefore suppressed.\n")}
      if (notecode[i,1]==8)
      {cat(" \n")
        cat("NOTE: The number of bootstrap samples was adjusted upward given your desired confidence. \n")}  
      if (notecode[i,1]==9)
      {cat(" \n")
        cat("NOTE: WMODVAL is ignored when W is specified as multicategorical. \n")}  
      if (notecode[i,1]==10)
      {cat(" \n")
        cat("NOTE: ZMODVAL is ignored when Z is specified as multicategorical. \n")}
      if (notecode[i,1]==11)
      {cat(" \n")
        cat("NOTE: Total effect model generated only when all covariates are specified\n")
        cat("      in all models of M and Y.\n")}
      if (notecode[i,1]==12)
      {cat(" \n")
        cat("NOTE: Total effect model generated only when X is freely estimated to\n")
        cat("      affect each M and both X and M are freely estimated to affect Y.\n")}
      if (notecode[i,1]==30)
      {cat(" \n")
        cat("NOTE: Your vector of linear hypothesis weights is of the wrong length for this model. \n")}  
      if (notecode[i,1]==13)
      {cat(" \n")
        cat("NOTE: There are too many pairwise contrasts to conduct with this model. \n")}  
      if (notecode[i,1]==14)
      {cat(" \n")
        cat("NOTE: The number of contrast weights must equal the number of indirect effects. \n")}  
      if (notecode[i,1]==15)
      {cat(" \n")
        cat("NOTE: Monte Carlo confidence intervals not available for this model. \n")
        cat("      Bootstrapping is used instead.\n")}
      if (notecode[i,1]==16)
      {cat(" \n")
        cat("NOTE: The number of Monte Carlo samples was adjusted upward given your desired confidence. \n")}  
      if (notecode[i,1]==19)
      {cat(" \n")
        cat("NOTE: Your contrast matrix is invalid or not applicable to this model. \n")}  
      if (notecode[i,1]==20)
      {cat(" \n")
        cat("NOTE: One of the groups specified by your contrast matrix does not exist in the data. \n")}  
      if (notecode[i,1]==24)
      {cat(" \n")
        cat("NOTE: Total effect model not available with dichotomous Y. \n")}
      if (notecode[i,1]==25)
      {cat(" \n")
        cat("NOTE: STAND/EFFSIZE options not available with dichotomous Y. \n")}
      if (notecode[i,1]==27)
      {cat(" \n")
        cat("NOTE: Standardized coefficients not available for models with moderators. \n")}
      if (notecode[i,1]==28)
      {cat(" \n")
        cat("NOTE: The contrast option is not available with a multicategorical X. \n")}
      if ((notecode[i,1]==29) & (listmiss==1))
      {cat(" \n")
        a<-missrow;conum<-ncol(missrow);allgood<-0;smremain<-12;largesti<-1;smallrow<-0;
        if (conum > 12)
        {
          for (ii in c(1:12))
          { 
            check<-(conum/ii);
            if (check==trunc(check))
            {
              check2<-conum/ii;aok<-ii
              if (aok > 2){allgood<-1}
            }
            if (check != trunc(check))
            {
              remain<-conum-(ii*trunc(check))
              if (remain <= smremain)
              {smremain<-remain;largesti<-ii;smallrow<-trunc(conum/largesti)}
            }
          }
          atemp<-a[1,1:(aok*check2)]
          atemp<-matrix(atemp,nrow=check2,byrow=TRUE)
          cat("NOTE: Missing data resulted in the deletion of the following row(s) of: \n")
          cat("      data from the analysis: \n")
          if (ncol(atemp) > 2)
          {prmatrix(atemp, rowlab=rep("    ",nrow(atemp)),collab=rep("    ",ncol(atemp)))}
          if (allgood==0)
          {
            atemp<-a[1,1:(smallrow*largesti)]
            atemp<-matrix(atemp,nrow=smallrow,byrow=TRUE)
            btemp<-matrix(a[1,((largesti*smallrow)+1):conum])
            prmatrix(atemp,rowlab=rep("    ",nrow(atemp)),collab=rep("    ",ncol(atemp)))
            prmatrix(btemp,rowlab=rep("    ",1),collab=rep("    ",ncol(btemp)))
          }
        }
        if (conum <= 12)
        {
          cat("NOTE: Missing data resulted in the deletion of the following row(s) of: \n")
          cat("      data from the analysis: \n")
          prmatrix(a, rowlab=rep("    ",nrow(a)), collab=rep("    ",ncol(a)))
        }

      }
    }
    if ((nmiss > 0) & (listmiss==0))
    {cat(" \n")
      cat("NOTE: Some cases with missing data were deleted. The number of deleted cases was: ")
      write.table(nmiss,quote=FALSE,row.names=FALSE,col.names=FALSE)}
  }

  # print errors
  for (i in c(1:100)) 
  {
    if (errcode[i,1]==1)
    {cat(" \n")
      cat("ERROR: You must specify a Y and an X variable. \n")}  
    if (errcode[i,1]==2)
    {cat(" \n")
      cat("ERROR: X, M, or Y variable used more than once or W and Z are the same variable. \n")}  
    if (errcode[i,1]==3)
    {cat(" \n")
      cat("ERROR: You have specified more than one variable for W, Y, X, or Z. \n")} 
    if (errcode[i,1]==4)
    {cat(" \n")
      cat("ERROR: A variable specified as multicategorical has more than nine categories. \n")} 
    if (errcode[i,1]==5)
    {cat(" \n")
      cat("ERROR: One of the categories contains only a single case. \n")} 
    if (errcode[i,1]==6)
    {cat(" \n")
      cat("ERROR: Invalid model number in this version of PROCESS. \n")}  
    if (errcode[i,1]==7)
    {cat(" \n")
      cat("ERROR: Invalid model number. \n")}  
    if (errcode[i,1]==8)
    {cat(" \n")
      cat("ERROR: You must specify an M variable for this model. \n")} 
    if (errcode[i,1]==9)
    {cat(" \n")
      cat("ERROR: You have specified an M variable in a model that does not use it. \n")}
    if (errcode[i,1]==10)
    {cat(" \n")
      cat("ERROR: You have specified a W variable in a model that does not use it. \n")}
    if (errcode[i,1]==11)
    {cat(" \n")
      cat("ERROR: You have not specified a W variable in a model that requires it. \n")}
    if (errcode[i,1]==12)
    {cat(" \n")
      cat("ERROR: You have specified a Z variable in a model that does not use it. \n")}
    if (errcode[i,1]==13)
    {cat(" \n")
      cat("ERROR: You have not specified a Z variable in a model that requires it. \n")}
    if (errcode[i,1]==15)
    {cat(" \n")
      cat("ERROR: One of your model variables exhibits no variation (it is a constant). \n")}
    if (errcode[i,1]==16)
    {cat(" \n")
      cat("ERROR: BMATRIX is not the correct length or is otherwise invalid. \n")}
    if (errcode[i,1]==17)
    {cat(" \n")
      cat("ERROR: WMATRIX is not the correct length or is otherwise invalid. \n")}
    if (errcode[i,1]==18)
    {cat(" \n")
      cat("ERROR: ZMATRIX is not the correct length or is otherwise invalid. \n")}
    if (errcode[i,1]==19)
    {cat(" \n")
      cat("ERROR: WZMATRIX is not the correct length or is otherwise invalid. \n")}
    if (errcode[i,1]==20)
    {cat(" \n")
      cat("ERROR: A path fixed at zero cannot be moderated. \n")}
    if (errcode[i,1]==60)
    {cat(" \n")
      cat("PROCESS is now ready for use.\n")
      cat("Copyright 2022 by Andrew F. Hayes ALL RIGHTS RESERVED\n")
      cat("Workshop schedule at http://haskayne.ucalgary.ca/CCRAM\n \n")}
    if (errcode[i,1]==21)
    {cat(" \n")
      cat("ERROR: If only one moderator is specified, it must be specified as W. \n")}
    if (errcode[i,1]==22)
    {cat(" \n")
      cat("ERROR: In BMATRIX, X must be specified to affect at least one variable. \n")}
    if (errcode[i,1]==23)
    {cat(" \n")
      cat("ERROR: In BMATRIX, at least one variable must be specified to affect Y. \n")}
    if (errcode[i,1]==24)
    {cat(" \n")
      cat("ERROR: You must specify a model number or a custom BMATRIX specification. \n")}  
    if (errcode[i,1]==25)
    {cat(" \n")
      cat("ERROR: BMATRIX cannot be used in conjunction with a model number. \n")}
    if (errcode[i,1]==26)
    {cat(" \n")
      cat("ERROR: Your model has a dangling mediator (all Ms must affect and be affected). \n")}
    if (errcode[i,1]==29)
    {cat(" \n")
      cat("ERROR: CMATRIX is not the correct length or is otherwise invalid. \n")}  
    if (errcode[i,1]==30)
    {cat(" \n")
      cat("ERROR: In CMATRIX, all covariates must be assigned to an M or a Y. \n")}
    if ((errcode[i,1]==31) & (singlr==0))
    {cat(" \n")
      singlr<-1
      cat("ERROR: A linear or near linear dependence (singularity) exists in the data. \n")}
    if (errcode[i,1]==32)
    {cat(" \n")
      cat("ERROR: Models 80 and 81 require between 3 and 6 mediators. \n")}  
    if (errcode[i,1]==33)
    {cat(" \n")
      cat("ERROR: Model 82 requires 4 mediators. \n")} 
    if (errcode[i,1]==34)
    {cat(" \n")
      cat("ERROR: This model number requires between 2 and 6 mediators. \n")} 
    if (errcode[i,1]==35)
    {cat(" \n")
      cat("ERROR: In a model with only one moderator, that moderator must be W. \n")}
    if (errcode[i,1]==36)
    {cat(" \n")
      cat("ERROR: A serial mediation model cannot have more than 6 mediators. \n")}  
    if (errcode[i,1]==37)
    {cat(" \n")
      cat("ERROR: No more than 10 mediators are allowed in a PROCESS command. \n")} 
    if (errcode[i,1]==38)
    {cat(" \n")
      cat("ERROR: XCATCODE is not provided, not the correct length, or is otherwise invalid. \n")} 
    if (errcode[i,1]==39)
    {cat(" \n")
      cat("ERROR: WCATCODE is not provided, not the correct length, or is otherwise invalid. \n")} 
    if (errcode[i,1]==40)
    {cat(" \n")
      cat("ERROR: ZCATCODE is not provided, not the correct length, or is otherwise invalid. \n")} 
    if (errcode[i,1]==41)
    {cat(" \n")
      cat("ERROR: Models 1, 2, and 3 cannot be customized. \n")}
    if (errcode[i,1]==43)
    {cat(" \n")
      cat("ERROR: PROCESS does not allow dichotomous mediators. \n")}
    if (errcode[i,1]==50)
    {cat(" \n")
      cat("ERROR: A multicategorical moderator cannot be specified as a covariate. \n")} 
    if (errcode[i,1]==51)
    {cat(" \n")
      cat("ERROR: A variable you specified as a covariate is a moderator in all equations. \n")} 
    if (errcode[i,1]==62)
    {cat(" \n")
      cat("ERROR: After listwise deletion of cases with missing data, too few cases remain. \n")} 
    if ((errcode[i,1]==52) & (mcerpt==0))
    {mcerpt<-1
    cat(" \n")
    cat("ERROR: A variable specified as multicategorical must have at least three categories. \n")}
    if (errcode[i,1]==53)
    {cat(" \n")
      cat("ERROR: Variables declared as factors are not accepted by PROCESS.\n")}  
  }
  #if (saveboot==1)
  #{resultm<-list(resultm,boots)}
  #invisible(resultm)
  resultms<-NULL
  if ((saveboot==0) & (saveest==1)){resultms<-resultm}
  if ((saveboot==1) & (saveest==0)){resultms<-boots}
  if ((saveboot==1) & (saveest==1)){resultms<-(list(boots,resultm))}
  invisible(resultms)
}

```

### Participants

* 568 respondents, 13.57% had been in their referent job less than 6 months, 19.20% between 6 months and a year, 49.12% between one and five years, 13.27% between 5 and 10 years, and 4.87% more than 10 years.
* Ages ranged from 18 to 65 with an average of 28.18 years old (SD = 7.53).
* Gender: female (52.58%) or male (46.83%). 
* Job classifications: International Standard Classification of Occupations (ISCO) via the package `labourR` [@R-labourR], and further categorized into “knowledge” (n = 320) versus “skilled” (n = 214) occupations with knowledge workers being identified via ISCO classifications of: 1) professionals, and 2) managers.

The data for this study were collected through Prolific sample,18 or older and holding a full-time or part-time job. Participants were asked to think about their primary job while answering the survey, and upon completion each participant was compensated in the amount of six US dollars.

# Materials

We used 98 statements taken directly from O$^{*}$Net’s “activity” and “context” classifications. Each of the 98 descriptors has potentially unique response categories, but scaling was consistently 1 (low) to 5 (high). Subsequent to these self-evaluations, respondents were asked to rate elements in terms of 1) …this aspect of your job is a resource that can be functional in achieving work goals, reduce job demands, or stimulate personal growth/development, 2) …this aspect of your job is a challenge that can promote mastery, personal growth, or future gains, and 3) …this aspect of your job is a hindrance that can inhibit personal growth, learning, and work goal attainment.

## Procedure

We used PROCESS for R Version 4.1.1 [@hayes2022] to assess the extent to which the relationship between demands and stress are moderated by resources.

# Results

```{r analyses, fig.cap="Interaction between hindrances and resources as predictors of stress"}

process(activate=1)
process(data=data, y="stress", x="overall.hindrance", w="overall.resource", model=1, jn=1, plot=1)

#Do resources moderate the relationship between hindrances and stress such 
#that more total resources weaken the relationship between
#hindrances and stress?

#Plot the interaction
x<-c(1.6667, 2.2894, 3.2416, 1.6667, 2.2894, 3.2416, 1.6667, 2.2894, 3.2416)
w<-c(3.2983, 3.2983, 3.2983, 3.7402, 3.7402, 3.7402, 4.2063, 4.2063,4.2063)
y<-c(2.6985, 2.8237, 3.0151, 2.7039, 2.7763, 2.8871, 2.7096, 2.7264, 2.7520)
wmarker<-c(15,15,15,16,16,16,17,17,17)
#par(mar=c(1,1,1,1))
plot(y=y,x=x,cex=1.2,pch=wmarker,xlab="Resources (W)",
     ylab="Stress (Y)")
legend.txt<-c("Low Resources","Medium Resources", "High Resources")
legend("topleft", legend = legend.txt,cex=1,lty=c(1,3,6),lwd=c(2,3,2),
       pch=c(15,16,17))
lines(x[w==3.2983],y[w==3.2983],lwd=2,col="black")
lines(x[w==3.7402],y[w==3.7402],lwd=3,lty=3,col="black")
lines(x[w==4.2063],y[w==4.2063],lwd=2,lty=6,col="black")

```

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup