### 2022-08-24 SIOP2023/SIOP_PROCESS.Rmd

---
title             : "JD-R Theory: Using the Content of the O*Net"
shorttitle        : "SIOP_Hindrances-Resources Interaction"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib", "SIOP_PROCESS.bib", "jdr.bib", "jobdemands_resources_bibtex.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

csl               : apa7.csl
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

The Job Demands-Resources Theory [JD-R; @demerouti2001job] has received wide support across contexts and varied research questions. We add to this literature via two routes: by utilizing some of the job characteristics in the popular O\*Net, and by focusing on perceptions of all characteristics as demands/resources. Specifically, we explore the interaction between perceptions of job demands and resources on the outcome of stress across a wide range of occupations.

Here, respondents made a series of evaluations that used: direct O\*Net terminology (both descriptor and response option), and JD-R influenced ratings of demand and hindrance stressors. Prior to a description of results, a brief overview of both the JD-R theory, the stress appraisal process, and O\*Net, is provided.

## The Job demands-Resources Theory

The job demands-resources theory is an expansion of the well-studied job demands-resources model [@demerouti2001job]. One of the major advantages of the job demands-resources theory is that it allows us to model both work environment and job characteristics via job resources and demands, which are thoroughly documented by job in O\*Net. *Resources* are defined as physical, psychological, social, or organizational aspects of the job that may help an employee achieve work goals, reduce job demands, or promote personal growth and development [@demerouti2001job]. *Demands*, on the other hand, include components of a job that require sustained effort, and as such, produce psychological or physiological strain (e.g., high work pressure; @demerouti2001job). 
Cognitively, the perception of an element of one's job as a resource or demand activates one of two unique processes: health impairment (resulting from demands) or motivation [resulting from resources; @bakker2014job]. Demanding job characteristics are frequently associated with negative outcomes [e.g., @bakker2003dual], whereas job characteristics deemed resources have been associated with positive organizational outcomes like engagement and motivation [@bakker2007job]. Our focus is on whether or not having more resources serves as a buffer to the demand-stress relationship. One of the stickier elements of this question surrounds the subjective nature of demands/resources, which we address next.

## Objective vs. Subjective Nature of Demands and Resources: The Role of Appraisal

@searle2015merits note that much of our research on workplace demands is based on apriori classifications of demands. However, the stress experience, or process, described early on by @lazarus1984stress is grounded in the assumption that individual appraisals of stressors/demands vary. Their transactional theory of stress and coping states that people continuously appraise stimuli in their environments. An appraisal is the cognitive process whereby meaning is assigned to a stimulus. If a stimulus is appraised as a stressor (threat, challenge, potentially harmful), emotional distress leads to coping of some kind. This action to cope is also associated with another appraisal about the outcome itself and the process continues if the outcomes is not appraised as favorable [@lazarus1984stress]. The stress appraisal process suggests that classifying a job characteristic or environmental condition as an objective demand or resource might be in error.

We next consider the (limited) empirical evidence on this topic. First, some relatively recent research suggests that job demands and resources may not be universally appraised or assigned as such. Starting with job demands, @webster2011extending, for example, studied workload, role ambiguity, and role conflict demands, and found that while each could be appraised primarily as a challenge or hindrance demand, they could also simultaneously be perceived as being both a challenge and hindrance to different degrees. While their study did include resources, it nonetheless points to individual differences on how people perceive stressors at work. Although part of a much larger study on retirement, @sonnega_comparison_2018 compared self-reported (subjective) ratings of degree of physical demand, stress, and need for intense concentration from the Health and Retirement Study with objective ratings from O*Net. Correlations physical demand (r = .52), stress (r = .10), and need for intense concentration (r = .14), again suggesting perhaps that our objective ratings of job demands (and resources) may be subject to a greater level of individual difference than assumed. Next considering resources, @schmitz_interpreting_2019 also captured subjective and objective resources in their study of retirement. Correlations of composite variables for the resources of autonomy (r = .12. p > .01), recognition of work (r = .07, p > .01), and decision freedom (r = .08, p > .01), while significant, certainly do not reflect high levels of overlap. We do acknowledge as well, that demands and resources are not necessarily consistent across days, or seasons, for many employees. @downes2021incorporating meta-analysis addresses this reality in depth, although it is beyond the scope of this project.

Thus, while it is cleaner to be able to categorize job characteristics as *either* a demand or a resource, the above research suggests that individual appraisal is an important consideration. It is quite possible that one person experiences high work pressure (commonly cited as a demand in the literature) as a hindrance stressor and thus experiences strain, and another thrives in a fast-paced pressured role and would thus find the environment motivating. Here, we asked respondents to rate all of the job characterstics in terms of hindrances and resources.  consider the here whether perceptions of demands (specifically hindrance demands) and stress is

## Why use the O\*Net Resource?

Originally, the Advisory Panel for the Dictionary of Occupational Titles recommended a system that would “…promote the effective education, training, counseling, and employment of the American workforce. It should accomplish its purpose by providing a database system that identifies, defines, classifies, and describes occupations in the economy in an accessible and flexible manner” (@advisory1993new, p. 6). The result was the now commonly used O\*NET. The Occupational Information Network (O\*NET; onetonline.org) contains a comprehensive description of occupations [@peterson2001understanding]. This widely accessed database houses hundreds of standardized and occupation-specific descriptors most occupations in the US and these descriptions are continually updated. In fact, there was a call to work with experienced I/O psychologists over the summer to update the content for the [Industrial and Organizational Psychologist listing on O\*Net](https://www.onetonline.org/link/summary/19-3032.00). These data, and the tools provided for free on the website (e.g., Career Exploration Tools, “My Next Move for Veterans”, “My Next Move”, Toolkit for Business) are frequently used by counselors, students, human resources departments, and researchers to assist potential applicants discover the skills and training they need for the job of their choice. It is also useful to employers by providing them with information with which to craft job descriptions and help employees determine what skills are needed for promotion.

Of greatest interest here are statements taken from O\*NET ["activity" and "context" classifications](https://www.O\*NETonline.org/find/descriptor/result/4.A.1.b.3) (e.g., items related to information input, interacting with others, physical work conditions, structural job characteristics). One of the first and basic questions is whether or not the categorical examples of "resources" and "demands" described in the Job Demands-Resources Theory [@demerouti2001job], for example, are generally deemed resources or demands as we objectively define them. The next logical question surrounds how "universal" such ratings are. For instance, it is quite possible, given the theoretical and empirical evidence presented above, that there is wide variability in individual appraisal of work activities and context such that some people may rate a given activity as a resource and others a hindrance. A second study extends the findings from Study 1 to a potentially key moderator - job categories/classifications, examining whether ratings of resources, challenge- and hindrance demands differ by job classification.

# Methods

## Participants

# Methods

## Participants

```{r getdata}

data <- read.csv("SIOPcleaned.csv")      ## from "datacleaning.R" script

## O*NET categories

data$onet.resource.ii <- rowMeans(data[c(162:166)], na.rm=TRUE)
data$onet.resource.mp <- rowMeans(data[c(167:176)], na.rm=TRUE)
data$onet.resource.wo <- rowMeans(data[c(177:185)], na.rm=TRUE)
data$onet.resource.io <- rowMeans(data[c(186:202)], na.rm=TRUE)
data$onet.resource.ir <- rowMeans(data[c(149:161)], na.rm=TRUE)
data$onet.resource.pc <- rowMeans(data[c(128:144,146:148)], na.rm=TRUE)
data$onet.resource.sc <- rowMeans(data[c(118:127)], na.rm=TRUE)

data$overall.resource <- rowMeans(data[c(118:144,147:202)], na.rm=TRUE)

data$onet.hindrance.ii <- rowMeans(data[c(246:247,249:252)], na.rm=TRUE)
data$onet.hindrance.mp <- rowMeans(data[c(252:261)], na.rm=TRUE)
data$onet.hindrance.wo <- rowMeans(data[c(262:270)], na.rm=TRUE)
data$onet.hindrance.io <- rowMeans(data[c(271:287)], na.rm=TRUE)
data$onet.hindrance.ir <- rowMeans(data[c(233:245)], na.rm=TRUE)
data$onet.hindrance.pc <- rowMeans(data[c(213:232)], na.rm=TRUE)
data$onet.hindrance.sc <- rowMeans(data[c(203:212)], na.rm=TRUE)

data$overall.hindrance <- rowMeans(data[c(203:247,249:287)], na.rm=TRUE)

data$onet.challenge.ii <- rowMeans(data[c(332:336)], na.rm=TRUE)
data$onet.challenge.mp <- rowMeans(data[c(337:346)], na.rm=TRUE)
data$onet.challenge.wo <- rowMeans(data[c(347:355)], na.rm=TRUE)
data$onet.challenge.io <- rowMeans(data[c(356:372)], na.rm=TRUE)
data$onet.challenge.ir <- rowMeans(data[c(319:331)], na.rm=TRUE)
data$onet.challenge.pc <- rowMeans(data[c(298:307,309:318)], na.rm=TRUE)
data$onet.challenge.sc <- rowMeans(data[c(288:297)], na.rm=TRUE)

data$overall.challenge <- rowMeans(data[c(288:307,309:372)], na.rm=TRUE)

data$burnout      <- rowMeans(data[c(373:376)], na.rm=TRUE)
data$stress       <- rowMeans(data[c(377:379)], na.rm=TRUE)
data$engagement   <- rowMeans(data[c(380:399)], na.rm=TRUE)        ## added 10/12 because analyses too wonky with 3 subscales

```

There were `r nrow(data)` respondents.

```{r PROCESS, include=FALSE}

source("process.R", local = knitr::knit_global())

```

### Participants

* 568 respondents, 13.57% had been in their referent job less than 6 months, 19.20% between 6 months and a year, 49.12% between one and five years, 13.27% between 5 and 10 years, and 4.87% more than 10 years.
* Ages ranged from 18 to 65 with an average of 28.18 years old (SD = 7.53).
* Gender: female (52.58%) or male (46.83%). 
* Job classifications: International Standard Classification of Occupations (ISCO) via the package `labourR` [@R-labourR], and further categorized into “knowledge” (n = 320) versus “skilled” (n = 214) occupations with knowledge workers being identified via ISCO classifications of: 1) professionals, and 2) managers.

The data for this study were collected through Prolific sample,18 or older and holding a full-time or part-time job. Participants were asked to think about their primary job while answering the survey, and upon completion each participant was compensated in the amount of six US dollars.

# Materials

We used 98 statements taken directly from O$^{*}$Net’s “activity” and “context” classifications. Each of the 98 descriptors has potentially unique response categories, but scaling was consistently 1 (low) to 5 (high). Subsequent to these self-evaluations, respondents were asked to rate elements in terms of 1) …this aspect of your job is a resource that can be functional in achieving work goals, reduce job demands, or stimulate personal growth/development, 2) …this aspect of your job is a challenge that can promote mastery, personal growth, or future gains, and 3) …this aspect of your job is a hindrance that can inhibit personal growth, learning, and work goal attainment.

## Procedure

We used PROCESS for R Version 4.1.1 [@hayes2022] to assess the extent to which the relationship between demands and stress are moderated by resources.

# Results

```{r analyses, fig.cap="Interaction between hindrances and resources as predictors of stress"}

process(activate=1)
results <- process(data=data, y="stress", x="overall.hindrance", w="overall.resource", model=1, jn=1, plot=1, save=2)

#Do resources moderate the relationship between hindrances and stress such 
#that more total resources weaken the relationship between
#hindrances and stress?

#Plot the interaction
x<-c(1.6667, 2.2894, 3.2416, 1.6667, 2.2894, 3.2416, 1.6667, 2.2894, 3.2416)
w<-c(3.2983, 3.2983, 3.2983, 3.7402, 3.7402, 3.7402, 4.2063, 4.2063,4.2063)
y<-c(2.6985, 2.8237, 3.0151, 2.7039, 2.7763, 2.8871, 2.7096, 2.7264, 2.7520)
wmarker<-c(15,15,15,16,16,16,17,17,17)
#par(mar=c(1,1,1,1))
plot(y=y,x=x,cex=1.2,pch=wmarker,xlab="Resources (W)",
     ylab="Stress (Y)")
legend.txt<-c("Low Resources","Medium Resources", "High Resources")
legend("topleft", legend = legend.txt,cex=1,lty=c(1,3,6),lwd=c(2,3,2),
       pch=c(15,16,17))
lines(x[w==3.2983],y[w==3.2983],lwd=2,col="black")
lines(x[w==3.7402],y[w==3.7402],lwd=3,lty=3,col="black")
lines(x[w==4.2063],y[w==4.2063],lwd=2,lty=6,col="black")

```

#Results

A moderated regression including hindrances, resources, and the interaction between them was done using PROCESS, version 4.1.1. First, the overall regression model including mean hindrances, mean resources, and the interaction between the two variables was significant, F(3, 564) = 3.29, p = .020.  The interaction between hindrance and resources (uncentered) revealed that the relationship between hindrances and stress was conditional on resources, F(3, 564) = 3.51, p = .061. As can be seen in Figure 1, those with fewer resources show a much stronger positive relationship between hindrances and stress than those with more resources. Upon exploring the interaction further, it was evident that this moderated effect happened at lower, but not higher levels of resources.

```{r table}

Component <- c("Constant", "Hindrance (X)", "Resource (W)", "Hindrance x Resource")
coeff <- format(round(results[c(2:5),1],2), nsmall=2)
SE    <- format(round(results[c(2:5),2],2), nsmall=2)
t     <- format(round(results[c(2:5),3],2), nsmall=2)
p     <- format(round(results[c(2:5),4],2), nsmall=2)

table <- as.data.frame(cbind(Component, coeff, SE, t, p))

papaja::apa_table(table,
                  caption = "Results from a regression analysis examining the moderation of resources on the relationship between hindrance demands and stress",
                  note = "R^2 etc here",
                  landscape = FALSE,
                  escape = TRUE)

```

Next steps:
1)	Make a prettier graph using ggplot. Note the percentiles that we graphed. 
2)	Or, could generate Johnson-Neyman Technique – shows specifically the range of W values that are significant. See p. 272 for visual and r-script. 
Table example on p. 286: includes variables, symbol, coeff, SE, t, p; below that, rsquared, MSE and f-string.

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup