### 2022-09-02 SIOP2023/SIOP_PROCESS.Rmd
Updated introduction. See notes near end of intro for next steps.

---
title             : "JD-R Theory: Using the Content of the O*Net"
shorttitle        : "SIOP_Hindrances-Resources Interaction"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib", "SIOP_PROCESS.bib", "jdr.bib", "jobdemands_resources_bibtex.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

csl               : apa7.csl
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

The Job Demands-Resources Theory [JD-R; @demerouti2001job] has received wide support across contexts and varied research questions. We add to this literature via two routes: by utilizing some of the job characteristics in the popular O\*Net, and by focusing on perceptions of all characteristics as demands/resources. Specifically, we explore the interaction between perceptions of job demands and resources on the outcome of stress across a wide range of occupations. Here, respondents made a series of evaluations that used: direct O\*Net terminology (both descriptor and response option), and JD-R influenced ratings of demand and hindrance stressors. Prior to a description of results, a brief overview of both the JD-R theory, the stress appraisal process, and O\*Net, is provided.

## The Job demands-Resources Theory

The job demands-resources theory is an expansion of the well-studied job demands-resources model [@demerouti2001job]. One of the major advantages of the job demands-resources theory is that it allows us to model both work environment and job characteristics via job resources and demands, which are thoroughly documented by job in O\*Net. *Resources* are defined as physical, psychological, social, or organizational aspects of the job that may help an employee achieve work goals, reduce job demands, or promote personal growth and development [@demerouti2001job]. *Demands*, on the other hand, include components of a job that require sustained effort, and as such, produce psychological or physiological strain (e.g., high work pressure; @demerouti2001job).

Cognitively, the perception of an element of one's job as a resource or demand activates one of two unique processes: health impairment (resulting from demands) or motivation [resulting from resources; @bakker2014job]. Demanding job characteristics are frequently associated with negative outcomes [e.g., @bakker2003dual], whereas job characteristics deemed resources have been associated with positive organizational outcomes like engagement and motivation [@bakker2007job]. Our focus is on whether or not having more resources serves as a buffer to the demand-stress relationship. One of the stickier elements of this question surrounds the subjective nature of demands/resources, which we address next.

## Objective vs. Subjective Nature of Demands and Resources: The Role of Appraisal

@searle2015merits note that much of our research on workplace demands is based on apriori classifications of demands. For instance, we assume that generally, time pressure is a negative demand on an employee. However, the stress experience, or process, described early on by @lazarus1984stress is grounded in the assumption that individual appraisals of stressors/demands vary. Their transactional theory of stress and coping states that people continuously appraise stimuli in their environments. An appraisal is the cognitive process whereby meaning is assigned to a stimulus. If a stimulus is appraised as a stressor (threat, challenge, potentially harmful), emotional distress leads to coping of some kind. This action to cope is also associated with another appraisal about the outcome itself and the process continues if the outcomes is not appraised as favorable [@lazarus1984stress]. As such, the stress appraisal process suggests that classifying a job characteristic or environmental condition as an objective demand or resource might be in error.

We next consider the empirical evidence on the subjective nature of demands and resources.  First, as hinted at above, some research suggests that job demands and resources may not be universally appraised or assigned as such. Starting with job demands, @webster2011extending  studied workload, role ambiguity, and role conflict demands, and found that while each could be appraised primarily as a challenge or hindrance demand, they could also simultaneously be perceived as being *both* a challenge and hindrance to different degrees. While their study not did include resources, it documents individual differences in how people perceive stressors at work. Although not the primary focus of their paper, @sonnega_comparison_2018 compared self-reported (subjective) ratings of degree of physical demand, stress, and need for intense concentration from the Health and Retirement Study with objective ratings from O*Net. Correlations physical demand (r = .52), stress (r = .10), and need for intense concentration (r = .14), again suggesting perhaps that our objective ratings of job demands (and resources) may be subject to a greater level of individual difference than we tend to think. While the above two studies provide evidence for variability in perception of demands, @schmitz_interpreting_2019 captured subjective and objective resources in their study of retirement. Correlations of composite variables between subjective and objective measures for the resources of autonomy (r = .12. p > .01), recognition of work (r = .07, p > .01), and decision freedom (r = .08, p > .01), while significant, certainly do not reflect high levels of overlap. We do acknowledge as well, that demands and resources are not necessarily consistent across days, or seasons, for many employees. @downes2021incorporating meta-analysis addresses this reality in depth, although it is beyond the scope of this project.

Thus, while it is cleaner to be able to categorize job characteristics as *either* a demand or a resource, the above research suggests that individual appraisal is an important consideration. It is quite possible that one person experiences high work pressure (commonly cited as a demand in the literature) as a hindrance stressor and thus experiences strain, and another thrives in a fast-paced pressured role and would thus find the environment motivating. Here, we asked respondents to rate all of the job characteristics in terms of hindrances, challenges, and resources.

## Value of exploring the O\*Net Resource

First, the Occupational Information Network (O\*NET; onetonline.org) contains a comprehensive description of occupations [@peterson2001understanding]. This widely accessed database houses hundreds of standardized and occupation-specific descriptors most occupations in the US and these descriptions are continually updated. These data, and the tools provided for free on the website (e.g., Career Exploration Tools, “My Next Move”, Toolkit for Business) are frequently used by counselors, students, human resources departments, and researchers to assist potential applicants discover the skills and training they need for the job of their choice. It is also useful to employers by providing them with information with which to craft job descriptions and help employees determine what skills are needed for promotion. We utilized statements taken from O\*NET ["activity" and "context" classifications](https://www.O\*NETonline.org/find/descriptor/result/4.A.1.b.3) (e.g., items related to information input, interacting with others, physical work conditions, structural job characteristics).

# Current Study and Hypotheses

These data were taken from a larger study on JD-R theory as it applies to O\*NET items. Our specific interest in the current study is in whether or not *perceptions* of hindrance demands are postitively related to perceived stress, and whether or not this relationship is moderated by perceived resources. The Job demands-Resources theory would suggest resources would buffer this relationship. We do have some existing evidence that this occurs with other outcomes. For example, Tadic et al. (2015) found that daily hindrance job demands were negatively related to both positive affect and engagement in a sample of primary school teachers. Daily job resources, in this sample, buffered the relationships between hindrances and affect and engagement. Here, we propose that perceived resources generally, as opposed to daily, would also buffer the relationship between perceived hindrance stressors and, in this, case, perceived stress. The following two predictions are made:

*H1. There is a positive relationship between perceived hindrance stressors and stress.*

*H2. The relationship between mean perceived hindrances and stress will be moderated by resources such that this relationship is diminished as perceived resources increase.*

NOTES:
Add: Challenge versus hindrance job demands and wellbeing: A diary study on the moderating role of job resources
Add a correlation table to speak to H1. Reference H1 and H2 in the results section.

# Methods

```{r getdata}

data <- read.csv("SIOPcleaned.csv")      ## from "datacleaning.R" script

## O*NET categories

data$onet.resource.ii <- rowMeans(data[c(162:166)], na.rm=TRUE)
data$onet.resource.mp <- rowMeans(data[c(167:176)], na.rm=TRUE)
data$onet.resource.wo <- rowMeans(data[c(177:185)], na.rm=TRUE)
data$onet.resource.io <- rowMeans(data[c(186:202)], na.rm=TRUE)
data$onet.resource.ir <- rowMeans(data[c(149:161)], na.rm=TRUE)
data$onet.resource.pc <- rowMeans(data[c(128:144,146:148)], na.rm=TRUE)
data$onet.resource.sc <- rowMeans(data[c(118:127)], na.rm=TRUE)

data$overall.resource <- rowMeans(data[c(118:144,147:202)], na.rm=TRUE)

data$onet.hindrance.ii <- rowMeans(data[c(246:247,249:252)], na.rm=TRUE)
data$onet.hindrance.mp <- rowMeans(data[c(252:261)], na.rm=TRUE)
data$onet.hindrance.wo <- rowMeans(data[c(262:270)], na.rm=TRUE)
data$onet.hindrance.io <- rowMeans(data[c(271:287)], na.rm=TRUE)
data$onet.hindrance.ir <- rowMeans(data[c(233:245)], na.rm=TRUE)
data$onet.hindrance.pc <- rowMeans(data[c(213:232)], na.rm=TRUE)
data$onet.hindrance.sc <- rowMeans(data[c(203:212)], na.rm=TRUE)

data$overall.hindrance <- rowMeans(data[c(203:247,249:287)], na.rm=TRUE)

data$onet.challenge.ii <- rowMeans(data[c(332:336)], na.rm=TRUE)
data$onet.challenge.mp <- rowMeans(data[c(337:346)], na.rm=TRUE)
data$onet.challenge.wo <- rowMeans(data[c(347:355)], na.rm=TRUE)
data$onet.challenge.io <- rowMeans(data[c(356:372)], na.rm=TRUE)
data$onet.challenge.ir <- rowMeans(data[c(319:331)], na.rm=TRUE)
data$onet.challenge.pc <- rowMeans(data[c(298:307,309:318)], na.rm=TRUE)
data$onet.challenge.sc <- rowMeans(data[c(288:297)], na.rm=TRUE)

data$overall.challenge <- rowMeans(data[c(288:307,309:372)], na.rm=TRUE)

data$burnout      <- rowMeans(data[c(373:376)], na.rm=TRUE)
data$stress       <- rowMeans(data[c(377:379)], na.rm=TRUE)
data$engagement   <- rowMeans(data[c(380:399)], na.rm=TRUE)        ## added 10/12 because analyses too wonky with 3 subscales

```

There were `r nrow(data)` respondents.

```{r PROCESS, include=FALSE}

source("process.R", local = knitr::knit_global())

```

### Participants

* 568 respondents, 13.57% had been in their referent job less than 6 months, 19.20% between 6 months and a year, 49.12% between one and five years, 13.27% between 5 and 10 years, and 4.87% more than 10 years.
* Ages ranged from 18 to 65 with an average of 28.18 years old (SD = 7.53).
* Gender: female (52.58%) or male (46.83%). 
* Job classifications: International Standard Classification of Occupations (ISCO) via the package `labourR` [@R-labourR], and further categorized into “knowledge” (n = 320) versus “skilled” (n = 214) occupations with knowledge workers being identified via ISCO classifications of: 1) professionals, and 2) managers.

The data for this study were collected through Prolific sample,18 or older and holding a full-time or part-time job. Participants were asked to think about their primary job while answering the survey, and upon completion each participant was compensated in the amount of six US dollars.

# Materials

We used 98 statements taken directly from O$^{*}$Net’s “activity” and “context” classifications. Each of the 98 descriptors has potentially unique response categories, but scaling was consistently 1 (low) to 5 (high). Subsequent to these self-evaluations, respondents were asked to rate elements in terms of 1) …this aspect of your job is a resource that can be functional in achieving work goals, reduce job demands, or stimulate personal growth/development, 2) …this aspect of your job is a challenge that can promote mastery, personal growth, or future gains, and 3) …this aspect of your job is a hindrance that can inhibit personal growth, learning, and work goal attainment.

## Procedure

We used PROCESS for R Version 4.1.1 [@hayes2022] to assess the extent to which the relationship between demands and stress are moderated by resources.

# Results

```{r analyses, fig.cap="Interaction between hindrances and resources as predictors of stress"}

process(activate=1)
results <- process(data=data, y="stress", x="overall.hindrance", w="overall.resource", model=1, jn=1, plot=1, save=2)

#Do resources moderate the relationship between hindrances and stress such 
#that more total resources weaken the relationship between
#hindrances and stress?

#Plot the interaction
x<-c(1.6667, 2.2894, 3.2416, 1.6667, 2.2894, 3.2416, 1.6667, 2.2894, 3.2416)
w<-c(3.2983, 3.2983, 3.2983, 3.7402, 3.7402, 3.7402, 4.2063, 4.2063,4.2063)
y<-c(2.6985, 2.8237, 3.0151, 2.7039, 2.7763, 2.8871, 2.7096, 2.7264, 2.7520)
wmarker<-c(15,15,15,16,16,16,17,17,17)
#par(mar=c(1,1,1,1))
plot(y=y,x=x,cex=1.2,pch=wmarker,xlab="Resources (W)",
     ylab="Stress (Y)")
legend.txt<-c("Low Resources","Medium Resources", "High Resources")
legend("topleft", legend = legend.txt,cex=1,lty=c(1,3,6),lwd=c(2,3,2),
       pch=c(15,16,17))
lines(x[w==3.2983],y[w==3.2983],lwd=2,col="black")
lines(x[w==3.7402],y[w==3.7402],lwd=3,lty=3,col="black")
lines(x[w==4.2063],y[w==4.2063],lwd=2,lty=6,col="black")

```

Table 1 presents correlations among the study variables of interest.

To explore H2, a moderated regression including hindrances, resources, and the interaction between them was done using PROCESS, version 4.1.1. First, the overall regression model including mean hindrances, mean resources, and the interaction between the two variables was significant, F(3, 564) = 3.29, p = .020.  The interaction between hindrance and resources (uncentered) revealed that the relationship between hindrances and stress was conditional on resources, F(3, 564) = 3.51, p = .061, providing tentative support for H2. As can be seen in Figure 1, those with fewer resources show a much stronger positive relationship between hindrances and stress than those with more resources. Upon exploring the interaction further, it was evident that this moderated effect happened at lower, but not higher levels of resources.

```{r table}

Component <- c("Constant", "Hindrance (X)", "Resource (W)", "Hindrance x Resource")
coeff <- format(round(results[c(2:5),1],2), nsmall=2)
SE    <- format(round(results[c(2:5),2],2), nsmall=2)
t     <- format(round(results[c(2:5),3],2), nsmall=2)
p     <- format(round(results[c(2:5),4],2), nsmall=2)

table <- as.data.frame(cbind(Component, coeff, SE, t, p))

papaja::apa_table(table,
                  caption = "Results from a regression analysis examining the moderation of resources on the relationship between hindrance demands and stress",
                  note = "R^2 etc here",
                  landscape = FALSE,
                  escape = TRUE)

```

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup