### 2022-09-05 Submission_ResourceCount.tex
new siop project

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Submission\_ResourceCount},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Submission\_ResourceCount}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

Next steps (thoughts from 8/24): Want this to be a more traditional
project: DVs engagement, burnout, stress. How to operationalize IVs of
resources, challenges, hindrances. May want to only count it if they
somewhat agree or strongly agree. Could do a count or a sum of these
values. The moderator of job type may work well in this idea.

Want to include idea of item level things in terms of agreement.
decreasing bar graphs relative to SD. Maybe we just present SD or SE.
Thinking on SD, but ensure we have sufficient numbers of people
responding to the item. Want to get a total count.

Alicia reflection 9/1/22: It actually appears that there may be more
here with the engagement variable when reviewing the correlations. Knit
and reflect more on the paper with that outcome being of focus.

John-o: maybe use this as basis for another SIOP - extent to which
resources and challenges are viewed similarly and also associations with
engagement, burnout, etc\ldots{} this operationalization is different
than previous SIOP so could submit anew. Also would need a different
approach to ``similarity'' - would need to look at item level

There were 568 retained respondents.

\begin{verbatim}
##            resource hindrance challenge
## resource  1.0000000 0.2255508 0.8622519
## hindrance 0.2255508 1.0000000 0.2204752
## challenge 0.8622519 0.2204752 1.0000000
\end{verbatim}

\begin{lltable}

\begin{longtable}{llllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:correlation_table}}\\
\toprule
 & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{3} & \multicolumn{1}{c}{4} & \multicolumn{1}{c}{5} & \multicolumn{1}{c}{$M$} & \multicolumn{1}{c}{$SD$}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:correlation_table} continued}}\\
\toprule
 & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{3} & \multicolumn{1}{c}{4} & \multicolumn{1}{c}{5} & \multicolumn{1}{c}{$M$} & \multicolumn{1}{c}{$SD$}\\
\midrule
\endhead
1. resource & - &  &  &  &  & 36.02 & 13.26\\
2. hindrance & .23*** & - &  &  &  & 13.09 & 13.62\\
3. challenge & .86*** & .22*** & - &  &  & 35.64 & 13.63\\
4. burnout & .05 & .04 & .07 & - &  & 3.04 & 0.87\\
5. stress & .06 & .09* & .08 & .70*** & - & 2.59 & 0.97\\
6. engagement & .34*** & .01 & .31*** & -.35*** & -.30*** & 4.04 & 0.83\\
\bottomrule
\end{longtable}

\end{lltable}

\end{document}