### 2023-02-14 SIOP2023/Counts/counts_poster.Rmd
formatting done - content should be double-checked

---
title: 'Perception of Work Demands and Resources: Does Sheer Volume Matter?'
author:
  - name: Alicia Stachowski
    affil: 1
  - name: John Kulas
    affil: 2
affiliation:
  - num: 1
    address: University of Wisconsin - Stout
  - num: 2
    address: eRg
column_numbers: 3
primary_colour: "#313ab3"
secondary_colour: "#B4B5B4"
accent_colour: "#FF0000"
logoright_name: 'Stoutlogo_2023.png'
logoleft_name: "erg.png"
output: 
  posterdown::posterdown_html:
    self_contained: false
csl               : "apa7.csl"
bibliography: jobdemands_resources_bibtex.bib
suppress-bibliography: true 
link-citations: true
floatsintext: yes
affiliation_textsize: "50pt"
body_textsize: "50px"
title_textsize:	"75pt"	
author_textsize:	"50pt"
font_family: 'Helvetica'
main_fontfamily: "Helvetica"
#logoleft_name	"Figures/posterdownlogo.png"	Location of the image to use.
#logoright_name	"Figures/posterdownlogo.png"	Location of the image to use.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

library(tinylabels) ## because of warning on later chunk

```

<style>
.logo_left {
 width: 15%;
}
.logo_right {
 width: 15%;
}
.poster_title{
  padding-top: 5%;
  width: 70%;
}
</style>

# Introduction

Using the the job demands-resources theory [@demerouti2001job], transactional theory of stress and coping [@lazarus1984stress], and  the challenge-hindrance framework [@cavanaugh2000empirical], we explored whether the amount or volume of perceived resources and demands would be related to three organizational outcomes: engagement [“a positive affective experience defined as a fulfilling, work-related state of mind characterized by vigor, dedication, and absorption”, @schaufeli2002measurement], workplace stress [“an individual state characterized by a combination of high arousal and displeasure”, p. 15, @pejtersen2010second] and burnout [“the degree of physical and psychological fatigue and exhaustion that is perceived by the person as related to his/her work”, p. 197, @kristensen2005copenhagen].

Perhaps not surprisingly, challenge stressors are typically associated with positive outcomes, whereas hindrance stressors are associated with more negative outcomes [e.g., @cavanaugh2000empirical]. Thus:

> *Hypothesis 1a-1c*: A greater overall quantity of resources is positively associated with engagement (1a), and negatively associated with stress (1b) and burnout (1c)

> *Hypothesis 2a-2c*: A greater overall quantity of challenge demands is positively associated with engagement (2a), and negatively associated with stress (2b) and burnout (2c).

> *Hypothesis 3a-3c*: A greater overall quantity of hindrance demands is negatively associated with engagement (3a), and positively associated with stress (3b) and burnout (3c).

# Methods

**Participants, Procedure, & Materials**

A Prolific panel sample initially resulted in 785 respondents, of which 568 were retained after cleaning. 13.57% had been in their job <6 months, 19.20% 6-12 months, 49.12% 1-5 years, 13.27% 5-10 year, and 4.87% >10 years. Mean age was 28.18 years (*SD* = 7.53). Self-identified gender was largely female (52.58%) or male (46.83%).

An email was sent to eligible Prolific respondents (those 18+ holding either a full-time or part-time job). A ~45 minute survey asked to think about their primary job. The items presented depended on the job characteristics they initially specified. Respondents were asked to report 1) how much a characteristic was a resource, 2) how much each characteristic was a hindrance, 3) how much each item was a challenge. Participants were compensated six dollars.

**Characteristics, Demands, and Resources**.

Items within O\*Net’s classifications of “work activity” and "work context" items were used here. We used O\*Net’s response scales, which all shared the same 1 to 5 scale options. Subsequent to providing ratings of whether or not a characteristic was relevant, respondents rated items in terms of, 1) . . . this aspect of your job is a resource that can be functional in achieving work goals, reduce job demands, or stimulate personal growth/development, 2) . . . this aspect of your job is a challenge that can promote mastery, personal growth, or future gains, and 3) . . . this aspect of your job is a hindrance that can inhibit personal growth, learning, and work goal attainment.

**Outcomes of Stress, Burnout, and Engagement**.

Three items taken from the Copenhagen Psychosocial Questionnaire [@burr_third_2019] to capture stress. Obtained alpha was .85 in this sample. Four items were taken from the Copenhagen Psychosocial Questionnaire [@burr_third_2019] to capture burnout. Alpha was 0.85 in this sample. The 18-item engagement measure was recently developed [@engage_2022] to capture engagement. For the purposes of the current study, we focused on an overall engagement score (18 item aggregate, alpha = 0.91).

# Results
```{r study1, include=FALSE, message = FALSE, warning = FALSE}
temp <- read.csv("..//JDRJournalSubmission//initial_data_screen.csv", header=FALSE, na.strings="")   ## NOTE: 404 vars (9/28/21) make sure to check indexing used throughout script

x <- paste("item", sep="",1:404)
y <- t(temp[2,])
                                        ## decluttering Qualtrics excess
data <- temp[-c(1:3),]                                           ## Getting rid of all 3 weird Qualtrics rows
colnames(data) <- x

incomplete <- read.csv("..//JDRJournalSubmission//inprogress.csv", header=FALSE, na.strings="")   ## NOTE: 404 vars (9/28/21) make sure to check indexing used throughout script

data2 <- incomplete[-c(1:2),-c(10,12,16,113,199,285,371,379,400, 406:411)]

data3 <- as.data.frame(cbind(data2$V8,data2$V9,data2$V1,data2$V1,data2$V1,data2$V1,data2$V1,data2$V1,data2))

colnames(data3) <- x

rm(x, y, temp)

use <- rbind(data,data3)

## STILL NEED FILE WITH BAD RESPONDENTS IDENTIFIED (E.G., Item_18=2, na > 200, longstring > 20, CARELESS CHECKS > 1)

library(careless)
use$careless_long <- longstring(use[18:399])
use$missing <- rowSums(is.na(use[18:404]))

different <- use[ which(use$item18 == 2 | use$careless_long > 20 | use$missing > 200), ]

attention <- use[which(use$item61 == 5 & use$item145 == 5 & use$item248 == 2 & use$item308 == 3), ]

data <- data.frame(lapply(attention, function(x) as.numeric(as.character(x))))
data <- data[,c(1:404)]

numtot <- nrow(use)
numkicked <- nrow(different)
numvalid <- nrow(data)

```

```{r demos}

a <- descr::freq(use$item401, plot=FALSE)

use$item402[use$item402 == "Female " | use$item402 == "female" | use$item402 == "FEMALE" | use$item402 == "Woman" | use$item402 == "Mujer" | use$item402 == "feminine" | use$item402 == "fEMALE" | use$item402 == "F" | use$item402 == "Femenale" | use$item402 == "feminino" | use$item402 == "female " | use$item402 == "Fem" | use$item402 == "Feminine" | use$item402 == "feminine" | use$item402 == "WOMAN" | use$item402 == "Femal" | use$item402 == "femenine" | use$item402 == "Femenine" | use$item402 == "femenino" | use$item402 == "woman"]  <- "Female"

use$item402[use$item402 == "gerente" | use$item402 == "M" | use$item402 == "Make" | use$item402 == "Mal" | use$item402 == "male" | use$item402 == "MALE" | use$item402 == "male" | use$item402 == "Male" | use$item402 == "mals" | use$item402 == "man" | use$item402 == "masculine" | use$item402 == "Masculine" | use$item402 == "Man" | use$item402 == "mals " | use$item402 == "male " | use$item402 == "Male "]  <- "Male"

b <- descr::freq(use$item402, plot=FALSE)

##Job title
library(labourR)

#ESCO/ISCO classification using the labourR package
#ISCO = International Standard Classification of Occupations

attention$id <- 1:nrow(attention) #adds unique identifier to satisfy inane requirement of classify_occupation()

ISCO <- classify_occupation(attention, text_col = 'item400', isco_level = 1)

m3 <-merge(attention, ISCO, by.x = "id", by.y = "id", all.x = TRUE)

c <- descr::freq(m3$preferredLabel, plot=FALSE)

knowledge <- data[ which(m3$iscoGroup == "1" | m3$iscoGroup == "2"), ]
service <- data[ which(m3$iscoGroup == "3" | m3$iscoGroup == "4" | m3$iscoGroup == "5" | m3$iscoGroup == "6" | m3$iscoGroup == "7" | m3$iscoGroup == "8" | m3$iscoGroup == "9"), ]

data$item403[data$item403 == 2000]  <- 20

```

```{r generalrecodes, include=FALSE}

source("..//JDRJournalSubmission//recodescript.R", local = knitr::knit_global())

```

```{r keeping45, include=FALSE}

## need to run "recodescript.R" prior to this if running chunk-by-chunk

## variables of interest (resource, challenge, hindrance): 118:372

library(dplyr)

 data22 <- data %>% 
 mutate_at(vars(118:372),
 function(x) recode(x,
 1==NA, 
 2==NA,
 3==NA,
 4==4,
 5==5))

## attention checks: 145, 248, 308

data22$resource <- rowSums(data22[c(118:144,146:202)], na.rm=TRUE)
data22$hindrance <- rowSums(data22[c(203:247,249:287)], na.rm=TRUE)
data22$challenge <- rowSums(data22[c(288:307,309:372)], na.rm=TRUE)

data22$item387 <- 7-data22$item387
data22$burnout      <- rowMeans(data22[c(373:376)], na.rm=TRUE)
data22$stress       <- rowMeans(data22[c(377:379)], na.rm=TRUE)
data22$engagement   <- rowMeans(data22[c(380:399)], na.rm=TRUE)        ## probably need to recode some

cor(data22[405:410], use="all.obs")   ## something to be told with resources and challenges

## Peculiar, but repeating for normal datafile so column identifiers line-up

data$resource <- rowSums(data22[c(118:144,146:202)], na.rm=TRUE)
data$hindrance <- rowSums(data22[c(203:247,249:287)], na.rm=TRUE)
data$challenge <- rowSums(data22[c(288:307,309:372)], na.rm=TRUE)

data$burnout      <- rowMeans(data22[c(373:376)], na.rm=TRUE)
data$stress       <- rowMeans(data22[c(377:379)], na.rm=TRUE)
data$engagement   <- rowMeans(data22[c(380:399)], na.rm=TRUE)        ## probably need to recode some

```

```{r matching, include=FALSE}

source("..\\JDRJournalSubmission\\SIOP2023\\Counts\\matching.R", local = knitr::knit_global())  ## adds 252 variables (count matches) to data object

```

```{r rankings, warnings = FALSE, include = FALSE, eval = FALSE}

write.csv(psych::describe(data[411:662]), "..\\JDRJournalSubmission\\SIOP2023\\Counts\\percents.csv")   ## all together

write.csv(psych::describe(data[411:494]), "..\\JDRJournalSubmission\\SIOP2023\\Counts\\resource_hind.csv")   ## 
write.csv(psych::describe(data[495:578]), "..\\JDRJournalSubmission\\SIOP2023\\Counts\\resource_challenge.csv")
write.csv(psych::describe(data[579:662]), "..\\JDRJournalSubmission\\SIOP2023\\Counts\\hind_challenge.csv")

```

Correlations, means and standard deviations are presented in Table \@ref(tab:correlationtable). The results reveal a positive association between resources and engagement (*r* = .34; H1a), but a lack of meaningful association between resources and stress (H1b) and burnout (H1c). Challenge demands were positively associated with engagement (*r* = .31; H2a), but were unrelated to stress or burnout (H2b and H2c). Total hindrance stressors were largely not significantly associated with our outcomes, although there was a small-moderate association with stress (H3b; *r* = .09).

```{r correlationtable, warning=FALSE}

r <- corx::corx(data22[,c(405:410)],                    
                triangle = "lower",
                stars = c(0.05, 0.01, 0.001),
                describe = c(`$M$` = mean, `$SD$` = sd))

papaja::apa_table(r$apa, # apa contains the data.frame needed for apa_table,
                 caption = "Focal variable correlations (counts data).",
                 placement = "H",
                 align = c('l','l','l','l','l','l','r','r'),
                 escape = FALSE)

# cor(data[405:410], use="all.obs")   ## something to be told with resources and challenges

```

```{r regressions, include=FALSE}

eng <- lm(engagement ~ hindrance + challenge + resource, data22)
bur <- lm(burnout ~ hindrance + challenge + resource, data22)
str <- lm(stress ~ hindrance + challenge + resource, data22)

summary(eng)
summary(bur)
summary(str)

```

We also conducted three regression analyses: regressing a) engagement, b) stress, and c) burnout separately onto total resources, challenge and hindrance demands. The total resources were predictive of engagement ($F_{(3, 564)}$ = 26.41, p < .001), but neither challenge nor hindrance demands predicted engagement; see Table \@ref(tab:regressiontableeng).

```{r regressiontableeng}
library(papaja)
apa_lm <- apa_print(eng)
apa_table(
  apa_lm$table, 
  caption = "Regression Predicting Engagement.",
  align = c('l','r','r','r','r','r'),
  placement = "H"
)
```

Stress was not predicted by total resources, challenge, or hindrance demands, $F_{(3, 564)}$ = 2.47, *p* = .060; see Table \@ref(tab:regressiontablestr).

```{r regressiontablestr}
library(papaja)
apa_lm <- apa_print(str)
apa_table(
  apa_lm$table, 
  placement = "H",
  align = c('l','r','r','r','r','r'),
  caption = "Regression Predicting Stress."
)
```

Similarly, burnout was not predicted by total resources, challenge, or hindrance demands, $F_{(3, 564)}$= 1.10, *p* = .349. See Table \@ref(tab:regressiontablebur).

```{r regressiontablebur}
library(papaja)
apa_lm <- apa_print(bur)
apa_table(
  apa_lm$table, 
  placement = "H",
  align = c('l','r','r','r','r','r'),
  caption = "Regression Predicting Burnout."
)
```

# Discussion

The major goal of this paper was to further explore the relationships among *total* perceived challenge demands, hindrance demands, and resources and outcomes of engagement, stress, and burnout. The results suggest a positive relationship between both resources and engagement (H1a) and challenge demands and engagement (H2a). Employers would benefit from understanding that at least the perception of having “more” resources and more challenge demands in a job is highly associated with reported engagement. While not a causal relationship, it points to the potential value of these kinds of employee support nonetheless.

## Limitations & Future Directions

Future study could consider additional or other O\*Net items, *kinds* of resources or challenges, additional outcomes (e.g., job satisfaction), or whether volume of resources and demands operationalized in this way are related to other behaviors. Finally, there was a moderate relationship between challenges and hindrances, which may be driven by job complexity – that is, some jobs simply have more elements. With more job elements come increased opportunities for those elements to be perceived as resources, challenges, or hindrances.

<center>
![](SIOP.png){height=350px width=1080px}
<center>
```{r bib, eval=FALSE}
knitr::write_bib(c('knitr','rmarkdown','posterdown','pagedown'), 'packages.bib')
```