### 2024-04-30 moderated_regressions.Rmd
updating tables

---
title: "moderated_regressions"
output: pdf_document
date: "2024-04-13"
---

```{r data, include = FALSE}
data <- read.csv("SIOPcleaned.csv")      ## from "datacleaning.R" script

## Study 2 DVs

data$item387      <- 7- data$item387           ## only reflected engagement item

data$absorption   <- rowMeans(data[c(380:386)], na.rm=TRUE)
data$vigor        <- rowMeans(data[c(387:392)], na.rm=TRUE) 
data$dedication   <- rowMeans(data[c(393:399)], na.rm=TRUE)

data$cognitive    <- rowMeans(data[c(380:382, 387:388, 393:395)], na.rm=TRUE)
data$affective    <- rowMeans(data[c(383:384, 389:390, 396:397)], na.rm=TRUE) 
data$behavioral   <- rowMeans(data[c(385:386, 391:392, 398:399)], na.rm=TRUE)

######################################################

data$burnout      <- rowMeans(data[c(373:376)], na.rm=TRUE)
data$stress       <- rowMeans(data[c(377:379)], na.rm=TRUE)
data$engagement   <- rowMeans(data[c(380:399)], na.rm=TRUE)        ## added 10/12 because analyses too wonky with 3 subscales

alpha.abs <- psych::alpha(data[c(380:386)])
alpha.vig <- psych::alpha(data[c(387:392)])
alpha.ded <- psych::alpha(data[c(393:399)])
alpha.eng <- psych::alpha(data[c(380:399)])

alpha.bur <- psych::alpha(data[c(373:376)])
alpha.str <- psych::alpha(data[c(377:379)])

## O*NET categories

data$onet.resource.ii <- rowMeans(data[c(162:166)], na.rm=TRUE)
data$onet.resource.mp <- rowMeans(data[c(167:176)], na.rm=TRUE)
data$onet.resource.wo <- rowMeans(data[c(177:185)], na.rm=TRUE)
data$onet.resource.io <- rowMeans(data[c(186:202)], na.rm=TRUE)
data$onet.resource.ir <- rowMeans(data[c(149:161)], na.rm=TRUE)
data$onet.resource.pc <- rowMeans(data[c(128:144,146:148)], na.rm=TRUE)
data$onet.resource.sc <- rowMeans(data[c(118:127)], na.rm=TRUE)

overall.resource <- rowMeans(data[c(118:144,147:202)], na.rm=TRUE)

data$onet.hindrance.ii <- rowMeans(data[c(246:247,249:252)], na.rm=TRUE)
data$onet.hindrance.mp <- rowMeans(data[c(252:261)], na.rm=TRUE)
data$onet.hindrance.wo <- rowMeans(data[c(262:270)], na.rm=TRUE)
data$onet.hindrance.io <- rowMeans(data[c(271:287)], na.rm=TRUE)
data$onet.hindrance.ir <- rowMeans(data[c(233:245)], na.rm=TRUE)
data$onet.hindrance.pc <- rowMeans(data[c(213:232)], na.rm=TRUE)
data$onet.hindrance.sc <- rowMeans(data[c(203:212)], na.rm=TRUE)

overall.hindrance <- rowMeans(data[c(203:247,249:287)], na.rm=TRUE)

data$onet.challenge.ii <- rowMeans(data[c(332:336)], na.rm=TRUE)
data$onet.challenge.mp <- rowMeans(data[c(337:346)], na.rm=TRUE)
data$onet.challenge.wo <- rowMeans(data[c(347:355)], na.rm=TRUE)
data$onet.challenge.io <- rowMeans(data[c(356:372)], na.rm=TRUE)
data$onet.challenge.ir <- rowMeans(data[c(319:331)], na.rm=TRUE)
data$onet.challenge.pc <- rowMeans(data[c(298:307,309:318)], na.rm=TRUE)
data$onet.challenge.sc <- rowMeans(data[c(288:297)], na.rm=TRUE)

overall.challenge <- rowMeans(data[c(288:307,309:372)], na.rm=TRUE)

```

```{r Analyses with Row Sums}
overall.resource <- rowSums(data[c(118:144,147:202)], na.rm=TRUE)
summary(overall.resource) #202.3
overall.hindrance <- rowSums(data[c(203:247,249:287)], na.rm=TRUE)
summary(overall.hindrance) #159.2
overall.challenge <- rowSums(data[c(288:307,309:372)], na.rm=TRUE)
summary(overall.challenge) #204.2

#centered predictors
overall.resource_center <- overall.resource - 202.3
overall.hindrance_center <- overall.hindrance - 159.2
overall.challenge_center <- overall.challenge - 204.2
```

```{r Moderated Regression Challenge-Resource Burnout}

cr_burn_m1 <- lm(burnout ~ overall.challenge_center, data = data)
summary(cr_burn_m1)

cr_burn_m2 <- lm(burnout ~ overall.challenge_center + overall.resource, data = data)
summary(cr_burn_m2)

cr_burn_m3 <- lm(burnout ~ overall.challenge_center + overall.resource_center + overall.challenge_center*overall.resource_center, data = data)
summary(cr_burn_m3)

aov_test <- anova(cr_burn_m2, cr_burn_m3)
```

```{r Moderated Regression Challenge-Resource Stress}

cr_stress_m1 <- lm(stress ~ overall.challenge_center, data = data)
summary(cr_stress_m1)

cr_stress_m2 <- lm(stress ~ overall.challenge_center + overall.resource_center, data = data)
summary(cr_stress_m2)

cr_stress_m3 <- lm(stress ~ overall.challenge_center + overall.resource_center + overall.challenge_center*overall.resource_center, data = data)
summary(cr_stress_m3)

anova(cr_stress_m2, cr_stress_m3)
```

```{r Moderated Regression Challenge-Resource Engagement}

cr_eng_m1 <- lm(engagement ~ overall.challenge_center, data = data)
summary(cr_eng_m1)

cr_eng_m2 <- lm(engagement ~ overall.challenge_center + overall.resource_center, data = data)
summary(cr_eng_m2)

cr_eng_m3 <- lm(engagement ~ overall.challenge_center + overall.resource_center + overall.challenge_center*overall.resource_center, data = data)
summary(cr_eng_m3)

anova(cr_eng_m2, cr_eng_m3)
```

```{r Moderated Regression Hindrance-Resource Burnout}

hr_burn_m1 <- lm(burnout ~ overall.hindrance_center, data = data)
summary(hr_burn_m1)

hr_burn_m2 <- lm(burnout ~ overall.hindrance_center + overall.resource_center, data = data)
summary(hr_burn_m2)

hr_burn_m3 <- lm(burnout ~ overall.hindrance_center + overall.resource_center + overall.hindrance_center*overall.resource_center, data = data)
summary(hr_burn_m3)

anova(hr_burn_m2, hr_burn_m3)
```

```{r Moderated Regression Hindrance-Resource Stress}

hr_stress_m1 <- lm(stress ~ overall.hindrance_center, data = data)
summary(hr_stress_m1)

hr_stress_m2 <- lm(stress ~ overall.hindrance_center + overall.resource_center, data = data)
summary(hr_stress_m2)

hr_stress_m3 <- lm(stress ~ overall.hindrance_center + overall.resource_center + overall.hindrance_center*overall.resource_center, data = data)
summary(hr_stress_m3)

anova(hr_stress_m2, hr_stress_m3)
```

```{r Moderated Regression Hindrance-Resource Engagement}

hr_eng_m1 <- lm(engagement ~ overall.hindrance_center, data = data)
summary(hr_eng_m1)

hr_eng_m2 <- lm(engagement ~ overall.hindrance_center + overall.resource_center, data = data)
summary(hr_eng_m2)

hr_eng_m3 <- lm(engagement ~ overall.hindrance_center + overall.resource_center + overall.hindrance_center*overall.resource_center, data = data)
summary(hr_eng_m3)

anova(hr_eng_m2, hr_eng_m3)
```

```{r Numbers for Table1}

options(scipen=999)  ## Suppresses scientific notation

#Challenge-Resource-Burnout

# b weights
chal.2.burn <- round(summary(cr_burn_m2)$coefficients[2, 1],2)
res.2.burn <- round(summary(cr_burn_m2)$coefficients[3, 1],2)

int.cr.3.burn <- round(summary(cr_burn_m3)$coefficients[4, 1],2)

#r squared
cr_burn_m2_r <- round(summary(cr_burn_m2)$r.squared,2)
cr_burn_m3_r <- round(summary(cr_burn_m3)$r.squared,2)

## R square change
cr_burn_anova <- anova(cr_burn_m2, cr_burn_m3)
cr_burn_r_square_change <- round(cr_burn_m3_r - cr_burn_m2_r,2)

#Challenge-Resource-Stress

# b weights
chal.2.stress <- round(summary(cr_stress_m2)$coefficients[2, 1],2)
res.2.stress <- round(summary(cr_stress_m2)$coefficients[3, 1],2)

int.cr.3.stress <- round(summary(cr_stress_m3)$coefficients[4, 1],2)

#r squared
cr_stress_m2_r <- round(summary(cr_stress_m2)$r.squared,2)
cr_stress_m3_r <- round(summary(cr_stress_m3)$r.squared,2)

## R square change
cr_stress_anova <- anova(cr_stress_m2, cr_stress_m3)
cr_stress_r_square_change <- round(cr_stress_m3_r - cr_stress_m2_r,2)

#Challenge-Resource-Engagement

# b weights
chal.2.eng <- round(summary(cr_eng_m2)$coefficients[2, 1],2)
res.2.eng <- round(summary(cr_eng_m2)$coefficients[3, 1],2)

int.cr.3.eng <- round(summary(cr_eng_m3)$coefficients[4, 1],2)

#r squared
cr_eng_m2_r <- round(summary(cr_eng_m2)$r.squared,2)
cr_eng_m3_r <- round(summary(cr_eng_m3)$r.squared,2)

## R square change
cr_eng_anova <- anova(cr_eng_m2, cr_eng_m3)
cr_eng_r_square_change <- round(cr_eng_m3_r - cr_eng_m2_r,2)
```

```{r Numbers for Table2}

options(scipen=999)  ## Suppresses scientific notation

#Hindrance-Resource-Burnout

# b weights
hind.2.burn <- round(summary(hr_burn_m2)$coefficients[2, 1],2)
res.2hr.burn <- round(summary(hr_burn_m2)$coefficients[3, 1],2)

int.hr.3.burn <- round(summary(hr_burn_m3)$coefficients[4, 1],2)

#r squared
hr_burn_m2_r <- round(summary(hr_burn_m2)$r.squared,2)
hr_burn_m3_r <- round(summary(hr_burn_m3)$r.squared,2)

## R square change
hr_burn_anova <- anova(hr_burn_m2, hr_burn_m3)
hr_burn_r_square_change <- round(hr_burn_m3_r - hr_burn_m2_r,2)

#Hindrance-Resource-Stress

# b weights
hind.2.stress <- round(summary(hr_stress_m2)$coefficients[2, 1],2)
res.2hr.stress <- round(summary(hr_stress_m2)$coefficients[3, 1],2)

int.hr.3.stress <- round(summary(hr_stress_m3)$coefficients[4, 1],2)

#r squared
hr_stress_m2_r <- round(summary(hr_stress_m2)$r.squared,2)
hr_stress_m3_r <- round(summary(hr_stress_m3)$r.squared,2)

## R square change
hr_stress_anova <- anova(hr_stress_m2, hr_stress_m3)
hr_stress_r_square_change <- round(hr_stress_m3_r - hr_stress_m2_r,2)

#Hindrance-Resource-Engagement

# b weights
hind.2.eng <- round(summary(hr_eng_m2)$coefficients[2, 1],2)
res.2hr.eng <- round(summary(hr_eng_m2)$coefficients[3, 1],2)

int.hr.3.eng <- round(summary(hr_eng_m3)$coefficients[4, 1],2)

#r squared
hr_eng_m2_r <- round(summary(hr_eng_m2)$r.squared,2)
hr_eng_m3_r <- round(summary(hr_eng_m3)$r.squared,2)

## R square change
hr_eng_anova <- anova(hr_eng_m2, hr_eng_m3)
hr_eng_r_square_change <- round(hr_eng_m3_r - hr_eng_m2_r,2)
```

```{r, results='asis'}
#I'm not touching this script - using it as a model and making two new chunks
library(kableExtra)

DV <- c("Burnout","","")
Step <- c("1", "","2")
Model <- c("Challenge", "Resource", "Challenge X Resource")
b <- c(chal.2.burn,res.2.burn,int.cr.3.burn)
DeltaR <- c("",paste(cr_burn_m2_r,"**"),cr_burn_r_square_change)

regtable <- cbind(DV, Step, Model, b, DeltaR)

papaja::apa_table(regtable)

```

```{r, results='asis'}

library(kableExtra)

DV <- c("Engagement","","", "Stress","","", "Burnout","","")
Step <- c("1", "","2", "1", "","2", "1", "","2")
Model <- c("Challenge", "Resource", "Challenge X Resource", "Challenge", "Resource", "Challenge X Resource", "Challenge", "Resource", "Challenge X Resource")
b <- c(chal.2.eng, res.2.eng, int.cr.3.eng, chal.2.stress, res.2.stress, int.cr.3.stress, chal.2.burn, res.2.burn, int.cr.3.burn)
R2 <- c("", cr_eng_m2_r, cr_eng_m3_r, "", cr_stress_m2_r, cr_stress_m3_r, "", cr_burn_m2_r, cr_burn_m3_r)
DeltaR <- c("",paste(cr_eng_m2_r),cr_eng_r_square_change, "",paste(cr_stress_m2_r),cr_stress_r_square_change, "",paste(cr_burn_m2_r),cr_burn_r_square_change)

regtable <- cbind(DV, Step, Model, b, R2, DeltaR)

papaja::apa_table(regtable)

```

```{r, results='asis'}

library(kableExtra)

DV <- c("Engagement","","", "Stress","","", "Burnout","","")
Step <- c("1", "","2", "1", "","2", "1", "","2")
Model <- c("Hindrance", "Resource", "Hindrance X Resource", "Hindrance", "Resource", "Hindrance X Resource", "Hindrance", "Resource", "Hindrance X Resource")
b <- c(hind.2.eng, res.2hr.eng, int.hr.3.eng, hind.2.stress, res.2hr.stress, int.hr.3.stress, hind.2.burn, res.2hr.burn, int.hr.3.burn)
R2 <- c("", hr_eng_m2_r, hr_eng_m3_r, "", hr_stress_m2_r, hr_stress_m3_r, "", hr_burn_m2_r, hr_burn_m3_r)
DeltaR <- c("",paste(hr_eng_m2_r,"**"),hr_eng_r_square_change, "",paste(hr_stress_m2_r,"**"),hr_stress_r_square_change, "",paste(hr_burn_m2_r,"**"),hr_burn_r_square_change)

regtable <- cbind(DV, Step, Model, b, R2, DeltaR)

papaja::apa_table(regtable)
# How to get delta symbol and superscript
#https://cran.r-project.org/web/packages/reporter/vignettes/reporter-super.html#:~:text=To%20get%20superscripts%20or%20subscripts,to%20the%20desired%20cell%20value.

```