### 2024-05-21 moderated_regressions.Rmd
updating results section

---
title: "moderated_regressions"
output: 
  pdf_document:
    latex_engine: xelatex
date: "2024-04-13"
---

```{r modreg, include = FALSE}

sigfill <- function(x, sigfigs = 3){
  out <- gsub("\\.$", "",
              formatC(signif(x, digits = sigfigs),
                      digits = sigfigs, format = "fg", flag = "#"))
  out[grepl(".", out, fixed = TRUE)] <- strtrim(out[grepl(".", out, fixed = TRUE)],
                                                sigfigs + c(1, 2)[grepl("-", out, fixed = TRUE) + 1])
  return(out)
}

modreg <- read.csv("SIOPcleaned.csv")      ## from "datacleaning.R" script

## Study 2 DVs

modreg$item387      <- 7- modreg$item387           ## only reflected engagement item

modreg$absorption   <- rowMeans(modreg[c(380:386)], na.rm=TRUE)
modreg$vigor        <- rowMeans(modreg[c(387:392)], na.rm=TRUE) 
modreg$dedication   <- rowMeans(modreg[c(393:399)], na.rm=TRUE)

modreg$cognitive    <- rowMeans(modreg[c(380:382, 387:388, 393:395)], na.rm=TRUE)
modreg$affective    <- rowMeans(modreg[c(383:384, 389:390, 396:397)], na.rm=TRUE) 
modreg$behavioral   <- rowMeans(modreg[c(385:386, 391:392, 398:399)], na.rm=TRUE)

######################################################

modreg$burnout      <- rowMeans(modreg[c(373:376)], na.rm=TRUE)
modreg$stress       <- rowMeans(modreg[c(377:379)], na.rm=TRUE)
modreg$engagement   <- rowMeans(modreg[c(380:399)], na.rm=TRUE)        ## added 10/12 because analyses too wonky with 3 subscales

alpha.abs <- psych::alpha(modreg[c(380:386)])
alpha.vig <- psych::alpha(modreg[c(387:392)])
alpha.ded <- psych::alpha(modreg[c(393:399)])
alpha.eng <- psych::alpha(modreg[c(380:399)])

alpha.bur <- psych::alpha(modreg[c(373:376)])
alpha.str <- psych::alpha(modreg[c(377:379)])

## O*NET categories

modreg$onet.resource.ii <- rowMeans(modreg[c(162:166)], na.rm=TRUE)
modreg$onet.resource.mp <- rowMeans(modreg[c(167:176)], na.rm=TRUE)
modreg$onet.resource.wo <- rowMeans(modreg[c(177:185)], na.rm=TRUE)
modreg$onet.resource.io <- rowMeans(modreg[c(186:202)], na.rm=TRUE)
modreg$onet.resource.ir <- rowMeans(modreg[c(149:161)], na.rm=TRUE)
modreg$onet.resource.pc <- rowMeans(modreg[c(128:144,146:148)], na.rm=TRUE)
modreg$onet.resource.sc <- rowMeans(modreg[c(118:127)], na.rm=TRUE)

overall.resource <- rowMeans(modreg[c(118:144,147:202)], na.rm=TRUE)

modreg$onet.hindrance.ii <- rowMeans(modreg[c(246:247,249:252)], na.rm=TRUE)
modreg$onet.hindrance.mp <- rowMeans(modreg[c(252:261)], na.rm=TRUE)
modreg$onet.hindrance.wo <- rowMeans(modreg[c(262:270)], na.rm=TRUE)
modreg$onet.hindrance.io <- rowMeans(modreg[c(271:287)], na.rm=TRUE)
modreg$onet.hindrance.ir <- rowMeans(modreg[c(233:245)], na.rm=TRUE)
modreg$onet.hindrance.pc <- rowMeans(modreg[c(213:232)], na.rm=TRUE)
modreg$onet.hindrance.sc <- rowMeans(modreg[c(203:212)], na.rm=TRUE)

overall.hindrance <- rowMeans(modreg[c(203:247,249:287)], na.rm=TRUE)

modreg$onet.challenge.ii <- rowMeans(modreg[c(332:336)], na.rm=TRUE)
modreg$onet.challenge.mp <- rowMeans(modreg[c(337:346)], na.rm=TRUE)
modreg$onet.challenge.wo <- rowMeans(modreg[c(347:355)], na.rm=TRUE)
modreg$onet.challenge.io <- rowMeans(modreg[c(356:372)], na.rm=TRUE)
modreg$onet.challenge.ir <- rowMeans(modreg[c(319:331)], na.rm=TRUE)
modreg$onet.challenge.pc <- rowMeans(modreg[c(298:307,309:318)], na.rm=TRUE)
modreg$onet.challenge.sc <- rowMeans(modreg[c(288:297)], na.rm=TRUE)

overall.challenge <- rowMeans(modreg[c(288:307,309:372)], na.rm=TRUE)

```

```{r Analyses with Row Sums}
overall.resource <- rowSums(modreg[c(118:144,147:202)], na.rm=TRUE)
summary(overall.resource) #202.3
overall.hindrance <- rowSums(modreg[c(203:247,249:287)], na.rm=TRUE)
summary(overall.hindrance) #159.2
overall.challenge <- rowSums(modreg[c(288:307,309:372)], na.rm=TRUE)
summary(overall.challenge) #204.2

#centered predictors
overall.resource_center <- overall.resource - 202.3
overall.hindrance_center <- overall.hindrance - 159.2
overall.challenge_center <- overall.challenge - 204.2
```

```{r Moderated Regression Challenge-Resource Burnout}

cr_burn_m1 <- lm(scale(burnout) ~ scale(overall.challenge_center), modreg)
summary(cr_burn_m1)

cr_burn_m2 <- lm(scale(burnout) ~ scale(overall.challenge_center) + scale(overall.resource), modreg)
summary(cr_burn_m2)

cr_burn_m3 <- lm(scale(burnout) ~ scale(overall.challenge_center) + scale(overall.resource_center) + scale(overall.challenge_center*overall.resource_center), modreg)
summary(cr_burn_m3)

aov_test <- anova(cr_burn_m2, cr_burn_m3)
aov_test
```

```{r Moderated Regression Challenge-Resource Stress}

cr_stress_m1 <- lm(scale(stress) ~ scale(overall.challenge_center), modreg)
summary(cr_stress_m1)

cr_stress_m2 <- lm(scale(stress) ~ scale(overall.challenge_center) + scale(overall.resource_center), modreg)
summary(cr_stress_m2)

cr_stress_m3 <- lm(scale(stress) ~ scale(overall.challenge_center) + scale(overall.resource_center) + scale(overall.challenge_center*overall.resource_center), modreg)
summary(cr_stress_m3)

cr_stress_anova <- anova(cr_stress_m2, cr_stress_m3)
cr_stress_anova
```

```{r Moderated Regression Challenge-Resource Engagement}

cr_eng_m1 <- lm(scale(engagement) ~ scale(overall.challenge_center), modreg)
summary(cr_eng_m1)

cr_eng_m2 <- lm(scale(engagement) ~ scale(overall.challenge_center) + scale(overall.resource_center), modreg)
summary(cr_eng_m2)

cr_eng_m3 <- lm(scale(engagement) ~ scale(overall.challenge_center) + scale(overall.resource_center) + scale(overall.challenge_center*overall.resource_center), modreg)
summary(cr_eng_m3)

cr_eng_anova <- anova(cr_eng_m2, cr_eng_m3)
cr_eng_anova
```

```{r Moderated Regression Hindrance-Resource Burnout}

hr_burn_m1 <- lm(scale(burnout) ~ scale(overall.hindrance_center), modreg)
summary(hr_burn_m1)

hr_burn_m2 <- lm(scale(burnout) ~ scale(overall.hindrance_center) + scale(overall.resource_center), modreg)
summary(hr_burn_m2)

hr_burn_m3 <- lm(scale(burnout) ~ scale(overall.hindrance_center) + scale(overall.resource_center) + scale(overall.hindrance_center*overall.resource_center), modreg)
summary(hr_burn_m3)

hr_burn_anova <- anova(hr_burn_m2, hr_burn_m3)
hr_burn_anova
```

```{r Moderated Regression Hindrance-Resource Stress}

hr_stress_m1 <- lm(scale(stress) ~ scale(overall.hindrance_center), modreg)
summary(hr_stress_m1)

hr_stress_m2 <- lm(scale(stress) ~ scale(overall.hindrance_center) + scale(overall.resource_center), modreg)
summary(hr_stress_m2)

hr_stress_m3 <- lm(scale(stress) ~ scale(overall.hindrance_center) + scale(overall.resource_center) + scale(overall.hindrance_center*overall.resource_center), modreg)
summary(hr_stress_m3)

hr_stress_anova <- anova(hr_stress_m2, hr_stress_m3)
hr_stress_anova
```

```{r Moderated Regression Hindrance-Resource Engagement}

hr_eng_m1 <- lm(scale(engagement) ~ scale(overall.hindrance_center), modreg)
summary(hr_eng_m1)

hr_eng_m2 <- lm(scale(engagement) ~ scale(overall.hindrance_center) + scale(overall.resource_center), modreg)
summary(hr_eng_m2)

hr_eng_m3 <- lm(scale(engagement) ~ scale(overall.hindrance_center) + scale(overall.resource_center) + scale(overall.hindrance_center*overall.resource_center), modreg)
summary(hr_eng_m3)

hr_eng_anova <- anova(hr_eng_m2, hr_eng_m3)
hr_eng_anova
```

```{r Numbers for Table1}

options(scipen=999)  ## Suppresses scientific notation

#Challenge-Resource-Burnout

# b weights
chal.2.burn <- round(summary(cr_burn_m2)$coefficients[2, 1],2)
res.2.burn <- round(summary(cr_burn_m2)$coefficients[3, 1],2)

int.cr.3.burn <- round(summary(cr_burn_m3)$coefficients[4, 1],2)

#r squared
cr_burn_m2_r <- summary(cr_burn_m2)$r.squared
cr_burn_m3_r <- summary(cr_burn_m3)$r.squared

## R square change
cr_burn_anova <- anova(cr_burn_m2, cr_burn_m3)
cr_burn_r_square_change <- sigfill(cr_burn_m3_r - cr_burn_m2_r)

#Challenge-Resource-Stress

# b weights
chal.2.stress <- round(summary(cr_stress_m2)$coefficients[2, 1],2)
res.2.stress <- round(summary(cr_stress_m2)$coefficients[3, 1],2)

int.cr.3.stress <- round(summary(cr_stress_m3)$coefficients[4, 1],2)

#r squared
cr_stress_m2_r <- summary(cr_stress_m2)$r.squared
cr_stress_m3_r <- summary(cr_stress_m3)$r.squared

## R square change
cr_stress_anova <- anova(cr_stress_m2, cr_stress_m3)
cr_stress_r_square_change <- sigfill(cr_stress_m3_r - cr_stress_m2_r)

#Challenge-Resource-Engagement

# b weights
chal.2.eng <- round(summary(cr_eng_m2)$coefficients[2, 1],2)
res.2.eng <- round(summary(cr_eng_m2)$coefficients[3, 1],2)

int.cr.3.eng <- round(summary(cr_eng_m3)$coefficients[4, 1],2)

#r squared
cr_eng_m2_r <- round(summary(cr_eng_m2)$r.squared,2)
cr_eng_m3_r <- round(summary(cr_eng_m3)$r.squared,2)

## R square change
cr_eng_anova <- anova(cr_eng_m2, cr_eng_m3)
cr_eng_r_square_change <- round(cr_eng_m3_r - cr_eng_m2_r,2)
```

```{r Numbers for Table2}

options(scipen=999)  ## Suppresses scientific notation

#Hindrance-Resource-Burnout

# b weights
hind.2.burn <- round(summary(hr_burn_m2)$coefficients[2, 1],2)
res.2hr.burn <- round(summary(hr_burn_m2)$coefficients[3, 1],2)

int.hr.3.burn <- round(summary(hr_burn_m3)$coefficients[4, 1],2)

#r squared
hr_burn_m2_r <- round(summary(hr_burn_m2)$r.squared,2)
hr_burn_m3_r <- round(summary(hr_burn_m3)$r.squared,2)

## R square change
hr_burn_anova <- anova(hr_burn_m2, hr_burn_m3)
hr_burn_r_square_change <- round(hr_burn_m3_r - hr_burn_m2_r,2)

#Hindrance-Resource-Stress

# b weights
hind.2.stress <- round(summary(hr_stress_m2)$coefficients[2, 1],2)
res.2hr.stress <- round(summary(hr_stress_m2)$coefficients[3, 1],2)

int.hr.3.stress <- round(summary(hr_stress_m3)$coefficients[4, 1],2)

#r squared
hr_stress_m2_r <- round(summary(hr_stress_m2)$r.squared,2)
hr_stress_m3_r <- round(summary(hr_stress_m3)$r.squared,2)

## R square change
hr_stress_anova <- anova(hr_stress_m2, hr_stress_m3)
hr_stress_r_square_change <- round(hr_stress_m3_r - hr_stress_m2_r,2)

#Hindrance-Resource-Engagement

# b weights
hind.2.eng <- round(summary(hr_eng_m2)$coefficients[2, 1],2)
res.2hr.eng <- round(summary(hr_eng_m2)$coefficients[3, 1],2)

int.hr.3.eng <- round(summary(hr_eng_m3)$coefficients[4, 1],2)

#r squared
hr_eng_m2_r <- summary(hr_eng_m2)$r.squared
hr_eng_m3_r <- summary(hr_eng_m3)$r.squared

## R square change
hr_eng_anova <- anova(hr_eng_m2, hr_eng_m3)
hr_eng_r_square_change <- sigfill(hr_eng_m3_r - hr_eng_m2_r)
```

```{r, results='asis'}
#I'm not touching this script - using it as a model and making two new chunks
library(kableExtra)

DV <- c("Burnout","","")
Step <- c("1", "","2")
Model <- c("Challenge", "Resource", "Challenge X Resource")
b <- c(chal.2.burn,res.2.burn,int.cr.3.burn)
DeltaR <- c("",paste(cr_burn_m2_r,"**"),cr_burn_r_square_change)

regtable <- cbind(DV, Step, Model, b, DeltaR)

papaja::apa_table(regtable)

```

```{r, results='asis'}

library(kableExtra)

DV <- c("Engagement","","", "Stress","","", "Burnout","","")
Step <- c("1", "","2", "1", "","2", "1", "","2")
Model <- c("Challenge", "Resource", "Challenge X Resource", "Challenge", "Resource", "Challenge X Resource", "Challenge", "Resource", "Challenge X Resource")
Beta <- c(sigfill(chal.2.eng), paste(res.2.eng,"**"), paste(int.cr.3.eng,"*"), chal.2.stress, res.2.stress, int.cr.3.stress, paste(chal.2.burn,"**"), res.2.burn, int.cr.3.burn)
R2 <- c("", paste(cr_eng_m2_r,"**"), paste(cr_eng_m3_r,"**"), "", round(cr_stress_m2_r,2), round(cr_stress_m3_r,2), "", paste(round(cr_burn_m2_r,2),"**"), paste(round(cr_burn_m3_r,2),"**"))
DeltaR <- c("","", paste(cr_eng_r_square_change,"*"), "","",cr_stress_r_square_change, "","",cr_burn_r_square_change)

regtable <- as.data.frame(cbind(DV, Step, Model, Beta, R2, DeltaR))

colnames(regtable)[4] <- "$\\beta$"
colnames(regtable)[5] <- "$R^2$"
colnames(regtable)[6] <- "$\\Delta R^2$"

papaja::apa_table(regtable, 
                 escape=FALSE,
                note="* = p < .05; ** = p < .01")
```

```{r, results='asis'}

library(kableExtra)

DV <- c("Engagement","","", "Stress","","", "Burnout","","")
Step <- c("1", "","2", "1", "","2", "1", "","2")
Model <- c("Hindrance", "Resource", "Hindrance X Resource", "Hindrance", "Resource", "Hindrance X Resource", "Hindrance", "Resource", "Hindrance X Resource")
Beta <- c(paste(hind.2.eng,"**"), paste(res.2hr.eng,"**"), int.hr.3.eng, paste(sigfill(hind.2.stress),"*"), paste(res.2hr.stress,"**"), paste(int.hr.3.stress,"**"), paste(sigfill(hind.2.burn),"*"), paste(res.2hr.burn,"*"), paste(int.hr.3.burn,"**"))
R2 <- c("", paste(round(hr_eng_m2_r,2),"**"), paste(round(hr_eng_m3_r,2),"**"), "", paste(hr_stress_m2_r,"*"), paste(hr_stress_m3_r, "**"), "", paste(hr_burn_m2_r,"**"), paste(hr_burn_m3_r,"**"))
#DeltaR <- c("",paste(hr_eng_m2_r,"**"),hr_eng_r_square_change, "",paste(hr_stress_m2_r,"**"),hr_stress_r_square_change, "",paste(hr_burn_m2_r,"**"),hr_burn_r_square_change)
DeltaR <- c("","",hr_eng_r_square_change, "","",paste(hr_stress_r_square_change,"**"), "","",paste(hr_burn_r_square_change, "**"))
regtable <- as.data.frame(cbind(DV, Step, Model, Beta, R2, DeltaR))

colnames(regtable)[4] <- "$\\beta$"
colnames(regtable)[5] <- "$R^2$"
colnames(regtable)[6] <- "$\\Delta R^2$"

#kableExtra::kable(regtable,
#                  format="latex",
#                  escape=FALSE)

papaja::apa_table(regtable,
                 escape=FALSE,
                 note="* = p < .05; ** = p < .01")
```

```{r fig.cap="Interaction between Challenge and Resources on Engagement"}
#install.packages("interactions")
library(interactions)

#running it with unstandardized bs here. 
cr_eng_m3u <- lm(engagement ~ overall.challenge_center + overall.resource_center + overall.challenge_center*overall.resource_center, modreg)
summary(cr_eng_m3u)

cr_eng_plot <- interact_plot(cr_eng_m3u, pred = overall.challenge_center, modx = overall.resource_center, interval = FALSE,
              int.width = 0.8)
cr_eng_plot
```

```{r fig.cap="Interaction between Hindrances and Resources on Stress"}
#install.packages("interactions")
library(interactions)

hr_stress_m3u <- lm(stress ~ overall.hindrance_center + overall.resource_center + overall.hindrance_center*overall.resource_center, modreg)
summary(hr_stress_m3u)

hr_stress_plot <- interact_plot(hr_stress_m3u, pred = overall.hindrance_center, modx = overall.resource_center, interval = FALSE, 
              int.width = 0.8)
hr_stress_plot
```

```{r fig.cap="Interaction between Hindrances and Resources on Burnout"}
#install.packages("interactions")
library(interactions)

hr_burn_m3u <- lm(burnout ~ overall.hindrance_center + overall.resource_center + overall.hindrance_center*overall.resource_center, modreg)
summary(hr_burn_m3u)

hr_burn_plot <- interact_plot(hr_burn_m3u, pred = overall.hindrance_center, modx = overall.resource_center, interval = FALSE, 
              int.width = 0.8)
hr_burn_plot
```